---
title: "Analysis"
author: "August E. G. Mikkelsen"
date: "`r Sys.Date()`"
output: pdf_document
header-includes:
  - \usepackage{xcolor}
---

This notebook provides a walkthrough of the analysis and results described in several papers, which as of June 2026 are: 

1) "Generalized Additive Models for Mass Appraisal: Evidence from the Danish Public Property Valuation System" (in review)
2) "Property and Land value Predictions with Generalized Additive Models" (in preparation)
3) "On the Accuracy of Hedonic Real Estate Models" (in preparation)

# 1. Reproducibility and setup 

```{r setup, eval=TRUE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggpubr)
library(mgcv)
library(gratia) 
library(knitr)
library(patchwork)
library(lubridate)
library(future)
library(future.apply)
library(rlang)
library(stringr)
library(furrr)
library(xgboost)
library(caret)
library(Hmisc)
library(kableExtra)
library(nabor)
library(purrr)
library(ggrastr)
library(ggridges)
library(gridExtra)
library(forcats)
library(pheatmap)
library(viridis)
library(viridisLite)
library(cowplot)
library(scico)
library(scales)

# Global parameters
set.seed(123) # for reproducibility
N <- 4 # number of CV folds
k_spatial = 2000 # initial number of basis functions used in the spatial smooths
k_other = 10 # initial number of basis functions used for other smooths
load <- FALSE # determines whether or not to preload models in the various chunks
plan(multisession, workers = N) #do as much as possible in parallel
options(future.globals.maxSize = 10 * 1024^3) #increase maxSize for future to handle large models

#Create a saved files folder and checkpoints folder is these do not already exist
for (f in c("saved_files", "checkpoints")) {
  if (!dir.exists(f)) {
    dir.create(f, recursive = TRUE)
    message(sprintf("Created folder: %s", f))
  } else {
    message(sprintf("Folder already exists: %s", f))
  }
}
```



# 2. Data 

## 2.1 The dataset
The analysis is based on a dataset of single-family home sales, single-family lot sales and properties in Denmark from 1 January 2014 to 1 January 2024. The dataset, included as a .rds file in this repository, is loaded and inspected below.

```{r Load dataset, eval=TRUE, echo=TRUE, warning=FALSE}

#Externally defined coordinate grid of Denmark
coordinate_grid <- read.csv2("coordinates_ocean_land.csv")

#Load data set
dataset <- readRDS("dataset.rds")

#The three types included in the dataset. 
table(dataset$type)

#Distribution of sales prices
plot_sales_price_distribution <- ggplot(dataset %>% filter(type %in% c("Lot sale", "Property sale")),
                                 aes(x = sales_price / 1e6,
                                     fill = type,
                                     color = type)) +
                                 geom_density(linewidth = 1,
                                              alpha = 0.3,
                                              adjust = 1.5,
                                              na.rm = TRUE) +
                                 coord_cartesian(xlim = c(0, 15)) +
                                 labs(title = "Sales price distribution",
                                      x = "Sales price [million DKK]",
                                      y = "Density",
                                      fill = NULL,
                                      color = NULL) +
                                 theme(legend.position = c(0.8, 0.8)) + 
                                 theme_minimal()

plot_sales_price_distribution

plot_geographic <- ggplot(dataset, aes(coordinate_east, coordinate_north)) +
                   geom_contour(data = coordinate_grid,
                                aes(x = coordinate_east,
                                    y = coordinate_north,
                                    z = as.numeric(!ocean)),
                                inherit.aes = FALSE,
                                color = "black",
                                linewidth = 0.1) +
                   geom_point(size = 0.2, alpha = 0.25, color = "steelblue") +
                   coord_fixed() +
                   facet_wrap(~ type) +
                   theme_void(base_size = 10)

plot_geographic

```
## 2.2 Descriptive statistics
In the subsequent cells histograms of selected variables are produced. These histograms are for the raw variables before any imputations and variable conversions take place.

```{r Continuous variables, eval=TRUE, echo=TRUE, warning=FALSE, fig.width=20, fig.height=30}

get_label <- function(var, labels_vec) {
  lbl <- tryCatch(labels_vec[[var]], error = function(e) NULL)
  if (is.null(lbl) || is.na(lbl)) var else lbl
}

plot_one <- function(df, var, label) {
  
  x <- df[[var]]
  
  if (label == "Sales date") {
    x_date <- as.Date(x)
    x <- decimal_date(x_date)
  }
  
  na_count <- sum(is.na(x))
  zero_count <- sum(x == 0, na.rm = TRUE)
  na_label <- paste0("Number of NA-values: ", format(na_count, big.mark = ",", scientific = FALSE))
  zero_label <- paste0("Number of zero-values: ", format(zero_count, big.mark = ",", scientific = FALSE))

  if (is.numeric(x)) {
    
    df2 <- tibble(x = x) %>% filter(is.finite(x) & x != 0)

    rng <- range(df2$x, na.rm = TRUE)
    
    rng_label <- paste0("Range: [", format(rng[1], big.mark = "", scientific = FALSE, digits = 2, trim = TRUE), " , ",
                                    format(rng[2], big.mark = "", scientific = FALSE, digits = 2, trim = TRUE), "]")

    x_pos = (0.75*min(x, na.rm = TRUE) + 0.25*max(x, na.rm = TRUE))
    
    ggplot(df2, aes(x = x)) +
    geom_histogram(bins = 30, fill = "#3182bd", color = "white") +
    labs(x = label, y = NULL) +
    annotate("text", x = x_pos, y = Inf, label = na_label, hjust = 0, vjust = 0, size = 4.5, color = "grey25", fontface = "bold") +
    annotate("text", x = x_pos, y = Inf, label = zero_label, hjust = 0, vjust = 1.5, size = 4.5, color = "grey25", fontface = "bold") +
    annotate("text", x = x_pos, y = Inf, label = rng_label, hjust = 0, vjust = 3, size = 4.5, color = "grey25", fontface = "bold") +
    coord_cartesian(clip = "off") +
    theme(text = element_text(size = 12),
          plot.margin = margin(t = 28, r = 12, b = 5, l = 5),
          panel.grid.minor = element_blank(),
          panel.grid.major.x = element_blank())

  } else {
    
    x2 <- fct_lump_n(factor(x), n = 30, other_level = "Other")
    
    df_counts <- tibble(x = x2) %>%
                 filter(!is.na(x)) %>%
                 count(x, name = "n")
    
    x_pos = (nlevels(df_counts$x) + 1) / 2
    
    ggplot(df_counts, aes(x = x, y = n)) +
    geom_col(aes(fill = x), color = "white", show.legend = FALSE) +
    scale_fill_brewer(palette = "Set3") +
    geom_text(aes(label = comma(n)), vjust = -0.2, size = 3.3, color = "grey15") +
    labs(x = NULL, y = NULL) +
    ggtitle(label) + 
    scale_y_continuous(expand = expansion(mult = c(0.02, 0.12))) +
    coord_cartesian(clip = "off") +
    theme(plot.margin = margin(t = 20, r = 12, b = 5, l = 5),
          axis.text.x = element_text(angle = 45, hjust = 1),
          panel.grid.minor = element_blank(),
          panel.grid.major.x = element_blank())}
}


variables_continuous <- c(sales_price = "Sales price [DKK]",
                          year_built = "Year built",
                          year_rebuilt = "Year rebuilt",
                          living_space = "Living space [m²]",
                          lot_size = "Lot size [m²]",
                          coast_distance = "Distance to coast [m]",
                          lake_distance = "Distance to lake [m]",
                          stream_distance = "Distance to stream [m]",
                          road_distance = "Distance to road [m]",
                          highway_distance = "Distance to highway [m]",
                          train_station_distance = "Distance to train station [m]",
                          railway_distance = "Distance to railway [m]",
                          windturbine_distance = "Distance to wind turbine [m]",
                          ocean_view = "Ocean view [m]",
                          lake_view = "Lake view [m]",
                          forest_size = "Forest size [ha]",
                          number_of_bathrooms = "Number of bathrooms",
                          number_of_rooms = "Number of rooms",
                          number_of_floors = "Number of floors",
                          sales_date = "Sales date")

vars_cont <- intersect(names(variables_continuous), names(dataset))

histograms_cont <- map(vars_cont, ~ plot_one(dataset, .x, get_label(.x, variables_continuous)))

saveRDS(histograms_cont, "saved_files/histograms_continuous.rds")
wrap_plots(histograms_cont, ncol = 3)
```

```{r Categorical variables, eval=TRUE, echo=TRUE, warning=FALSE, fig.width=20, fig.height=30}

variables_categorical <- c(house_type = "House type",
                           roof_type = "Roof type",
                           wall_type = "Wall type",
                           heating_type = "Heating type",
                           municipality_type = "Municipality type",
                           province = "Province",
                           region = "Region",
                           zone = "Zone")

vars_cat <- intersect(names(variables_categorical), names(dataset))

histograms_cat <- map(vars_cat, ~ plot_one(dataset, .x, get_label(.x, variables_categorical)))

saveRDS(histograms_cat, "saved_files/histograms_categorical.rds")

wrap_plots(histograms_cat, ncol = 3)

```


## 2.3 Transformation and imputation
Prior to model fitting a set of variable conversions and imputations are performed in order to: 

1) Convert categorical variables to factors
2) Handle missing values in the dataset
3) Create auxiliary variables

Missing variables with NA-entries are imputed based on the following rules: 

- year_rebuilt: For this variable all NA's as well as values less than 1970 are replaced by 1970. 
- maximum_coverage_percentage: For this variable all NA's are imputed to 30% corresponding to the most typical allowed building coverage percentage in Denmark.
- subdivision_size: For this variable all NA's are imputed to 700m2 corresponding to the most typical allowed subdivision size in Denmark. 
- distance-variables: For these variables all NA's are replaced by the maximum value of the respective variable. 

Furthermore, the following auxiliary variables are created: 

- Numeric sales date: A numeric version of the sales date spanning the interval [2014.0, 2024.0]
- Number of subdivisions: Defined as lot size / subdivision size rounded down to nearest integer and bounded at 1. 

The code for the above transformations is provided below. 

```{r Variable conversion and imputation, eval=TRUE, echo=TRUE, warning=FALSE, results= 'hide'}

### Factor conversion
dataset$municipality_type <- relevel(factor(dataset$municipality_type), ref = "Provincial")
dataset$house_type <- relevel(factor(dataset$house_type), ref = "Detached")
dataset$roof_type <- relevel(factor(dataset$roof_type), ref = "Fiber cement")
dataset$wall_type <- relevel(factor(dataset$wall_type), ref = "Brick")
dataset$heating_type <- relevel(factor(dataset$heating_type), ref = "District heating")

### Missing values

#year_rebuilt
dataset <- dataset %>% mutate(year_rebuilt = if_else(is.na(year_rebuilt) | year_rebuilt < 1970, 1970, year_rebuilt))

#impute NA-values of maximum_coverage_percentage to 30% 
dataset <- dataset %>% mutate(maximum_coverage_percentage = if_else(is.na(maximum_coverage_percentage), 30, maximum_coverage_percentage))

#impute NA-values of subdivision size to 700
dataset <- dataset %>% mutate(subdivision_size = if_else(is.na(subdivision_size), 700, subdivision_size))

#impute NA in distance variables to their cap values 
dataset <- dataset %>% mutate(across(c(lake_distance,
                                       stream_distance,
                                       coast_distance,
                                       road_distance,
                                       highway_distance,
                                       train_station_distance,
                                       railway_distance,
                                       windturbine_distance),
                                     ~ replace_na(., max(., na.rm = TRUE))))


### Auxiliary variables

#create a numerical variable for sales_date
dataset <- dataset %>% mutate(sales_date = as.Date(sales_date), sales_date_numeric = decimal_date(sales_date))

#create an integer variable for the number of possible subdivisions
dataset <- dataset %>% mutate(number_of_divisions = pmax(1, floor(lot_size / subdivision_size)))

#check that there are no remaining NA's in sales data (lot sales have NA for building variables and properties for sales date)
which(is.na(subset(dataset, type == "Property sale")), arr.ind = TRUE)

```

## 2.4 Cross-validation design 
The code below partitions the property sales and lot sales into a standard four-fold cross-validation split ($K=4$). The resulting folds are subsequently used both to estimate out-of-sample predictive performance and to assess the robustness of the fitted model to perturbations of the training data.

```{r Cross-validation, eval=TRUE, echo=TRUE, warning=FALSE, results = 'hide'}
# pick out sales and properties and make factor and ordered factor of type variable for sales
dataset_property_sales <- dataset %>% filter(type == "Property sale")  %>% 
                                      mutate(type = factor(type, levels = c("Lot sale", "Property sale")),
                                             type_ordered = ordered(type, levels = c("Lot sale", "Property sale")))

dataset_lot_sales <- dataset %>% filter(type == "Lot sale") %>% 
                                 mutate(type = factor(type, levels = c("Lot sale", "Property sale")),
                                        type_ordered = ordered(type, levels = c("Lot sale", "Property sale")))

dataset_properties <- dataset %>% filter(type == "Property") %>% mutate(sales_date_numeric = 2024.0) #we want to value properties 1 January 2024


#assign folds
assign_cv_folds <- function(data, N, seed = 123) {
  set.seed(seed)
  data %>% mutate(validation_set = sample(rep(1:N, length.out = n())), .rand = runif(n()))
}

dataset_property_sales_cv <- assign_cv_folds(dataset_property_sales, N = N, seed = 123)
dataset_lot_sales_cv <- assign_cv_folds(dataset_lot_sales, N = N, seed = 123)

table(dataset_property_sales_cv$validation_set, useNA = "ifany")
table(dataset_lot_sales_cv$validation_set, useNA = "ifany")

```


# 3. Paper I: "Generalized Additive Models for Mass Appraisal: Evidence from the Danish Public Property Valuation System" 

## 3.1 Model specification and fitting
The code below fits the GAM model structure described in "Generalized Additive Models for Mass Appraisal: Evidence from the Danish Public Property Valuation System". A model is fitted for each cross-validation fold using the remaining $K-1$ folds as the training data. 

```{r Model training, eval=TRUE, echo=TRUE, warning=FALSE, results='hide'}
fit_fold <- function(i, data, k_spatial, k_other = k_other, fraction = 1.0, checkpoint_prefix = "Model", save_checkpoint = TRUE) {
  
  training_subset <- data %>% filter(validation_set != i)

  if (fraction < 1.0) {
    
    training_subset <- training_subset %>%
                       group_by(municipality_type) %>%
                       arrange(.rand, .by_group = TRUE) %>%
                       slice_head(prop = fraction) %>%
                       ungroup() 
  }
  
  model <- bam(sales_price ~ 
               s(coordinate_east, coordinate_north, k = k_spatial, pc = c(697000, 6150000)) +
               s(sales_date_numeric, k = k_other, by = municipality_type, pc = 2024) +
               s(year_built, k = k_other, by = municipality_type, pc = 1970) +
               s(year_rebuilt, k = k_other, by = municipality_type, pc = 1970) +
               s(living_space, k = k_other, by = municipality_type, pc = 140) +
               s(lot_size, k = k_other, by = municipality_type, pc = 800) +
               s(coast_distance, k = k_other, pc = 1500) +
               s(lake_distance, k = k_other, pc = 500) +
               s(stream_distance, k = k_other, pc = 1000) +
               s(road_distance, k = k_other, pc = 2000) +
               s(highway_distance, k = k_other, pc = 2000) +
               s(train_station_distance, k = k_other, pc = 2000) +
               s(railway_distance, k = k_other, pc = 1000) +
               s(windturbine_distance, k = k_other, pc = 3000) +
               s(ocean_view, k = k_other, pc = 0) +
               s(lake_view, k = k_other, pc = 0) +
               s(forest_size, k = k_other, pc = 0) +
               s(number_of_bathrooms, by = municipality_type, k = 2, pc = 1) +
               house_type * municipality_type +
               roof_type * municipality_type +
               wall_type * municipality_type +
               heating_type * municipality_type,
               data = training_subset,
               chunk.size = nrow(training_subset),
               family = tw(link = log),
               method = "fREML")
  
  if (save_checkpoint) {
    filename <- sprintf("%s_fold%d_k%d_frac%.2f.rds", checkpoint_prefix, i, k_spatial, fraction)
    saveRDS(model, file = file.path("checkpoints", filename))
  }
  
  model
}

fit_model_ensemble <- function(data, N, k_spatial, k_other = 10, fraction = 1.0, checkpoint_prefix = "Model") {
  
  models <- future_lapply(1:N, function(i) {
    fit_fold(
      i = i,
      data = data,
      k_spatial = k_spatial,
      k_other = k_other,
      fraction = fraction,
      checkpoint_prefix = checkpoint_prefix
    )
  })
  names(models) <- paste0("Model", 1:N)
  models
}

if (load) {
  
  models <- readRDS("saved_files/models.rds")
  
} else {
  
  models <- fit_model_ensemble(data = dataset_property_sales_cv,
                                N = N,
                                k_spatial = k_spatial,
                                k_other = k_other,
                                fraction = 1.0,
                                checkpoint_prefix = "main")
  
  saveRDS(models, "saved_files/models.rds")

}

```


## 3.2 Partial effects

#### Spatial smooths 
The code below extracts the spatial smooth for each model and plots it alongside the log-scale standard error. The plotting relies on a premade spatial grid of Denmark, where ocean and land are marked. 

```{r Spatial smooths property models, eval=TRUE, echo=TRUE, warning=FALSE, results = 'hide', fig.width=20, fig.height=30}


# 1a) Predict once per model and store results
prediction_results <- future_lapply(seq_along(models), function(i) {
  model <- models[[i]]
  model_data <- model.frame(model)
  grid <- coordinate_grid

  for (var in setdiff(all.vars(formula(model)), names(grid))) {
    x <- model_data[[var]]
    if (is.factor(x)) {
      grid[[var]] <- factor(rep(levels(x)[1], nrow(grid)), levels = levels(x))
    } else {
      grid[[var]] <- 0
    }
  }

  terms <- predict.gam(model, grid, type = "terms", se.fit = TRUE)
  df_predictions <- as.data.frame(terms$fit)
  df_se <- as.data.frame(terms$se.fit)

  effect <- df_predictions$`s(coordinate_east,coordinate_north)`
  se <- df_se$`s(coordinate_east,coordinate_north)`
  effect[grid$ocean] <- NA
  se[grid$ocean] <- NA

  grid$effect_coordinates <- effect
  grid$se_coordinates <- se

  list(model_name = names(models)[i],
       grid = grid)
})


# 1b) Compute SD across models (per grid cell) ---
effects_mat <- do.call(cbind, lapply(prediction_results, function(x) x$grid$effect_coordinates))
base_grid   <- prediction_results[[1]]$grid

se_across_models <- apply(effects_mat, 1, sd, na.rm = TRUE)
se_across_models[base_grid$ocean] <- NA

# 2. Global ranges (normalized across everything) ---
# Effects: symmetric across *all model* effects
all_effects <- unlist(lapply(prediction_results, function(x) x$grid$effect_coordinates))
max_abs_effect <- max(abs(all_effects), na.rm = TRUE)
effect_limits  <- c(-max_abs_effect, max_abs_effect)

# SEs: include per-model SEs AND the across-models SD
all_ses_models <- unlist(lapply(prediction_results, function(x) x$grid$se_coordinates))
se_limits <- c(0, max(c(all_ses_models, se_across_models), na.rm = TRUE))  # SEs are >= 0

# 3. Plotting function (adds land/ocean outline) ---
plot_model_terms_from_grid <- function(grid, model_name, effect_range, se_range) {

  plot_effect <- ggplot(data = grid) +
                 geom_tile(aes(coordinate_east, coordinate_north, fill = effect_coordinates)) +
                 labs(x = "", y = "") +
                 scale_fill_scico(palette = "vik",
                                  name = paste("Partial effect", sub("^Model", "Model ", model_name)),
                                  na.value = "white",
                                  limits = effect_range) +
                 geom_contour(aes(x = coordinate_east, y = coordinate_north, z = as.numeric(!ocean)),
                              color = "black", 
                              size = 0.4) +
                 coord_fixed() +
                 theme_void() +
                 theme(text = element_text(size = 12),
                       legend.position = c(0.85, 0.75),
                       legend.title = element_text(size = 9, margin = margin(b = 5)),
                       legend.text = element_text(size = 9),
                       legend.title.align = 0.5)
  
  model_label <- sub("^Model", "Model ", model_name)

  plot_se <- ggplot(data = grid) +
             geom_tile(aes(coordinate_east, coordinate_north, fill = se_coordinates)) +
             labs(x = "", y = "") +
             scale_fill_scico(
             palette = "vik",
             name = bquote(sigma[.(model_label)]),
             na.value = "white",
             limits = se_range) +
             geom_contour(aes(x = coordinate_east, y = coordinate_north, z = as.numeric(!ocean)),
                          color = "black", 
                          size = 0.4) +
             coord_fixed() +
             theme_void() +
             theme(text = element_text(size = 12),
                   legend.position = c(0.85, 0.75),
                   legend.title = element_text(size = 9, margin = margin(b = 5)),
                   legend.text = element_text(size = 9),
                   legend.title.align = 0.5)

  list(effect = plot_effect, se = plot_se)
}

# 4. Build plots for each model (effect + SE) ---
plot_lists <- lapply(prediction_results, function(result) {
  plot_model_terms_from_grid(
    result$grid, result$model_name,
    effect_range = effect_limits,
    se_range     = se_limits
  )
})

plot_spatial_smooths <- unlist(plot_lists, recursive = FALSE)

# 5. ONE extra plot: SD of effects across models (uses same se_limits) ---
base_grid2 <- base_grid
base_grid2$se_across_models <- se_across_models  # <- bake it into the data

plot_sd_across <- ggplot(data = base_grid2) +
                  geom_tile(aes(coordinate_east, coordinate_north, fill = se_across_models)) +
                  labs(x = "", y = "") +
                  scale_fill_scico(palette = "vik",
                                   name = expression(sigma[Ensemble]),
                                   na.value = "white",
                                   limits = se_limits) +
                  geom_contour(aes(x = coordinate_east, y = coordinate_north, z = as.numeric(!ocean)),
                               color = "black", 
                               size = 0.4) +
                  coord_fixed() +
                  theme_void() +
                  theme(text = element_text(size = 12),
                        legend.position = c(0.85, 0.75),
                        legend.title = element_text(size = 9, margin = margin(b = 5)),
                        legend.text = element_text(size = 9),
                        legend.title.align = 0.5)

plot_spatial_smooths <- c(plot_spatial_smooths, list(plot_sd_across))
saveRDS(plot_spatial_smooths, "saved_files/plot_spatial_smooths.rds")
wrap_plots(plot_spatial_smooths, ncol = 3)

```


#### Other smooth effects 
The code below extracts and plots the remaining smooth effects of the trained GAM-models. A smooth corresponding to each trained model is plotted, but for readability confidence intervals are only displayed for Model 1. 

```{r Other smooth effects property models, eval=TRUE, echo=TRUE, warning=FALSE, results = 'hide', fig.width=20, fig.height=30}

smooth_variables <- c(
  sales_date_numeric = "Sales date",
  year_built = "Year built",
  year_rebuilt = "Year rebuilt",
  living_space = "Living space [m²]",
  lot_size = "Lot size [m²]",
  coast_distance = "Distance to coast [m]",
  lake_distance = "Distance to lake [m]",
  stream_distance = "Distance to stream [m]",
  road_distance = "Distance to road [m]",
  highway_distance = "Distance to highway [m]",
  train_station_distance = "Distance to train station [m]",
  railway_distance = "Distance to railway [m]",
  windturbine_distance = "Distance to wind turbine [m]",
  ocean_view = "Ocean view [m]",
  lake_view = "Lake view [m]",
  forest_size = "Forest size [ha]",
  number_of_bathrooms = "Number of bathrooms")

plot_smooths <- list()

for (var in names(smooth_variables)) {

  smooth_dfs <- lapply(names(models), function(model_name) {
    mod <- models[[model_name]]
    df <- tryCatch(smooth_estimates(mod, smooth = paste0("s(", var, ")"), partial_match = TRUE), error = function(e) {return(NULL)})
    if (!is.null(df)) {df$Model <- sub("^Model", "Model ", model_name)}
    df
  })

  smooth_dfs <- Filter(Negate(is.null), smooth_dfs)  # remove NULLs
  if (length(smooth_dfs) == 0) next

  combined_df <- bind_rows(smooth_dfs)

  has_by <- any(!is.na(combined_df$.by))
  plot_title <- paste("Partial effect of", gsub("_", " ", var))

  if (has_by) {

    by_var_name <- unique(na.omit(combined_df$.by))[[1]]

    p <- ggplot(combined_df, aes(x = .data[[var]], y = .estimate,
                                 color = .data[[by_var_name]],
                                 linetype = Model, 
                                 group = interaction(Model, .data[[by_var_name]]))) +
         geom_ribbon(data = subset(combined_df, Model == "Model 1"),
                     aes(ymin = .estimate - 1.96*.se, ymax = .estimate + 1.96*.se),
                     alpha = 0.5, 
                     color = NA) +
         geom_line(linewidth = 1) +
         labs(title = plot_title,
              x = smooth_variables[[var]],
              y = "Partial effect") +
         scale_color_brewer(palette = "Dark2", name = "Municipality type") +
         scale_linetype_manual(values = c("solid", 
                                          "dashed", 
                                          "dotted", 
                                          "dotdash", 
                                          "twodash", 
                                          "longdash")) +
         guides(fill = guide_legend(override.aes = list(linetype = 0)))

  } else {

    p <- ggplot(combined_df, aes(x = .data[[var]], y = .estimate, linetype = Model)) +
         geom_ribbon(data = subset(combined_df, Model == "Model 1"),
                     aes(ymin = .estimate - 1.96*.se, ymax = .estimate + 1.96*.se),
                     alpha = 0.5, 
                     color = NA) +
         geom_line(linewidth = 1, color = "black") +
         labs(title = plot_title,
              x = smooth_variables[[var]],
              y = "Partial effect") +
         scale_linetype_manual(values = c("solid", 
                                          "dashed", 
                                          "dotted", 
                                          "dotdash", 
                                          "twodash", 
                                          "longdash"))
  }
  
  p <- p + theme(legend.position = "none")

  plot_smooths[[var]] <- p
}


#manual axis specifications of some ill-behaved smooths to make plots easier to understand
plot_smooths$stream_distance <- plot_smooths$stream_distance + coord_cartesian(ylim = c(-0.03, 0.1))
plot_smooths$lake_view <- plot_smooths$lake_view + coord_cartesian(xlim = c(0.0, 10000), ylim = c(0, 0.6))
plot_smooths$forest_size <- plot_smooths$forest_size + coord_cartesian(xlim = c(0.0,40), ylim = c(-0.02, 0.1))

saveRDS(plot_smooths, "saved_files/plot_smooths.rds")
wrap_plots(plot_smooths, ncol = 2) + 
           plot_layout(guides = "collect") &
           theme(legend.position = "bottom",
                 legend.text = element_text(size = 12),
                 legend.title = element_text(size = 13, face = "bold"),
                 legend.box = "horizontal",
                 legend.justification = "center")
```

#### Factor effects
The code below extracts and plots the factor effects of the trained GAM-models. A point is shown for the effect of each model but error bars are shown only for Model 1. 

```{r Factor effects property models, eval=TRUE, echo=TRUE, warning=FALSE, results = 'hide', fig.width=20, fig.height=20}
factor_variables <- c(
  house_type = "House type",
  municipality_type = "Municipality type",
  roof_type = "Roof type",
  wall_type = "Wall type",
  heating_type = "Heating type")

make_termwise_df <- function(model, model_data, variable, interaction_by = "municipality_type") {

  v_vals <- unique(model_data[[variable]])
  if (is.null(interaction_by)) {
    df <- data.frame(setNames(list(v_vals), variable), stringsAsFactors = FALSE)
  } else {
    m_vals <- unique(model_data[[interaction_by]])
    grid_list <- setNames(list(v_vals, m_vals), c(variable, interaction_by))
    df <- expand.grid(grid_list, stringsAsFactors = FALSE)
    df[[interaction_by]] <- factor(df[[interaction_by]], 
                                   levels = levels(model_data[[interaction_by]]))
  }
  for (var in setdiff(all.vars(formula(model)), names(df))) {
    x <- model_data[[var]]
    if (is.factor(x)) {
      df[[var]] <- factor(rep(levels(x)[1], nrow(df)), levels = levels(x))
    } else {
      df[[var]] <- 0
    }
  }
  return(df)
}

combine_all_term_effects <- function(df, term, interaction_by = "municipality_type") {
  pattern <- paste0("^", term, "$|^", term, ":" ,"|:", term, "$")
  cols_to_sum <- grep(pattern, names(df), value = TRUE)
  df[[paste0("effect_", term)]] <- rowSums(as.matrix(df[cols_to_sum]))
  return(df)
}

combine_all_term_se <- function(df, term, interaction_by = "municipality_type") {
  pattern <- paste0("^", term, "$|^", term, ":" ,"|:", term, "$")
  cols_to_sum <- grep(pattern, names(df), value = TRUE)
  df[[paste0("se_", term)]] <- sqrt(rowSums(as.matrix(df[cols_to_sum])^2))
  return(df)
}

plot_term_effect <- function(df, term, interaction_by = "municipality_type") {
  effect_col <- paste0("effect_", term)
  lower_col <- paste0("lower_", effect_col)
  upper_col <- paste0("upper_", effect_col)
  df_sub <- df %>% filter(term == !!term)
  plot_title <- paste("Partial effect of", gsub("_", " ", term))

  p <- ggplot(df_sub, aes(x = .data[[term]], 
                          y = .data[[effect_col]], 
                          shape = Model,
                          group = interaction(Model, if (identical(term, interaction_by) || 
                                                        !(interaction_by %in% names(df_sub))) NULL 
                                                     else .data[[interaction_by]]))) +
      geom_pointrange(data = subset(df_sub, Model == "Model 1"),
                      aes(ymin = .data[[lower_col]], ymax = .data[[upper_col]]),
                      position = position_dodge(width = 0.5),
                      size = 0.8) +
      geom_point(data = subset(df_sub, Model != "Model 1"),
                 position = position_dodge(width = 0.5),
                 size = 2.5) +
      coord_flip() +
      labs(x = NULL, y = "Partial effect", title = plot_title, shape = "Model") +
      theme(axis.text.y = element_text(angle = 45, hjust = 1))

  if (interaction_by %in% names(df_sub)) {
      color_var <- if (term == interaction_by) term else interaction_by
      p <- p + aes(color = .data[[color_var]]) +
      scale_color_brewer(palette = "Dark2", name = "Municipality type")
  }

  return(p)
}

factor_dfs <- list()

for (model_name in names(models)) {

  model <- models[[model_name]]
  model_data <- model.frame(model)
  relevant_factors <- factor_variables[names(factor_variables) %in% all.vars(formula(model))]

  for (var in names(relevant_factors)) {

    interaction <- if (var == "municipality_type") NULL else "municipality_type"
    df <- make_termwise_df(model, model_data, var, interaction_by = interaction)
    df$term <- var

    terms <- predict.gam(model, df, type = "terms", se.fit = TRUE)
    df_predictions <- as.data.frame(terms$fit)
    df_se <- as.data.frame(terms$se.fit)

    combined_predictions <- combine_all_term_effects(df_predictions, 
                                                     var, 
                                                     interaction_by = interaction) %>% 
                            select(starts_with("effect_"))
    
    combined_se <- combine_all_term_se(df_se, 
                                       var, 
                                       interaction_by = interaction) %>% 
                   select(starts_with("se_"))

    df_final <- df %>% select(all_of(c(var, interaction)), term) %>%
                       bind_cols(combined_predictions) %>%
                       bind_cols(combined_se) %>%
                       mutate(across(starts_with("effect_"),
                                     ~ . - 1.96 * get(str_replace(cur_column(), 
                                     "effect_", 
                                     "se_")),
                                     .names = "lower_{.col}")) %>%
                       mutate(across(starts_with("effect_"),
                                     ~ . + 1.96 * get(str_replace(cur_column(), "effect_", "se_")),
                                     .names = "upper_{.col}"))

    df_final$Model <- sub("^Model", "Model ", model_name)
    factor_dfs[[length(factor_dfs) + 1]] <- df_final
  }
}

df_all_factors <- bind_rows(factor_dfs)
plot_factors <- lapply(names(factor_variables), 
                       function(term) {plot_term_effect(df_all_factors, term)})
names(plot_factors) <- names(factor_variables)

saveRDS(plot_factors, "saved_files/plot_factors.rds")

wrap_plots(plot_factors, ncol = 2) +
           plot_layout(guides = "collect") &
           theme(legend.position = "bottom",
                 legend.text = element_text(size = 12),
                 legend.title = element_text(size = 13, face = "bold"),
                 legend.box = "horizontal",
                 legend.justification = "center")
```


## 3.3 Predictive performance
The code below extracts various residual types and calculates different performance measures for the training and validation set. Note that using N-fold cross-validation means that there are N-1 estimates for each training point, and the training error is simply taken as the average over these. 


```{r Performance property sales, eval=TRUE, echo=TRUE, warning=FALSE}

get_residuals <- function(models, dataset, fold_from_name = TRUE) {
  
  residual_list <- future_lapply(seq_along(models), function(i) {
    
    model_name <- names(models)[i]
    model <- models[[i]]
    
    if (fold_from_name) {
      fold <- as.integer(sub("^Model(\\d+).*$", "\\1", model_name))
    } else {
      fold <- i
    }
    
    predicted_sales_price <- predict.gam(model, newdata = dataset, type = "response")
    
    tibble(sales_price = dataset$sales_price,
           sales_date_numeric = dataset$sales_date_numeric,
           predicted_sales_price = predicted_sales_price,
           residual_sales_price = predicted_sales_price - dataset$sales_price,
           residual_sales_price_percent = (predicted_sales_price - dataset$sales_price) / dataset$sales_price * 100,
           residual_sales_price_log = log(predicted_sales_price) - log(dataset$sales_price),
           ratio = predicted_sales_price / dataset$sales_price,
           validation = dataset$validation_set == fold,
           id = dataset$id,
           municipality_type = dataset$municipality_type,
           Model = model_name)
    
  })
  
  bind_rows(residual_list)
}

residual_df <- get_residuals(models, dataset_property_sales_cv)
saveRDS(residual_df, "saved_files/residual_df.rds")

performance_countrywide <- residual_df %>%
                           mutate(dataset_type = if_else(validation, "Validation", "Training")) %>%
                           group_by(dataset_type) %>%
                           summarise(municipality_type = "Countrywide",
                                     MALE = mean(abs(residual_sales_price_log)),
                                     MAPE = mean(abs(residual_sales_price_percent)),
                                     MAE = mean(abs(residual_sales_price)),
                                     PM20 = mean(abs(residual_sales_price_percent) <= 20) * 100,
                                     COD = mean(abs(ratio - median(ratio))) / median(ratio) * 100,
                                     PRD = mean(ratio) / (sum(predicted_sales_price) / sum(sales_price)),
                                     .groups = "drop")

performance_municipality_type <- residual_df %>%
                                 mutate(dataset_type = if_else(validation, "Validation", "Training")) %>%
                                 group_by(dataset_type, municipality_type) %>%
                                 summarise(MALE = mean(abs(residual_sales_price_log)),
                                           MAPE = mean(abs(residual_sales_price_percent)),
                                           MAE = mean(abs(residual_sales_price)),
                                           PM20 = mean(abs(residual_sales_price_percent) <= 20) * 100,
                                           COD = mean(abs(ratio - median(ratio))) / median(ratio) * 100,
                                           .groups = "drop")

performance <- bind_rows(performance_countrywide, performance_municipality_type)

municipality_levels <- c("Countrywide", sort(setdiff(unique(performance$municipality_type), "Countrywide")))

performance_sorted <- performance %>%
                      mutate(dataset_type = factor(dataset_type, levels = c("Training", "Validation")),
                             municipality_type = factor(municipality_type, levels = municipality_levels),
                             MAE = round(MAE,0),
                             PRD = ifelse(is.na(PRD), "", round(PRD, 2))) %>%
                      arrange(municipality_type, dataset_type)  %>% 
                      rename(Type = dataset_type,
                             `Municipality type` = municipality_type,
                             `MAPE [%]` = MAPE,
                             `MAE [DKK]` = MAE,
                             `PM20 [%]` = PM20,
                             `COD [%]` = COD)

#standard table
#knitr::kable(performance_sorted, digits = 2, caption = "Model Performance by Municipality Type")

#nicer table
kable(performance_sorted %>% 
select(-`Municipality type`), 
       digits = 2, 
       caption = "Model Performance by Municipality Type", 
       booktabs = TRUE) %>%
kable_styling(full_width = FALSE, position = "center") %>%
group_rows("Countrywide", 1, 2) %>%
group_rows("Capital", 3, 4) %>%
group_rows("Commuter", 5, 6) %>%
group_rows("Large city", 7, 8) %>%
group_rows("Provincial", 9, 10) %>%
group_rows("Rural", 11, 12)

tbl_latex <- kable(performance_sorted %>% select(-`Municipality type`),
                   format = "latex", booktabs = TRUE, digits = 2,
                   caption = "Model performance by municipality type.",
                   label = "perf_by_muni") %>%
             kable_styling(full_width = FALSE, 
                           position = "center", 
                           latex_options = c("hold_position")) %>%
             group_rows("Countrywide", 1, 2) %>%
             group_rows("Capital", 3, 4) %>%
             group_rows("Commuter", 5, 6) %>%
             group_rows("Large city", 7, 8) %>%
             group_rows("Provincial", 9, 10) %>%
             group_rows("Rural", 11, 12)

writeLines(tbl_latex, "saved_files/performance_by_muni.tex")

residual_df_validation <- residual_df %>% 
                          filter(validation) %>% 
                          mutate(pm20 = as.integer(abs(residual_sales_price_percent) <= 20))

### Various residual plots on validation set
plot_validation_residuals_log <- ggplot(residual_df_validation,
                                        aes(x = sales_price / 1e6, 
                                            y = residual_sales_price_log)) +
                                 geom_point(alpha = 0.6, color = "#6BAED6", size = 0.25) +
                                 geom_smooth(aes(x = sales_price / 1e6, 
                                                 y = abs(residual_sales_price_log), 
                                                 linetype = "MALE"),
                                             se = FALSE, 
                                             color = "black", 
                                             size = 0.8, 
                                             inherit.aes = FALSE) +
                                scale_linetype_manual(name = NULL,
                                                      values = c(MALE = "solid"),
                                guide = guide_legend(override.aes = list(color = "black", size = 0.8),
                                                     nrow = 1, 
                                                     byrow = TRUE)) +
                                coord_cartesian(ylim = c(-1, 1)) +
                                labs(x = "Sales price [million DKK]", 
                                     y = "Log residuals") +
                                theme(legend.position = c(0.5, 0.4))

saveRDS(plot_validation_residuals_log, "saved_files/plot_validation_residuals_log.rds")
print(plot_validation_residuals_log)



plot_validation_residuals_percent <- ggplot(residual_df_validation, 
                                            aes(x = sales_price / 1e6, 
                                                y = residual_sales_price_percent)) +
                                     geom_point(alpha = 0.6, color = "#6BAED6", size = 0.25) +
  
                                     # Absolute residuals line (MAPE) — inherit.aes = FALSE to decouple from color
                                     geom_smooth(aes(x = sales_price / 1e6, 
                                                     y = abs(residual_sales_price_percent), 
                                                     linetype = "MAPE"),
                                                 se = FALSE, 
                                                 color = "black", 
                                                 size = 0.8, 
                                                 inherit.aes = FALSE) +
  
                                     # pm20 probability line 
                                     geom_smooth(aes(x = sales_price / 1e6, 
                                                     y = pm20 * 100, 
                                                     linetype = "PM20"),
                                                 se = FALSE, 
                                                 color = "black", 
                                                 size = 0.8, 
                                                 inherit.aes = FALSE) +
                                     scale_linetype_manual(name = NULL,
                                                           values = c(MAPE = "solid", PM20 = "dashed"),
                                                           guide = guide_legend(override.aes = list(color = "black", 
                                                                                size = 0.8),
                                                           nrow = 1, byrow = TRUE)) +
                                     ylim(-100, 100) +
                                     labs(x = "Sales price [million DKK]", 
                                          y = "Percentage residuals [%]") +
                                     theme(legend.position = c(0.5, 0.4))


saveRDS(plot_validation_residuals_percent, "saved_files/plot_validation_residuals_percent.rds")
print(plot_validation_residuals_percent)

plot_validation_residuals <- ggplot(residual_df_validation, 
                                    aes(x = sales_price / 1e6, 
                                        y = residual_sales_price / 1e6)) +
                             geom_point(alpha = 0.6, color = "#6BAED6", size = 0.25) +
  
                             # MAE line, decoupled from color mapping
                             geom_smooth(aes(x = sales_price / 1e6, 
                                             y = abs(residual_sales_price) / 1e6, 
                                             linetype = "MAE"),
                                         se = FALSE, 
                                         color = "black", 
                                         size = 0.8, 
                                         inherit.aes = FALSE) +
                      
                             scale_linetype_manual(name = NULL,
                                                   values = c(MAE = "solid"),
                                                   guide = guide_legend(override.aes = list(color = "black", size = 0.8), 
                                                   nrow = 1, 
                                                   byrow = TRUE)) +
                             ylim(-3, 3) +
                             labs(x = "Sales price [million DKK]", 
                                  y = "Residuals [million DKK]") +
                             theme(legend.position = c(0.5, 0.4)) # inside top-right



saveRDS(plot_validation_residuals, "saved_files/plot_validation_residuals.rds")
print(plot_validation_residuals)

### Smoothed error plots for training and validation error vs sales date

residual_df$type <- with(residual_df, ifelse(validation, "Validation error", "Training error"))


plot_log_errors <- ggplot(residual_df) + 
                   geom_smooth(aes(x = sales_date_numeric, 
                                   y = abs(residual_sales_price_log), 
                                   linetype = type),
                                   se = FALSE, 
                                   color = "black", 
                                   size = 0.8, 
                                   inherit.aes = FALSE) +
                   labs(x = "Sales date", 
                        y = "MALE") +
                   theme(legend.title = element_blank(),
                         legend.position = c(0.5, 0.4))


saveRDS(plot_log_errors, "saved_files/plot_log_errors.rds")
print(plot_log_errors)

plot_percentage_errors <- ggplot(residual_df) + 
                          geom_smooth(aes(x = sales_date_numeric, 
                                          y = 100*as.integer(abs(residual_sales_price_percent) <= 20), 
                                          linetype = type),
                                          se = FALSE, 
                                          color = "black", 
                                          size = 0.8, 
                                          inherit.aes = FALSE) +
                          labs(x = "Sales date", 
                               y = "PM20 [%]") +
                          theme(legend.title = element_blank(),
                                legend.position = c(0.5, 0.4))


saveRDS(plot_percentage_errors, "saved_files/plot_percentage_errors.rds")
print(plot_percentage_errors)

plot_errors <- ggplot(residual_df) + 
               geom_smooth(aes(x = sales_date_numeric, 
                               y = abs(residual_sales_price) / 1e6, 
                               linetype = type),
                               se = FALSE, 
                               color = "black", 
                               size = 0.8, 
                               inherit.aes = FALSE) +
               labs(x = "Sales date", 
                    y = "MAE [million DKK]") +
               theme(legend.title = element_blank(),
                     legend.position = c(0.5, 0.4))

saveRDS(plot_errors, "saved_files/plot_errors.rds")
print(plot_errors)

```


## 3.4 Illustrative valuations
The code below performs valuations using the fitted models a set of simplified, fictitious properties, which are described in the main paper. 

```{r Interpretation property models, eval=TRUE, echo=TRUE, warning=FALSE}

make_interpretation_results <- function(model, model_name) {
  
  model_data <- model.frame(model)
  
  # generate fictitious properties
  grid <- data.frame(coordinate_east = c(697000, 462037, 469142, 725599),
                     coordinate_north = c(6150000, 6150225, 6322298, 6185920))
  
  point_constraints <- list()
  for (sm in model$smooth) {
    if (!is.null(sm$point.con)) {point_constraints <- c(point_constraints, sm$point.con)}
  }
  
  vars_in_model <- all.vars(formula(model))
  
  # fill missing variables in grid
  for (var in setdiff(vars_in_model, names(grid))) {
    
    if (is.factor(model_data[[var]])) {
      
      grid[[var]] <- factor(rep(levels(model_data[[var]])[1], nrow(grid)),
                            levels = levels(model_data[[var]]))
      
    } else {
      value <- if (var %in% names(point_constraints)) {
        point_constraints[[var]]
      } else {
        mean(model_data[[var]], na.rm = TRUE)
      }
      grid[[var]] <- rep(value, nrow(grid))
    }
  }
  
  # augment properties
  grid$coast_distance[2:4] <- 100
  grid$year_built[2:4] <- 2020
  grid$municipality_type[2:4] <- factor(c("Provincial", "Rural", "Capital"), levels = levels(model_data$municipality_type))
  grid$type <- c("Reference", "Esbjerg", "Nordjylland", "Strandvejen")
  
  terms <- as.data.frame(predict.gam(model, grid, type = "terms"))
  
  # helper for summing possibly missing columns
  sum_terms <- function(df, cols) {
    cols_present <- intersect(cols, names(df))
    if (length(cols_present) == 0) {return(rep(0, nrow(df)))}
    rowSums(df[, cols_present, drop = FALSE])
  }
  
  adjustments <- terms %>%
                 mutate(standard_price = exp(as.vector(coef(model))[1]),
      
                        adjustment_coordinates = exp(`s(coordinate_east,coordinate_north)`),
      
                        # Smooths with municipality_type interactions – 
                        #might as well sum over all is it only non-zero for the relevant smooth
                         
                        adjustment_sales_date = exp(`s(sales_date_numeric):municipality_typeCapital` +
                                                    `s(sales_date_numeric):municipality_typeRural` +
                                                    `s(sales_date_numeric):municipality_typeCommuter` +
                                                    `s(sales_date_numeric):municipality_typeProvincial` +
                                                    `s(sales_date_numeric):municipality_typeLarge city`),
      
                        adjustment_year_built = exp(`s(year_built):municipality_typeCapital` +
                                                    `s(year_built):municipality_typeRural` +
                                                    `s(year_built):municipality_typeCommuter` +
                                                    `s(year_built):municipality_typeProvincial` +
                                                    `s(year_built):municipality_typeLarge city`),
      
                        adjustment_year_rebuilt = exp(`s(year_rebuilt):municipality_typeCapital` +
                                                      `s(year_rebuilt):municipality_typeRural` +
                                                      `s(year_rebuilt):municipality_typeCommuter` +
                                                      `s(year_rebuilt):municipality_typeProvincial` +
                                                      `s(year_rebuilt):municipality_typeLarge city`),
      
                        adjustment_living_space = exp(`s(living_space):municipality_typeCapital` +
                                                      `s(living_space):municipality_typeRural` +
                                                      `s(living_space):municipality_typeCommuter` +
                                                      `s(living_space):municipality_typeProvincial` +
                                                      `s(living_space):municipality_typeLarge city`),
      
                        adjustment_lot_size = exp(`s(lot_size):municipality_typeCapital` +
                                                  `s(lot_size):municipality_typeRural` +
                                                  `s(lot_size):municipality_typeCommuter` +
                                                  `s(lot_size):municipality_typeProvincial` +
                                                  `s(lot_size):municipality_typeLarge city`),
                                              
                        adjustment_coast_distance = exp(`s(coast_distance)`),
          
                        adjustment_lake_distance = exp(`s(lake_distance)`),
          
                        adjustment_stream_distance = exp(`s(stream_distance)`),
          
                        adjustment_road_distance = exp(`s(road_distance)`),
          
                        adjustment_highway_distance = exp(`s(highway_distance)`),
          
                        adjustment_train_station_distance = exp(`s(train_station_distance)`),
          
                        adjustment_railway_distance = exp(`s(railway_distance)`),
          
                        adjustment_windturbine_distance = exp(`s(windturbine_distance)`),
          
                        adjustment_ocean_view = exp(`s(ocean_view)`),
          
                        adjustment_lake_view = exp(`s(lake_view)`),
          
                        adjustment_forest_size = exp(`s(forest_size)`),
          
                        adjustment_number_of_bathrooms = exp(`s(number_of_bathrooms):municipality_typeCapital` +
                                                             `s(number_of_bathrooms):municipality_typeRural` +
                                                             `s(number_of_bathrooms):municipality_typeCommuter` +
                                                             `s(number_of_bathrooms):municipality_typeProvincial` +
                                                             `s(number_of_bathrooms):municipality_typeLarge city`),
      
                        #Factor terms with interactions
                        adjustment_municipality_type = exp(municipality_type),
                        adjustment_house_type = exp(house_type + `house_type:municipality_type`),
                        adjustment_roof_type = exp(roof_type + `municipality_type:roof_type`),
                        adjustment_wall_type = exp(wall_type + `municipality_type:wall_type`),
                        adjustment_heating_type = exp(heating_type + `municipality_type:heating_type`),
 
                        adjustment_location = adjustment_coordinates *
                                              adjustment_sales_date *
                                              adjustment_lot_size *
                                              adjustment_coast_distance *
                                              adjustment_lake_distance *
                                              adjustment_stream_distance *
                                              adjustment_road_distance *
                                              adjustment_highway_distance *
                                              adjustment_train_station_distance *
                                              adjustment_railway_distance *
                                              adjustment_windturbine_distance *
                                              adjustment_ocean_view *
                                              adjustment_lake_view *
                                              adjustment_forest_size *
                                              adjustment_municipality_type,
      
                        adjustment_property = adjustment_year_built *
                                              adjustment_year_rebuilt *
                                              adjustment_living_space *
                                              adjustment_house_type *
                                              adjustment_roof_type *
                                              adjustment_wall_type *
                                              adjustment_heating_type * 
                                              adjustment_number_of_bathrooms,
      
                        adjustment_total = adjustment_location * adjustment_property,
                        predicted_price = standard_price * adjustment_total) %>%
                        select(standard_price,
                               adjustment_municipality_type,
                               adjustment_coordinates,
                               adjustment_sales_date,
                               adjustment_year_built,
                               adjustment_year_rebuilt,
                               adjustment_living_space,
                               adjustment_lot_size,
                               adjustment_coast_distance,
                               adjustment_lake_distance,
                               adjustment_stream_distance,
                               adjustment_road_distance,
                               adjustment_highway_distance,
                               adjustment_train_station_distance,
                               adjustment_railway_distance,
                               adjustment_windturbine_distance,
                               adjustment_ocean_view,
                               adjustment_lake_view,
                               adjustment_forest_size,
                               adjustment_number_of_bathrooms,
                               adjustment_house_type,
                               adjustment_roof_type,
                               adjustment_wall_type,
                               adjustment_heating_type,
                               adjustment_location,
                               adjustment_property,
                               adjustment_total,
                               predicted_price)
  
  grid_out <- grid %>% select(coast_distance, year_built, municipality_type, type)
  
  bind_cols(grid_out, adjustments) %>% mutate(Model = model_name, .before = 1)
}

results_all_models <- bind_rows(
  lapply(names(models), function(model_name) {
    make_interpretation_results(models[[model_name]], model_name)
  })
)

saveRDS(results_all_models, "saved_files/results_interpretation_property_models.rds")

```



## 3.5 Preditiction uncertainty and stability
The code below flags outliers based on both the GAM-based uncertainty estimate of Model 1 as well as the ensemble-based robustness measure described in the main paper. 

```{r Uncertainty detection, eval=TRUE, echo=TRUE, warning=FALSE, results = 'hide'}

predictions <- imap(models, function(model, name) {
  preds <- predict.gam(model, dataset_properties, type = "link", se.fit = TRUE)
  tibble(!!paste0("pred_", name) := as.vector(preds$fit),
         !!paste0("sigma_", name) := as.vector(preds$se.fit))
})

pred_matrix <- bind_cols(predictions)
property_predictions <- bind_cols(dataset_properties, pred_matrix)
saveRDS(property_predictions, "saved_files/property_predictions.rds")

pred_cols <- grep("^pred_", names(property_predictions), value = TRUE)
sigma_limit <- 0.2

property_predictions <- property_predictions %>%
                        rowwise() %>%
                        mutate(sigma_Ensemble = sd(c_across(all_of(pred_cols))), # ensemble spread
                               sigma_combined = pmax(sigma_Model1, sigma_Ensemble),   # max-rule
                               above_limit    = sigma_combined > sigma_limit) %>%
                        ungroup()

n_total <- nrow(property_predictions)
n_flagged <- sum(property_predictions$above_limit)
pct_flagged <- 100 * n_flagged / n_total


label_expr <- sprintf(
  "atop('Flagged: %s / %s (%.2f%%)', 'Rule: max(' * sigma[\"Model 1\"] * ', ' * sigma[Ensemble] * ') > %.2f')",
  format(n_flagged, big.mark = ","),
  format(n_total,   big.mark = ","),
  pct_flagged,
  sigma_limit
)

plot_uncertainties <- ggplot(property_predictions, aes(x = sigma_Model1, y = sigma_Ensemble)) +
                      geom_point(alpha = 0.1, size = 0.8, color = "#08306B") +
                      geom_point(data = subset(property_predictions, above_limit),
                                 shape = 21, fill = NA, color = "#08306B",
                                 size = 1.5, stroke = 0.7, alpha = 1.0) +
                      geom_density_2d(color = "grey80", linewidth = 0.6, alpha = 0.8) +
                      geom_vline(xintercept = sigma_limit, linetype = "dotted", linewidth = 0.7) +
                      geom_hline(yintercept = sigma_limit, linetype = "dotted", linewidth = 0.7) +
                      annotate("label",
                               x = 0.075, y = 0.24,
                               label = label_expr,
                               parse = TRUE,
                               hjust = 0, vjust = 1,
                               size = 3.5, label.size = 0.2) +
                      coord_equal(xlim = c(0, 0.5), ylim = c(0, 0.5), expand = FALSE) +
                      labs(x = expression(sigma["Model 1"]),
                           y = expression(sigma["Ensemble"]))

saveRDS(plot_uncertainties, "saved_files/plot_uncertainties.rds")
print(plot_uncertainties)

#Plot outliers geographically
outliers <- property_predictions %>% filter(above_limit)

plot_outliers_geographically <- ggplot() +
  
                                # Add land/ocean contour
                                geom_contour(data = coordinate_grid,
                                             aes(x = coordinate_east, 
                                                 y = coordinate_north, 
                                                 z = as.numeric(!ocean)),
                                             color = "black", size = 0.4) +
                                # Outliers
                                geom_point(data = outliers,
                                           aes(x = coordinate_east,
                                               y = coordinate_north,
                                               color = sigma_combined),
                                           size = 1.0, alpha = 0.7) +
                                scale_color_viridis_c(option = "plasma", name = expression(sigma["Combined"])) + 
                                coord_fixed() +
                                theme_void(base_size = 10) +
                                theme(plot.background = element_rect(fill = "white", color = NA),
                                      panel.background = element_rect(fill = "white", color = NA))

saveRDS(plot_outliers_geographically, "saved_files/plot_outliers_geographically.rds")
print(plot_outliers_geographically)


```

## 3.6 Model checking

### Standard Model Checks
The code below plots various standard diagnostics for Model 1 using the appraise() function from the gratia library as described in https://gavinsimpson.github.io/gratia/reference/appraise.html. 

```{r Standard model checks, eval=TRUE, echo=TRUE, warning=FALSE}
k.check(models$Model1)
plot_appraise <- appraise(models$Model1) 
saveRDS(plot_appraise, "saved_files/plot_appraise.rds")

```


### Distributional Assumptions
The subsequent cell generates posterior predictive draws from Model 1 and compares their marginal sales-price distributions with the empirical distribution (see https://gavinsimpson.github.io/gratia/articles/posterior-simulation.html for details on posterior simulation with gratia).

```{r Posterior distribution, eval=TRUE, echo=TRUE, warning=FALSE, results = 'hide'}

n_draws <- 10
samples <-  posterior_samples(models$Model1, n = n_draws, seed = 123, mvn_method = "mgcv") %>%
            select(.draw, .response) %>%
            mutate(draw = paste("Draw", .draw),
                   type = "Posterior draws",
                   sales_price = .response) %>%
            select(draw, type, sales_price)

samples <- samples %>%
           group_by(draw) %>%
           ungroup()

data <- as.data.frame(list(type = "Empirical distribution",
                           draw = "Empirical distribution",
                           sales_price = models$Model1$y))


plot_distribution <- ggplot(data = data,
                            aes(sales_price / 1e6, group = draw, color = type)) +
                     geom_density(linewidth = 1.0, key_glyph = "path") +
                     geom_density(data = samples,
                                  aes(sales_price / 1e6, color = type),
                                      linewidth = 0.25,
                                      key_glyph = "path") +
                     xlim(0, 15) +
                     xlab('Sales price [million DKK]') +
                     ylab('Density') +
                     theme(legend.position = c(0.8, 0.9),
                           legend.text = element_text(size = 12)) +
                     scale_color_manual(name = NULL,
                                        values = c("Empirical distribution" = "black",
                                                   "Posterior draws" = "#F8766D"))

saveRDS(plot_distribution, "saved_files/plot_distribution.rds")
print(plot_distribution)


```

### Concurvity Analysis
The code investigates concurvity between the various smooths for Model 1 based on https://stat.ethz.ch/R-manual/R-devel/library/mgcv/html/concurvity.html. 

```{r Concurvity analysis, eval=TRUE, echo=TRUE, warning=FALSE, results = 'hide', fig.width=20, fig.height=30}

cc <- concurvity(models$Model1, full = FALSE)

collapse_concurvity <- function(conc_mat, pretty_names) {
  
  # remove self-concurvity first
  diag(conc_mat) <- NA
  
  # base smooth names (strip :municipality_typeXXXX)
  base_names <- sub(":municipality_type.*$", "", colnames(conc_mat))
  uniq_terms <- unique(base_names)
  
  # prepare new collapsed matrix
  conc_grouped <- matrix(NA_real_,
                         nrow = length(uniq_terms),
                         ncol = length(uniq_terms),
                         dimnames = list(uniq_terms, uniq_terms))
  
  # fill blocks
  for (i in seq_along(uniq_terms)) {
    for (j in seq_along(uniq_terms)) {
      rows <- base_names == uniq_terms[i]
      cols <- base_names == uniq_terms[j]
      block <- conc_mat[rows, cols, drop = FALSE]
      conc_grouped[i, j] <- max(block)
    }
  }
  
  # drop parametric part
  conc_grouped <- conc_grouped[
    rownames(conc_grouped) != "para",
    colnames(conc_grouped) != "para",
    drop = FALSE
  ]
  
  # apply nice labels
  rownames(conc_grouped) <- pretty_names[rownames(conc_grouped)]
  colnames(conc_grouped) <- pretty_names[colnames(conc_grouped)]
  
  as.data.frame(conc_grouped)
}

# 0) Your pretty names
pretty_names <- c("s(coordinate_east,coordinate_north)" = "Spatial",
                  "s(sales_date_numeric)" = "Sales date",
                  "s(living_space)" = "Living space",
                  "s(lot_size)" = "Lot size",
                  "s(number_of_bathrooms)" = "Number of bathrooms",
                  "s(year_built)" = "Year built",
                  "s(year_rebuilt)" = "Year rebuilt",
                  "s(coast_distance)" = "Coast distance",
                  "s(lake_distance)" = "Lake distance",
                  "s(stream_distance)" = "Stream distance",
                  "s(train_station_distance)" = "Train distance",
                  "s(railway_distance)" = "Railway distance",
                  "s(road_distance)" = "Road distance",
                  "s(highway_distance)" = "Highway distance",
                  "s(windturbine_distance)" = "Wind turbine distance",
                  "s(forest_size)" = "Forest size",
                  "s(ocean_view)" = "Ocean view",
                  "s(lake_view)" = "Lake view")

# 1) collapse concurvity 
conc_grouped_worst <- collapse_concurvity(cc$worst, pretty_names)
conc_grouped_observed <- collapse_concurvity(cc$observed, pretty_names)
conc_grouped_estimate <- collapse_concurvity(cc$estimate, pretty_names)

# 2) pheatmaps WITHOUT legends
concurvity_heatmap_worst <- pheatmap(conc_grouped_worst,
                                     cluster_rows = FALSE,
                                     cluster_cols = FALSE,
                                     na_col       = "black",
                                     fontsize_row = 12,
                                     fontsize_col = 12,
                                     labels_col   = rep("", ncol(conc_grouped_worst)),
                                     color        = viridis(100),
                                     legend       = FALSE,                          
                                     breaks       = seq(0, 1, length.out = 100),
                                     border_color = NA,
                                     silent       = TRUE)

concurvity_heatmap_observed <- pheatmap(conc_grouped_observed,
                                        cluster_rows = FALSE,
                                        cluster_cols = FALSE,
                                        na_col       = "black",
                                        fontsize_row = 12,
                                        fontsize_col = 12,
                                        labels_col   = rep("", ncol(conc_grouped_observed)),
                                        color        = viridis(100),
                                        legend       = FALSE,                         
                                        breaks       = seq(0, 1, length.out = 100),
                                        border_color = NA,
                                        silent       = TRUE)

concurvity_heatmap_estimate <- pheatmap(conc_grouped_estimate,
                                        cluster_rows = FALSE,
                                        cluster_cols = FALSE,
                                        na_col       = "black",
                                        fontsize_row = 12,
                                        fontsize_col = 12,
                                        color        = viridis(100),
                                        legend       = FALSE,                           
                                        breaks       = seq(0, 1, length.out = 100),
                                        border_color = NA,
                                        silent       = TRUE)

#one with legend true to be saved separately
concurvity_heatmap_est <- pheatmap(conc_grouped_estimate,
                                   cluster_rows = FALSE,
                                   cluster_cols = FALSE,
                                   na_col       = "black",
                                   fontsize_row = 12,
                                   fontsize_col = 12,
                                   color        = viridis(100),
                                   legend       = FALSE,                           
                                   breaks       = seq(0, 1, length.out = 100),
                                   border_color = NA,
                                   silent       = TRUE)

# 3) stack the three heatmaps

concurvity_heatmaps <- plot_grid(concurvity_heatmap_worst$gtable,
                                 concurvity_heatmap_observed$gtable,
                                 concurvity_heatmap_estimate$gtable,
                                 ncol = 1,
                                 align = "v",
                                 rel_heights = c(1, 1, 1.2))


legend_df <- data.frame(x = 1, y = 1, z = 0.5)  

legend_plot <- ggplot(legend_df, aes(x = x, y = y, fill = z)) +
               geom_tile() +
               scale_fill_viridis_c(limits = c(0, 1),
                                    breaks = c(0, 0.5, 1),
                                    labels = c("0", "0.5", "1"),
                                    name   = "Concurvity") +
               theme_void() +
               theme(legend.position   = "bottom",
                     legend.title      = element_text(size = 11),
                     legend.text       = element_text(size = 10),
                     legend.key.width  = unit(4.0, "cm"),
                     legend.key.height = unit(0.6, "cm"),
                     legend.margin = margin(t = -20, b = -20))

g <- ggplotGrob(legend_plot)
guide_ix  <- which(sapply(g$grobs, function(x) x$name) == "guide-box")
legend_grob <- g$grobs[[guide_ix]]

concurvity_heatmaps <- plot_grid(concurvity_heatmaps,
                                 legend_grob,
                                 ncol = 1,
                                 rel_heights = c(1, 0.04))

saveRDS(concurvity_heatmap_est, "saved_files/concurvity_heatmap_est.rds")
saveRDS(concurvity_heatmaps, "saved_files/concurvity_heatmaps.rds")
print(concurvity_heatmaps)


```


### Convergence of the Spatial Smooth
The code below fits the model ensemble for different basis set sizes of the spatial smooth to examine the convergence of the validation error.

```{r Convergence basis set, eval=TRUE, echo=TRUE, warning=FALSE, results='hide'}
k_values <- c(100, 200, 500, 1000, 1500, 2000, 3000, 4000, 5000)
k_values <- k_values[k_values <= k_spatial]
  
model_grid_basis <- expand.grid(fold = 1:N, k_spatial = k_values)


if (load) {
  
  models_basis <- readRDS("saved_files/models_basis.rds")
  
} else {
  
  models_basis <- future_lapply(seq_len(nrow(model_grid_basis)), function(idx) {
    row <- model_grid_basis[idx, ]
    fit_fold(
      i = row$fold,
      data = dataset_property_sales_cv,
      k_spatial = row$k_spatial,
      k_other = k_other,
      fraction = 1.0,
      checkpoint_prefix = paste0("basis_k", row$k_spatial)
    )
  })
  
  names(models_basis) <- paste0("Model", model_grid_basis$fold, "_", model_grid_basis$k_spatial)
  saveRDS(models_basis, "saved_files/models_basis.rds")
}

get_residuals_grid <- function(models, dataset, grid_df, grid_name) {
  
  residual_list <- future_lapply(seq_along(models), function(i) {
    model_name <- names(models)[i]
    model <- models[[i]]
    row <- grid_df[i, ]
    fold <- row$fold
    
    predicted_sales_price <- predict.gam(model, newdata = dataset, type = "response")
    
    tibble(sales_price = dataset$sales_price,
           predicted_sales_price = predicted_sales_price,
           residual_sales_price = predicted_sales_price - dataset$sales_price,
           residual_sales_price_percent = (predicted_sales_price - dataset$sales_price) / dataset$sales_price * 100,
           residual_sales_price_log = log(predicted_sales_price) - log(dataset$sales_price),
           validation = dataset$validation_set == fold,
           municipality_type = dataset$municipality_type,
           Model = model_name,
           value = row[[grid_name]])
  })
  
  bind_rows(residual_list)
}

residual_df_basis <- get_residuals_grid(models = models_basis,
                                        dataset = dataset_property_sales_cv,
                                        grid_df = model_grid_basis,
                                        grid_name = "k_spatial")

saveRDS(residual_df_basis, "saved_files/residual_df_basis.rds")

performance_basis <- residual_df_basis %>%
                     mutate(type = if_else(validation, "Validation error", "Training error")) %>%
                     group_by(type, municipality_type, value) %>%
                     summarise(MALE = mean(abs(residual_sales_price_log)), .groups = "drop") %>%
                     rename(k = value)

plot_basis <- ggplot(performance_basis,
                     aes(x = k, y = MALE, color = municipality_type, lty = type)) +
              geom_line() +
              geom_point() +
              labs(x = "Number of spatial basis functions (k)",
                   y = "MALE",
                   color = "Municipality type")

saveRDS(plot_basis, "saved_files/plot_basis.rds")
print(plot_basis)

```

# 4. Paper II: "Property and Land Value Predictions with Generalized Additive Models" 

### 4.1 Standard-house construction
In the code below, lot sales and properties are augmented with a fictitious single-family house corresponding to a set of standard building characteristics. Their standardized property value is subsequently calculated. 

```{r Standardization, eval=TRUE, echo=TRUE, warning=FALSE, results= 'hide'}

standardize_property <- function(data, reference_data) {
  
    data %>% mutate(year_built = 1970,
                    year_rebuilt = 1970,
                    living_space = 140,
                    lot_size = 800,
                    maximum_coverage_percentage = 30,
                    number_of_divisions = 1,
                    number_of_bathrooms = 1,
                    house_type = factor("Detached", levels = levels(factor(reference_data$house_type))),
                    roof_type = factor("Fiber cement", levels = levels(factor(reference_data$roof_type))),
                    wall_type = factor("Brick", levels = levels(factor(reference_data$wall_type))),
                    heating_type = factor("District heating", levels = levels(factor(reference_data$heating_type))))
  }


if (load) {

  dataset_property_sales_cv <- readRDS("saved_files/dataset_property_sales_cv.rds")
  dataset_lot_sales_cv_standardized <- readRDS("saved_files/dataset_lot_sales_cv_standardized.rds")
  dataset_properties <- readRDS("saved_files/dataset_properties.rds")
  dataset_properties_standardized <- readRDS("saved_files/dataset_properties_standardized.rds")

} else {
  
  #standardize 
  dataset_lot_sales_cv_standardized <- standardize_property(dataset_lot_sales_cv, reference_data = dataset)
  dataset_properties_standardized <- standardize_property(dataset_properties, reference_data = dataset)
  
  #calculate standarized prices for each model 
  for (model_name in names(models)) {
    
    dataset_lot_sales_cv_standardized[[paste0("standard_property_value_", model_name)]] <- as.vector(predict.gam(models[[model_name]], dataset_lot_sales_cv_standardized, type = "response"))
    dataset_properties_standardized[[paste0("standard_property_value_", model_name)]] <- as.vector(predict.gam(models[[model_name]], dataset_properties_standardized, type = "response"))
    dataset_properties[[paste0("standard_property_value_", model_name)]] <- as.vector(predict.gam(models[[model_name]], dataset_properties_standardized, type = "response"))
    dataset_properties[[paste0("property_value_", model_name)]] <- as.vector(predict.gam(models[[model_name]], dataset_properties, type = "response"))
  
  }
  
  #restore original lot variables but leave standard building characteristics on the lot
  dataset_lot_sales_cv_standardized$lot_size <- dataset_lot_sales_cv$lot_size
  dataset_lot_sales_cv_standardized$maximum_coverage_percentage <- dataset_lot_sales_cv$maximum_coverage_percentage
  dataset_lot_sales_cv_standardized$number_of_divisions <- dataset_lot_sales_cv$number_of_divisions

  #save
  saveRDS(dataset_lot_sales_cv_standardized, "saved_files/dataset_lot_sales_cv_standardized.rds")
  saveRDS(dataset_property_sales_cv, "saved_files/dataset_property_sales_cv.rds")
  saveRDS(dataset_properties, "saved_files/dataset_properties.rds")
  saveRDS(dataset_properties_standardized, "saved_files/dataset_properties_standardized.rds")

}

```


### 4.2 Model specifications and fitting
The code below fits three alternative GAM approaches for estimating lot values. The direct lot model is trained exclusively on lot sales using location, time, lot characteristics, and surrounding amenities. The standard-property model relates lot-sale prices to the predicted value of a standardized property, while additionally controlling for selected lot and spatial characteristics. The combined model is fitted jointly to property sales and lot sales and distinguishes the two sale types through a type effect and a type-specific spatial adjustment. For each approach, one model is fitted per cross-validation fold using the remaining $K-1$ folds as the training data.

```{r Lot-value models, eval=TRUE, echo=TRUE, warning=FALSE, results= 'hide'}

if (load) {
  
  models_direct <- readRDS("saved_files/models_direct.rds")
  models_standard_property <- readRDS("saved_files/models_standard_property.rds")
  models_combined <- readRDS("saved_files/models_combined.rds")
  
} else {
  
  # -------------------------
  # 1. Direct lot model
  # -------------------------
  
  fit_direct_fold <- function(i,
                              data,
                              k_spatial,
                              k_other = k_other,
                              checkpoint_prefix = "direct_lot",
                              save_checkpoint = TRUE) {
    
    training_subset <- data %>% filter(validation_set != i)
    
    model <- bam(sales_price ~ s(coordinate_east, coordinate_north, k = k_spatial, pc = c(697000, 6150000)) +
                               s(sales_date_numeric, k = k_other, by = municipality_type, pc = 2024) +
                               s(lot_size, k = k_other, by = municipality_type, pc = 800) +
                               s(coast_distance, k = k_other, pc = 1500) +
                               s(lake_distance, k = k_other, pc = 500) +
                               s(stream_distance, k = k_other, pc = 1000) +
                               s(road_distance, k = k_other, pc = 2000) +
                               s(highway_distance, k = k_other, pc = 2000) +
                               s(train_station_distance, k = k_other, pc = 2000) +
                               s(railway_distance, k = k_other, pc = 1000) +
                               s(windturbine_distance, k = k_other, pc = 3000) +
                               s(ocean_view, k = k_other, pc = 0) +
                               s(lake_view, k = k_other, pc = 0) +
                               s(forest_size, k = k_other, pc = 0),
                data = training_subset,
                chunk.size = nrow(training_subset),
                family = tw(link = log),
                method = "fREML")
    
    if (save_checkpoint) {
      
      filename <- sprintf("%s_fold%d_kspatial%d_kother%d.rds",
                          checkpoint_prefix,
                          i,
                          k_spatial,
                          k_other)
      
      saveRDS(model, file = file.path("checkpoints", filename))
      
    }
    
    model
  }
  
  # -------------------------
  # 2. Standard-property model
  # -------------------------
  
  fit_standard_property_fold <- function(i, 
                                         data, 
                                         k_spatial, 
                                         k_standard_property = 10, 
                                         checkpoint_prefix = "standard_property", 
                                         save_checkpoint = TRUE) {
    
    training_subset <- data %>% filter(validation_set != i)
    
    standard_property_var <- paste0("standard_property_value_Model", i)
    
    formula_standard_property <- as.formula(paste0("sales_price ~ ",
                                                   "s(", standard_property_var, ", k = ", k_standard_property, ") + ",
                                                   "s(lot_size, k = 6, pc = 800) + ",
                                                   "s(maximum_coverage_percentage, k = 3, pc = 30) + ",
                                                   "s(number_of_divisions, k = 3, pc = 1) + ",
                                                   "s(coordinate_east, coordinate_north, ",
                                                   "k = ", k_spatial,
                                                   ", pc = c(697000, 6150000))"))
    
    model <- bam(formula_standard_property,
                 data = training_subset,
                 chunk.size = nrow(training_subset),
                 family = tw(link = log),
                 method = "fREML")
    
    if (save_checkpoint) {
      
      filename <- sprintf("%s_fold%d_kspatial%d_kstandard%d.rds",
                          checkpoint_prefix,
                          i,
                          k_spatial,
                          k_standard_property)
      
      saveRDS(model, file = file.path("checkpoints", filename))
    }
    
    model
  }
  
  # -------------------------
  # 3. Combined model
  # -------------------------
  
  fit_combined_fold <- function(i,
                                lot_data,
                                property_data,
                                k_spatial,
                                k_other = k_other,
                                checkpoint_prefix = "combined",
                                save_checkpoint = TRUE) {
    
    lot_training_subset <- lot_data %>% filter(validation_set != i)
    
    property_training_subset <- property_data %>% filter(validation_set != i)
    
    training_subset <- bind_rows(lot_training_subset, property_training_subset)
    
    model <- bam(sales_price ~ s(coordinate_east, coordinate_north, k = k_spatial, pc = c(697000, 6150000)) +
                               s(coordinate_east, coordinate_north, k = k_spatial, by = type_ordered, pc = c(697000, 6150000)) +
                               s(sales_date_numeric, k = k_other, by = municipality_type, pc = 2024) +
                               s(year_built, k = k_other, by = municipality_type, pc = 1970) +
                               s(year_rebuilt, k = k_other, by = municipality_type, pc = 1970) +
                               s(living_space, k = k_other, by = municipality_type, pc = 140) +
                               s(lot_size, k = k_other, by = municipality_type, pc = 800) +
                               s(coast_distance, k = k_other, pc = 1500) +
                               s(lake_distance, k = k_other, pc = 500) +
                               s(stream_distance, k = k_other, pc = 1000) +
                               s(road_distance, k = k_other, pc = 2000) +
                               s(highway_distance, k = k_other, pc = 2000) +
                               s(train_station_distance, k = k_other, pc = 2000) +
                               s(railway_distance, k = k_other, pc = 1000) +
                               s(windturbine_distance, k = k_other, pc = 3000) +
                               s(ocean_view, k = k_other, pc = 0) +
                               s(lake_view, k = k_other, pc = 0) +
                               s(forest_size, k = k_other, pc = 0) +
                               s(number_of_bathrooms, k = 2, by = municipality_type, pc = 1) +
                               house_type * municipality_type +
                               roof_type * municipality_type +
                               wall_type * municipality_type +
                               heating_type * municipality_type +
                               type,
                 data = training_subset,
                 chunk.size = nrow(training_subset),
                 family = tw(link = log),
                 method = "fREML")
    
    if (save_checkpoint) {
      
      filename <- sprintf("%s_fold%d_kspatial%d_kother%d.rds",
                          checkpoint_prefix,
                          i,
                          k_spatial,
                          k_other)
      
      saveRDS(model, file = file.path("checkpoints", filename))
    }
    
    model
  }
  
  # -------------------------
  # 4. Shared ensemble helper
  # -------------------------
  
  fit_model_ensemble <- function(fit_fun,
                                 N,
                                 checkpoint_prefix,
                                 ...) {
    
    models <- future_lapply(seq_len(N), function(i) {
      
      message("Fitting ", checkpoint_prefix, ", fold ", i, " of ", N, " ...")
      
      fit_fun(i = i, checkpoint_prefix = checkpoint_prefix, ...)
      
    })
    
    names(models) <- paste0("Model", seq_len(N))
    
    models
  }
  
  # -------------------------
  # 5. Fit model ensembles
  # -------------------------
  
  models_direct <- fit_model_ensemble(fit_fun = fit_direct_fold,
                                      N = N,
                                      checkpoint_prefix = "direct_lot",
                                      data = dataset_lot_sales_cv_standardized,
                                      k_spatial = k_spatial,
                                      k_other = k_other)
  
  models_standard_property <- fit_model_ensemble(fit_fun = fit_standard_property_fold,
                                                 N = N,
                                                 checkpoint_prefix = "standard_property",
                                                 data = dataset_lot_sales_cv_standardized,
                                                 k_spatial = k_spatial,
                                                 k_standard_property = 10)
  
  models_combined <- fit_model_ensemble(fit_fun = fit_combined_fold,
                                        N = N,
                                        checkpoint_prefix = "combined",
                                        lot_data = dataset_lot_sales_cv_standardized,
                                        property_data = dataset_property_sales_cv,
                                        k_spatial = k_spatial,
                                        k_other = k_other)
  
  # -------------------------
  # 6. Save complete ensembles
  # -------------------------
  
  saveRDS(models_direct, "saved_files/models_direct.rds")
  
  saveRDS(models_standard_property, "saved_files/models_standard_property.rds")
  
  saveRDS(models_combined, "saved_files/models_combined.rds")
  
}

```

## 4.3 Partial effects

#### Spatial smooths 
The code below extracts and visualizes the spatial smooths from each of the three lot-value model ensembles. For every fold-specific model, the estimated spatial effect and its standard error are evaluated on a common coordinate grid covering Denmark. For the combined model, both the shared spatial surface and the sale-type-specific spatial adjustment are extracted.

```{r Spatial smooths lot models, eval=TRUE, echo=TRUE, warning=FALSE, results = 'hide', fig.width=20, fig.height=30}

# -------------------------------
# Helper 1: complete prediction grid
# -------------------------------
complete_grid_for_model <- function(model, grid) {
  
  model_data <- model.frame(model)

  for (var in setdiff(all.vars(formula(model)), names(grid))) {
    
    x <- model_data[[var]]

    if (is.factor(x)) {
      grid[[var]] <- factor(rep(levels(x)[1], nrow(grid)), levels = levels(x))
    } else if (is.ordered(x)) {
      grid[[var]] <- ordered(rep(levels(x)[1], nrow(grid)), levels = levels(x))
    } else {
      grid[[var]] <- 0
    }
  }

  grid
}


# -------------------------------
# Helper 2: extract spatial terms
# -------------------------------
extract_spatial_terms <- function(model, grid, type_level = NULL) {

  grid <- complete_grid_for_model(model, grid)

  if (!is.null(type_level) && "type_ordered" %in% names(grid)) {
    grid$type_ordered <- ordered(
      rep(type_level, nrow(grid)),
      levels = levels(model.frame(model)$type_ordered)
    )
  }

  predictions <- predict.gam(model, grid, type = "terms", se.fit = TRUE)

  fit <- as.data.frame(predictions$fit)
  se  <- as.data.frame(predictions$se.fit)

  term_names <- names(fit)

  global_term <- term_names[term_names == "s(coordinate_east,coordinate_north)"]

  by_terms <- grep("^s\\(coordinate_east,coordinate_north\\):type_ordered", term_names, value = TRUE)

  out <- list()

  if (length(global_term) == 1) {
    out$global <- list(effect = fit[[global_term]],
                       se = se[[global_term]],
                       term = global_term)
  }

  if (!is.null(type_level) && length(by_terms) > 0) {
    matched <- by_terms[grepl(paste0(type_level, "$"), by_terms)]

    if (length(matched) == 1) {
      out$type_specific <- list(effect = fit[[matched]],
                                se = se[[matched]],
                                term   = matched)
    } else {
      out$type_specific <- list(effect = rep(0, nrow(grid)),
                                se = rep(0, nrow(grid)),
                                term   = paste0("reference type: ", type_level))
    }
  }

  out$grid <- grid
  out
}


# -------------------------------
# Helper 3: predict spatial grids
# -------------------------------
make_spatial_prediction_results <- function(
    models,
    coordinate_grid,
    include_type_specific = FALSE
) {

  future_lapply(seq_along(models), function(i) {

    model <- models[[i]]
    model_name <- names(models)[i]

    if (include_type_specific) {
      type_levels <- levels(model.frame(model)$type_ordered)
    } else {
      type_levels <- NULL
    }

    global_res <- extract_spatial_terms(model, coordinate_grid)

    results <- list()

    if (!is.null(global_res$global)) {
      
      grid <- global_res$grid
      grid$effect_coordinates <- global_res$global$effect
      grid$se_coordinates     <- global_res$global$se
      grid$effect_coordinates[grid$ocean] <- NA
      grid$se_coordinates[grid$ocean]     <- NA

      results[[paste0(model_name, "_global")]] <- list(model_name = model_name,
                                                       smooth_name = "Global spatial smooth",
                                                       grid = grid)
    }

    if (include_type_specific) {
      for (tl in type_levels) {
        type_res <- extract_spatial_terms(model,
                                          coordinate_grid,
                                          type_level = tl)

        grid <- type_res$grid
        grid$effect_coordinates <- type_res$type_specific$effect
        grid$se_coordinates     <- type_res$type_specific$se
        grid$effect_coordinates[grid$ocean] <- NA
        grid$se_coordinates[grid$ocean]     <- NA

        results[[paste0(model_name, "type", tl)]] <- list(model_name = model_name,
                                                          smooth_name = paste("Type-specific spatial smooth:", tl),
                                                          type_level = tl,
                                                          grid = grid)
      }
    }

    results
  }) |>
    unlist(recursive = FALSE)
}


# -------------------------------
# Helper 4: plotting
# -------------------------------
plot_model_terms_from_grid <- function(
    grid,
    model_name,
    smooth_name,
    effect_range,
    se_range
) {

  plot_effect <- ggplot(grid) +
                 geom_tile(aes(coordinate_east, coordinate_north,
                           fill = effect_coordinates)) +
                 scale_fill_scico(palette = "vik",
                                  name = paste("Partial effect\n", model_name),
                                  na.value = "white",
                                  limits = effect_range) +
                 geom_contour(aes(coordinate_east, coordinate_north, z = as.numeric(!ocean)),
                              color = "black",
                              size = 0.4) +
                 coord_fixed() +
                 labs(title = smooth_name, x = "", y = "") +
                 theme_void() +
                 theme(plot.title = element_text(size = 10, hjust = 0.5),
                       legend.position = c(0.85, 0.75),
                       legend.title = element_text(size = 9, margin = margin(b = 5)),
                       legend.text = element_text(size = 9),
                       legend.title.align = 0.5)

  plot_se <- ggplot(grid) +
             geom_tile(aes(coordinate_east, coordinate_north,
                       fill = se_coordinates)) +
             scale_fill_scico(palette = "vik",
                              name = bquote(sigma[.(model_name)]),
                              na.value = "white",
                              limits = se_range) +
             geom_contour(aes(coordinate_east, coordinate_north, z = as.numeric(!ocean)),
                          color = "black",
                          size = 0.4) +
             coord_fixed() +
             labs(title = paste("SE:", smooth_name), x = "", y = "") +
             theme_void() +
             theme(plot.title = element_text(size = 10, hjust = 0.5),
                   legend.position = c(0.85, 0.75),
                   legend.title = element_text(size = 9, margin = margin(b = 5)),
                   legend.text = element_text(size = 9),
                   legend.title.align = 0.5)

  list(effect = plot_effect, se = plot_se)
}


# -------------------------------
# Helper 5: full workflow
# -------------------------------
build_spatial_smooth_plots <- function(models,
                                       coordinate_grid,
                                       include_type_specific = FALSE,
                                       save_path = NULL) {

  prediction_results <- make_spatial_prediction_results(
                        models = models,
                        coordinate_grid = coordinate_grid,
                        include_type_specific = include_type_specific)

  all_effects <- unlist(lapply(prediction_results, \(x) x$grid$effect_coordinates))

  max_abs_effect <- max(abs(all_effects), na.rm = TRUE)
  effect_limits <- c(-max_abs_effect, max_abs_effect)

  all_ses <- unlist(lapply(prediction_results, \(x) x$grid$se_coordinates))

  effects_mat <- do.call(cbind, lapply(prediction_results, \(x) x$grid$effect_coordinates))

  base_grid <- prediction_results[[1]]$grid
  base_grid$se_across_models <- apply(effects_mat, 1, sd, na.rm = TRUE)
  base_grid$se_across_models[base_grid$ocean] <- NA

  se_limits <- c(0, max(c(all_ses, base_grid$se_across_models), na.rm = TRUE))

  plot_lists <- lapply(prediction_results, function(res) {
    
    plot_model_terms_from_grid(grid = res$grid,
                               model_name = sub("^Model", "Model ", res$model_name),
                               smooth_name = res$smooth_name,
                               effect_range = effect_limits,
                               se_range = se_limits)
  })

  plots <- unlist(plot_lists, recursive = FALSE)

  plot_sd_across <- ggplot(base_grid) +
                    geom_tile(aes(coordinate_east, coordinate_north, fill = se_across_models)) +
                    scale_fill_scico(palette = "vik",
                                     name = expression(sigma[Ensemble]),
                                     na.value = "white",
                                     limits = se_limits) +
                    geom_contour(aes(coordinate_east, coordinate_north, z = as.numeric(!ocean)),
                                 color = "black",
                                 size = 0.4) +
                    coord_fixed() +
                    labs(title = "SD across spatial smooths", x = "", y = "") +
                    theme_void() +
                    theme(plot.title = element_text(size = 10, hjust = 0.5),
                          legend.position = c(0.85, 0.75),
                          legend.title = element_text(size = 9, margin = margin(b = 5)),
                          legend.text = element_text(size = 9),
                          legend.title.align = 0.5)

  plots <- c(plots, list(sd_across_models = plot_sd_across))

  if (!is.null(save_path)) {
    saveRDS(plots, save_path)
  }

  plots
}

plots_spatial_direct <- build_spatial_smooth_plots(models_direct, 
                                                   coordinate_grid,
                                                   include_type_specific = FALSE,
                                                   save_path = "saved_files/plots_spatial_direct.rds")

plots_spatial_standard_property <- build_spatial_smooth_plots(models_standard_property, 
                                                              coordinate_grid,
                                                              include_type_specific = FALSE,
                                                              save_path = "saved_files/plots_spatial_standard_property.rds")

plots_spatial_combined <- build_spatial_smooth_plots(models_combined, 
                                                     coordinate_grid,
                                                     include_type_specific = TRUE,
                                                     save_path = "saved_files/plots_spatial_combined.rds")


#plots
wrap_plots(plots_spatial_direct, ncol = 3)
wrap_plots(plots_spatial_standard_property, ncol = 3)
#no need to plot the type specific spatial smooths for lot sales as they are 0
plots_spatial_combined_filtered <- plots_spatial_combined[!grepl("typeLot sale", names(plots_spatial_combined))]
wrap_plots(plots_spatial_combined_filtered, ncol = 3)


```

#### Other smooth effects 
The code below extracts and plots the remaining smooth effects for each type of lot model. A smooth corresponding to each trained model is plotted, but for readability confidence intervals are only displayed for Model 1. 

```{r Other smooth effects lot models, eval=TRUE, echo=TRUE, warning=FALSE, results = 'hide'}

# -------------------------
# 1. Smooth variables
# -------------------------

smooth_variables <- c(standard_property_value = "Standard property value",
                      sales_date_numeric = "Sales date",
                      year_built = "Year built",
                      year_rebuilt = "Year rebuilt",
                      living_space = "Living space [m²]",
                      lot_size = "Lot size [m²]",
                      maximum_coverage_percentage = "Maximum coverage percentage",
                      number_of_divisions = "Number of divisions",
                      coast_distance = "Distance to coast [m]",
                      lake_distance = "Distance to lake [m]",
                      stream_distance = "Distance to stream [m]",
                      road_distance = "Distance to road [m]",
                      highway_distance = "Distance to highway [m]",
                      train_station_distance = "Distance to train station [m]",
                      railway_distance = "Distance to railway [m]",
                      windturbine_distance = "Distance to wind turbine [m]",
                      ocean_view = "Ocean view [m]",
                      lake_view = "Lake view [m]",
                      forest_size = "Forest size [ha]",
                      number_of_bathrooms = "Number of bathrooms")

# -------------------------
# 2. Helper: standard-property smooth
# -------------------------

make_standard_property_smooth_plot <- function(models, model_set_name) {
  
  smooth_dfs <- imap(models, function(mod, model_name) {
    
    fold_number <- str_extract(model_name, "\\d+")
    var_name <- paste0("standard_property_value_Model", fold_number)
    
    df <- tryCatch(smooth_estimates(mod, smooth = paste0("s(", var_name, ")"), partial_match = TRUE),
                   error = function(e) {return(NULL)}
    )
    
    if (is.null(df)) return(NULL)
    
    df %>% rename(standard_property_value = all_of(var_name)) %>%
           mutate(Model = sub("^Model", "Model ", model_name))
  })
  
  smooth_dfs <- Filter(Negate(is.null), smooth_dfs)
  
  if (length(smooth_dfs) == 0) { return(NULL) }
   
  combined_df <- bind_rows(smooth_dfs)
  
  ggplot(combined_df,
  aes(x = standard_property_value,
      y = .estimate,
      linetype = Model)) +
  geom_ribbon(data = subset(combined_df, Model == "Model 1"),
              aes(ymin = .estimate - 1.96 * .se,
                  ymax = .estimate + 1.96 * .se),
              alpha = 0.15,
              color = NA) +
  geom_line(linewidth = 1, color = "black") +
  labs(title = paste0(model_set_name, ": Standard property value"),
       x = "Standard property value",
       y = "Partial effect",
       linetype = NULL) +
  theme_minimal()
  
}

# -------------------------
# 3. Helper: generic smooth plot
# -------------------------

make_single_smooth_plot <- function(models, model_set_name, var, label) {
  
  smooth_dfs <- lapply(names(models), function(model_name) {
    
    mod <- models[[model_name]]
    
    df <- tryCatch(smooth_estimates(mod, smooth = paste0("s(", var, ")"), partial_match = TRUE),
                   error = function(e) {return(NULL)}
    )
    
    if (!is.null(df)) {df$Model <- sub("^Model", "Model ", model_name)}
    
    df
    
  })
  
  smooth_dfs <- Filter(Negate(is.null), smooth_dfs)
  
  if (length(smooth_dfs) == 0) {return(NULL)}
  
  combined_df <- bind_rows(smooth_dfs)
  
  has_by <- ".by" %in% names(combined_df) && any(!is.na(combined_df$.by))
  
  plot_title <- paste0(model_set_name, ": ", label)
  
  if (has_by) {
    
    by_var_name <- unique(na.omit(combined_df$.by))[[1]]
    
    p <- ggplot(combined_df,
                aes(x = .data[[var]],
                    y = .estimate,
                    color = .data[[by_var_name]],
                    linetype = Model,
                    group = interaction(Model, .data[[by_var_name]]))) +
        geom_ribbon(data = subset(combined_df, Model == "Model 1"),
        aes(ymin = .estimate - 1.96 * .se,
            ymax = .estimate + 1.96 * .se,
            fill = .data[[by_var_name]]),
        alpha = 0.15,
        color = NA) +
        geom_line(linewidth = 1) +
        labs(title = plot_title,
             x = label,
             y = "Partial effect",
             color = NULL,
             fill = NULL,
             linetype = NULL) +
        theme_minimal()
    
  } else {
    
    p <- ggplot(combined_df,
                aes(x = .data[[var]],
                    y = .estimate,
                    linetype = Model)) +
         geom_ribbon(data = subset(combined_df, Model == "Model 1"),
                     aes(ymin = .estimate - 1.96 * .se,
                     ymax = .estimate + 1.96 * .se),
                     alpha = 0.15,
                     color = NA) +
         geom_line(linewidth = 1, color = "black") +
         labs(title = plot_title,
              x = label,
              y = "Partial effect",
              linetype = NULL) +
         theme_minimal()
  }
  
  p
}

# -------------------------
# 4. Main plot function
# -------------------------

make_smooth_plots <- function(models, model_set_name, smooth_variables) {
  
  plot_smooths <- list()
  
  for (var in names(smooth_variables)) {
    
    if (var == "standard_property_value") {
      
      p <- make_standard_property_smooth_plot(models = models, model_set_name = model_set_name)
      
    } else {
      
      p <- make_single_smooth_plot(models = models,
                                   model_set_name = model_set_name,
                                   var = var,
                                   label = smooth_variables[[var]])
    }
    
    if (!is.null(p)) {plot_smooths[[var]] <- p}
  }
  
  plot_smooths
}

# -------------------------
# 5. Create plots
# -------------------------

plot_smooths_direct <- make_smooth_plots(models = models_direct,
                                         model_set_name = "Direct lot model",
                                         smooth_variables = smooth_variables)

plot_smooths_standard_property <- make_smooth_plots(models = models_standard_property,
                                                    model_set_name = "Standard property model",
                                                    smooth_variables = smooth_variables)

plot_smooths_combined <- make_smooth_plots(models = models_combined,
                                           model_set_name = "Combined model",
                                           smooth_variables = smooth_variables)

# -------------------------
# 6. Save plots
# -------------------------

saveRDS(plot_smooths_direct, "saved_files/plot_smooths_direct.rds")
saveRDS(plot_smooths_standard_property, "saved_files/plot_smooths_standard_property.rds")
saveRDS(plot_smooths_combined, "saved_files/plot_smooths_combined.rds")

# -------------------------
# 7. Display selected plots (due to figure size mismatch some are moved to subsequent chunk)
# -------------------------
#

wrap_plots(plot_smooths_standard_property, ncol = 2)

```


```{r Plot smooths direct and combined, eval=TRUE, echo=TRUE, warning=FALSE, results = 'hide', fig.width=20, fig.height=30}
wrap_plots(plot_smooths_direct, ncol = 2)
wrap_plots(plot_smooths_combined, ncol = 2)
```

#### Factor effects
The code below extracts and plots the factor effects for the combined model. A point is shown for the effect of each model but error bars are shown only for Model 1. 

```{r Factor effects lot models, eval=TRUE, echo=TRUE, warning=FALSE, results = 'hide', fig.width=20, fig.height=20}

factor_variables <- c(house_type = "House type",
                      municipality_type = "Municipality type",
                      type = "Type",
                      roof_type = "Roof type",
                      wall_type = "Wall type",
                      heating_type = "Heating type")

make_termwise_df <- function(model, model_data, variable, interaction_by = "municipality_type") {

  v_vals <- unique(model_data[[variable]])
  
  if (is.null(interaction_by)) {
    df <- data.frame(setNames(list(v_vals), variable), stringsAsFactors = FALSE)
  } else {
    m_vals <- unique(model_data[[interaction_by]])
    grid_list <- setNames(list(v_vals, m_vals), c(variable, interaction_by))
    df <- expand.grid(grid_list, stringsAsFactors = FALSE)
    df[[interaction_by]] <- factor(df[[interaction_by]], 
                                   levels = levels(model_data[[interaction_by]]))
  }
  for (var in setdiff(all.vars(formula(model)), names(df))) {
    x <- model_data[[var]]
    if (is.factor(x)) {
      df[[var]] <- factor(rep(levels(x)[1], nrow(df)), levels = levels(x))
    } else {
      df[[var]] <- 0
    }
  }
  return(df)
}

combine_all_term_effects <- function(df, term, interaction_by = "municipality_type") {
  pattern <- paste0("^", term, "$|^", term, ":" ,"|:", term, "$")
  cols_to_sum <- grep(pattern, names(df), value = TRUE)
  df[[paste0("effect_", term)]] <- rowSums(as.matrix(df[cols_to_sum]))
  return(df)
}

combine_all_term_se <- function(df, term, interaction_by = "municipality_type") {
  pattern <- paste0("^", term, "$|^", term, ":" ,"|:", term, "$")
  cols_to_sum <- grep(pattern, names(df), value = TRUE)
  df[[paste0("se_", term)]] <- sqrt(rowSums(as.matrix(df[cols_to_sum])^2))
  return(df)
}

plot_term_effect <- function(df, term, interaction_by = "municipality_type") {
  
  effect_col <- paste0("effect_", term)
  lower_col <- paste0("lower_", effect_col)
  upper_col <- paste0("upper_", effect_col)
  df_sub <- df %>% filter(term == !!term)
  plot_title <- paste("Partial effect of", gsub("_", " ", term))

  p <- ggplot(df_sub, aes(x = .data[[term]], 
                          y = .data[[effect_col]], 
                          shape = Model,
                          group = interaction(Model, if (identical(term, interaction_by) || 
                                                        !(interaction_by %in% names(df_sub))) NULL 
                                                     else .data[[interaction_by]]))) +
      geom_pointrange(data = subset(df_sub, Model == "Model 1"),
                      aes(ymin = .data[[lower_col]], ymax = .data[[upper_col]]),
                      position = position_dodge(width = 0.5),
                      size = 0.8) +
      geom_point(data = subset(df_sub, Model != "Model 1"),
                 position = position_dodge(width = 0.5),
                 size = 2.5) +
      coord_flip() +
      labs(x = NULL, y = "Partial effect", title = plot_title, shape = "Model") +
      theme(axis.text.y = element_text(angle = 45, hjust = 1))

  if (interaction_by %in% names(df_sub)) {
      color_var <- if (term == interaction_by) term else interaction_by
      p <- p + aes(color = .data[[color_var]]) +
      scale_color_brewer(palette = "Dark2", name = "Municipality type")
  }

  return(p)
}

factor_dfs <- list()

for (model_name in names(models_combined)) {

  model <- models_combined[[model_name]]
  model_data <- model.frame(model)
  relevant_factors <- factor_variables[names(factor_variables) %in% all.vars(formula(model))]

  for (var in names(relevant_factors)) {

    interaction <- if (var == "municipality_type") NULL else "municipality_type"
    df <- make_termwise_df(model, model_data, var, interaction_by = interaction)
    df$term <- var

    terms <- predict.gam(model, df, type = "terms", se.fit = TRUE)
    df_predictions <- as.data.frame(terms$fit)
    df_se <- as.data.frame(terms$se.fit)

    combined_predictions <- combine_all_term_effects(df_predictions, var, interaction_by = interaction) %>% 
                            select(starts_with("effect_"))
    
    combined_se <- combine_all_term_se(df_se, var, interaction_by = interaction) %>% 
                   select(starts_with("se_"))

    df_final <- df %>% select(all_of(c(var, interaction)), term) %>%
                       bind_cols(combined_predictions) %>%
                       bind_cols(combined_se) %>%
                       mutate(across(starts_with("effect_"),
                                     ~ . - 1.96 * get(str_replace(cur_column(), 
                                     "effect_", 
                                     "se_")),
                                     .names = "lower_{.col}")) %>%
                       mutate(across(starts_with("effect_"),
                                     ~ . + 1.96 * get(str_replace(cur_column(), "effect_", "se_")),
                                     .names = "upper_{.col}"))

    df_final$Model <- sub("^Model", "Model ", model_name)
    factor_dfs[[length(factor_dfs) + 1]] <- df_final
  }
}

df_all_factors <- bind_rows(factor_dfs)
plot_factors <- lapply(names(factor_variables), function(term) {plot_term_effect(df_all_factors, term)})
names(plot_factors) <- names(factor_variables)

wrap_plots(plot_factors, ncol = 2) +
           plot_layout(guides = "collect") &
           theme(legend.position = "bottom",
                 legend.text = element_text(size = 12),
                 legend.title = element_text(size = 13, face = "bold"),
                 legend.box = "horizontal",
                 legend.justification = "center")

```

## 4.4 Predictive performance
The code below extracts various residual types for each type of lot model and calculates different performance measures for the training and validation set. Note that using N-fold cross-validation means that there are N-1 estimates for each training point, and the training error is simply taken as the average over these. 

```{r Performance lot sales, eval=TRUE, echo=TRUE, warning=FALSE}

# -------------------------
# 1. Residuals for lot-sales evaluation
# -------------------------

get_lot_residuals <- function(models, dataset_lot_sales_cv, model_approach,
                              fold_from_name = TRUE) {
  
  residual_list <- future_lapply(seq_along(models), function(i) {
    
    model_name <- names(models)[i]
    model <- models[[i]]
    
    fold <- if (fold_from_name) {
      as.integer(sub("^Model(\\d+).*$", "\\1", model_name))
    } else {
      i
    }
    
    predicted_sales_price <- as.vector(predict.gam(model, newdata = dataset_lot_sales_cv, type = "response"))
    
    tibble(model_approach = model_approach,
           Model = model_name,
           sales_price = dataset_lot_sales_cv$sales_price,
           sales_date_numeric = dataset_lot_sales_cv$sales_date_numeric,
           predicted_sales_price = predicted_sales_price,
           residual_sales_price = predicted_sales_price - dataset_lot_sales_cv$sales_price,
           residual_sales_price_percent = (predicted_sales_price - dataset_lot_sales_cv$sales_price) / dataset_lot_sales_cv$sales_price * 100,
           residual_sales_price_log = log(predicted_sales_price) - log(dataset_lot_sales_cv$sales_price),
           ratio = predicted_sales_price / dataset_lot_sales_cv$sales_price,
           validation = dataset_lot_sales_cv$validation_set == fold,
           id = dataset_lot_sales_cv$id,
           municipality_type = dataset_lot_sales_cv$municipality_type)
  })
  
  bind_rows(residual_list)
}

# -------------------------
# 2. Residuals for each approach
# -------------------------

residual_direct_lot <- get_lot_residuals(models = models_direct,
                                         dataset_lot_sales_cv = dataset_lot_sales_cv_standardized,
                                         model_approach = "Direct lot model")

residual_standard_property <- get_lot_residuals(models = models_standard_property,
                                                dataset_lot_sales_cv = dataset_lot_sales_cv_standardized,
                                                model_approach = "Standard property model")

residual_combined <- get_lot_residuals(models = models_combined,
                                       dataset_lot_sales_cv = dataset_lot_sales_cv_standardized,
                                       model_approach = "Combined model")

residual_df_lot_models <- bind_rows(residual_direct_lot,
                                    residual_standard_property,
                                    residual_combined)

saveRDS(residual_df_lot_models, "saved_files/residual_df_lot_models.rds")


# -------------------------
# 3. Performance helper
# -------------------------

summarise_lot_performance <- function(residual_df, group_vars) {
  
  residual_df %>% mutate(dataset_type = if_else(validation, "Validation", "Training")) %>%
                  group_by(across(all_of(group_vars))) %>%
                  summarise(MALE = mean(abs(residual_sales_price_log), na.rm = TRUE),
                            MAPE = mean(abs(residual_sales_price_percent), na.rm = TRUE),
                            MAE = mean(abs(residual_sales_price), na.rm = TRUE),
                            PM20 = mean(abs(residual_sales_price_percent) <= 20, na.rm = TRUE) * 100,
                            COD = mean(abs(ratio - median(ratio, na.rm = TRUE)), na.rm = TRUE) / median(ratio, na.rm = TRUE) * 100,
                            PRD = mean(ratio, na.rm = TRUE) /(sum(predicted_sales_price, na.rm = TRUE) / sum(sales_price, na.rm = TRUE)),
                            .groups = "drop")
}

# -------------------------
# 4. Countrywide and municipality-type performance
# -------------------------

performance_countrywide_lot_models <- summarise_lot_performance(residual_df_lot_models, group_vars = c("model_approach", "dataset_type")) %>%
                                      mutate(municipality_type = "Countrywide")

performance_municipality_type_lot_models <- summarise_lot_performance(residual_df_lot_models, group_vars = c("model_approach", "dataset_type", "municipality_type")) %>%
                                            mutate(PRD = NA_real_)

performance_lot_models <- bind_rows(performance_countrywide_lot_models, performance_municipality_type_lot_models)

municipality_levels <- c("Countrywide", sort(setdiff(unique(performance_lot_models$municipality_type), "Countrywide")))

performance_sorted_lot_models <- performance_lot_models %>%
                                 mutate(model_approach = factor(model_approach, levels = c("Direct lot model", "Standard property model", "Combined model")),
                                        dataset_type = factor(dataset_type, levels = c("Training", "Validation")),
                                        municipality_type = factor(municipality_type, levels = municipality_levels),
                                        MAE = round(MAE, 0),
                                        PRD = ifelse(is.na(PRD), "", round(PRD, 2))) %>%
                                 arrange(model_approach, municipality_type, dataset_type) %>%
                                 rename(Approach = model_approach,
                                        Type = dataset_type,
                                        `Municipality type` = municipality_type,
                                        `MAPE [%]` = MAPE,
                                        `MAE [DKK]` = MAE,
                                        `PM20 [%]` = PM20,
                                        `COD [%]` = COD)
       
saveRDS(performance_sorted_lot_models, "saved_files/performance_sorted_lot_models.rds")

# -------------------------
# 6. Tables
# -------------------------

kable(performance_sorted_lot_models,
      digits = 2,
      caption = "Lot-sales performance by model approach and municipality type",
      booktabs = TRUE) %>%
kable_styling(full_width = FALSE, position = "center")

tbl_latex_lot_models <- kable(performance_sorted_lot_models,
                              format = "latex",
                              booktabs = TRUE,
                              digits = 2,
                              caption = "Lot-sales performance by model approach and municipality type.",
                              label = "perf_lot_models") %>%
                        kable_styling(full_width = FALSE,
                                      position = "center",
                                      latex_options = c("hold_position"))

writeLines(tbl_latex_lot_models, "saved_files/performance_lot_models.tex")

# -------------------------
# 7. Validation residuals
# -------------------------

residual_df_validation_lot_models <- residual_df_lot_models %>%
                                     filter(validation) %>%
                                     mutate(pm20 = as.integer(abs(residual_sales_price_percent) <= 20),
                                            model_approach = factor(model_approach, levels = c("Direct lot model", "Standard property model", "Combined model")))


# -------------------------
# 8. Residual plots: log residuals
# -------------------------

plot_validation_residuals_log_lot_models <- ggplot(residual_df_validation_lot_models,
                                                   aes(x = sales_price / 1e6,
                                                       y = residual_sales_price_log,
                                                       color = model_approach)) +
                                            geom_point(alpha = 0.25, size = 0.25) +
                                            geom_smooth(aes(y = abs(residual_sales_price_log), linetype = "MALE"),
                                                        se = FALSE,
                                                        linewidth = 0.8) +
                                            scale_linetype_manual(name = NULL, values = c(MALE = "solid")) +
                                            ylim(-1, 1) +
                                            labs(x = "Sales price [million DKK]", 
                                                 y = "Log residuals", 
                                                 color = NULL) +
                                            theme_minimal()

saveRDS(plot_validation_residuals_log_lot_models, "saved_files/plot_validation_residuals_log_lot_models.rds")
print(plot_validation_residuals_log_lot_models)


# -------------------------
# 9. Residual plots: percentage residuals
# -------------------------

plot_validation_residuals_percent_lot_models <- ggplot(residual_df_validation_lot_models,
                                                       aes(x = sales_price / 1e6,
                                                           y = residual_sales_price_percent,
                                                           color = model_approach)) +
                                                geom_point(alpha = 0.25, size = 0.25) +
                                                geom_smooth(aes(y = abs(residual_sales_price_percent), linetype = "MAPE"),
                                                            se = FALSE,
                                                            linewidth = 0.8) +
                                                geom_smooth(aes(y = pm20 * 100, linetype = "PM20"),
                                                            se = FALSE,
                                                            linewidth = 0.8) +
                                                scale_linetype_manual(name = NULL, values = c(MAPE = "solid", PM20 = "dashed")) +
                                                ylim(-100, 100) +
                                                labs(x = "Sales price [million DKK]", 
                                                     y = "Percentage residuals [%]", 
                                                     color = NULL) +
                                                theme_minimal()

saveRDS(plot_validation_residuals_percent_lot_models, "saved_files/plot_validation_residuals_percent_lot_models.rds")
print(plot_validation_residuals_percent_lot_models)


# -------------------------
# 10. Residual plots: DKK residuals
# -------------------------

plot_validation_residuals_lot_models <- ggplot(residual_df_validation_lot_models,
                                               aes(x = sales_price / 1e6,
                                                   y = residual_sales_price / 1e6,
                                                   color = model_approach)) +
                                        geom_point(alpha = 0.25, size = 0.25) +
                                        geom_smooth(aes(y = abs(residual_sales_price) / 1e6, linetype = "MAE"),
                                                    se = FALSE,
                                                    linewidth = 0.8) +
                                        scale_linetype_manual(name = NULL, values = c(MAE = "solid")) +
                                        ylim(-3, 3) +
                                        labs(x = "Sales price [million DKK]",
                                             y = "Residuals [million DKK]",
                                             color = NULL) +
                                        theme_minimal()

saveRDS(plot_validation_residuals_lot_models, "saved_files/plot_validation_residuals_lot_models.rds")
print(plot_validation_residuals_lot_models)

# -------------------------
# 11. Error over time
# -------------------------

residual_df_lot_models <- residual_df_lot_models %>% mutate(error_type = if_else(validation, "Validation error", "Training error"),
                                                            model_approach = factor(model_approach, levels = c("Direct lot model", "Standard property model", "Combined model")))

plot_log_errors_lot_models <- ggplot(residual_df_lot_models) +
                              geom_smooth(aes(x = sales_date_numeric,
                                              y = abs(residual_sales_price_log),
                                              linetype = error_type,
                                              color = model_approach),
                                          se = FALSE,
                                          linewidth = 0.8) +
                              labs(x = "Sales date",
                                   y = "MALE",
                                   color = NULL,
                                   linetype = NULL) +
                              theme_minimal()

saveRDS(plot_log_errors_lot_models, "saved_files/plot_log_errors_lot_models.rds")
print(plot_log_errors_lot_models)

plot_percentage_errors_lot_models <- ggplot(residual_df_lot_models) +
                                     geom_smooth(aes(x = sales_date_numeric, y = 100 * as.integer(abs(residual_sales_price_percent) <= 20),
                                                 linetype = error_type,
                                                 color = model_approach),
                                                 se = FALSE,
                                                 linewidth = 0.8) +
                                     labs(x = "Sales date",
                                          y = "PM20 [%]",
                                          color = NULL,
                                          linetype = NULL) +
                                     theme_minimal()

saveRDS(plot_percentage_errors_lot_models, "saved_files/plot_percentage_errors_lot_models.rds")

print(plot_percentage_errors_lot_models)

plot_errors_lot_models <- ggplot(residual_df_lot_models) +
                          geom_smooth(aes(x = sales_date_numeric,
                                          y = abs(residual_sales_price) / 1e6,
                                          linetype = error_type,
                                          color = model_approach),
                                      se = FALSE,
                                      linewidth = 0.8) +
                          labs(x = "Sales date",
                               y = "MAE [million DKK]",
                               color = NULL,
                               linetype = NULL) +
                          theme_minimal()

saveRDS(plot_errors_lot_models, "saved_files/plot_errors_lot_models.rds")
print(plot_errors_lot_models)

```

## 4.5 Prediction stability
The code below evaluates the stability of predicted lot values across the fold-specific models for each of the three modelling approaches. All models within an ensemble are used to predict lot values for the same set of properties on the link scale. For each property, the mean prediction and the standard deviation across the fold-specific predictions are calculated. The standard deviation across the model ensembles is used as a measure of sensitivity to perturbations of the training data.

```{r Stability lot models, eval=TRUE, echo=TRUE, warning=FALSE}

# -------------------------
# 1. Settings
# -------------------------

cap <- 0.2

model_sets <- list("Direct lot model" = models_direct,
                   "Standard property model" = models_standard_property,
                   "Combined model" = models_combined)

# -------------------------
# 2. Prepare prediction datasets
# -------------------------

dataset_properties_combined_lot <- dataset_properties_standardized %>% 
                                   mutate(type = factor("Lot sale", levels = c("Lot sale", "Property sale")),
                                          type_ordered = ordered("Lot sale", levels = c("Lot sale", "Property sale")))

prediction_datasets <- list("Direct lot model" = dataset_properties,
                            "Standard property model" = dataset_properties,
                            "Combined model" = dataset_properties_combined_lot)

# -------------------------
# 3. Ensemble spread function
# -------------------------

spread_model_set <- function(model_set, dataset) {
  
  predictions <- imap(model_set, function(model, fold_name) {
    
    preds <- predict.gam(model,
                         newdata = dataset,
                         type = "link")
    
    tibble(!!paste0("pred_", fold_name) := as.vector(preds))
    
  })
  
  pred_matrix <- bind_cols(predictions)
  pred_mat <- as.matrix(pred_matrix)
  
  bind_cols(dataset, pred_matrix) %>%
  mutate(mean_Ensemble = rowMeans(pred_mat, na.rm = TRUE),
         sigma_Ensemble = apply(pred_mat, 1, sd, na.rm = TRUE))
}

# -------------------------
# 4. Predict for all three approaches
# -------------------------

if (load) {
  
  dataset_spread <- readRDS("saved_files/dataset_spread.rds")
  
} else {
  
  dataset_spread <- imap(
    model_sets,
    function(model_set, model_approach) {
      
      message("Predicting ensemble spread for ", model_approach, " ...")
      
      spread_model_set(model_set = model_set, dataset = prediction_datasets[[model_approach]])
      
    }
  )
  
  saveRDS(dataset_spread, "saved_files/dataset_spread.rds")
}

# -------------------------
# 5. Long-format spread data
# -------------------------

spread_long <- imap_dfr(
  dataset_spread,
  function(df, model_approach) {
    
    df %>% transmute(model_approach = model_approach, sigma_Ensemble = sigma_Ensemble)
    
  }
)

saveRDS(spread_long, "saved_files/spread_long.rds")

# -------------------------
# 6. Summary table
# -------------------------

spread_summary <- spread_long %>%
                  group_by(model_approach) %>%
                  summarise(mean_sigma = mean(sigma_Ensemble, na.rm = TRUE),
                            median_sigma = median(sigma_Ensemble, na.rm = TRUE),
                            p75_sigma = quantile(sigma_Ensemble, 0.75, na.rm = TRUE),
                            p90_sigma = quantile(sigma_Ensemble, 0.90, na.rm = TRUE),
                            p95_sigma = quantile(sigma_Ensemble, 0.95, na.rm = TRUE),
                            .groups = "drop")

saveRDS(spread_summary, "saved_files/spread_summary.rds")

# -------------------------
# 7. Density plot: all models in one plot
# -------------------------

plot_ensemble_density <- ggplot(filter(spread_long, sigma_Ensemble < 1.0),
                                aes(x = sigma_Ensemble,
                                    color = model_approach,
                                    fill = model_approach)) +
                         geom_density(alpha = 0.15,
                                      linewidth = 0.8,
                                      na.rm = TRUE) +
                         coord_cartesian(xlim = c(0, cap)) +
                         labs(title = "Ensemble spread across lot-value approaches",
                              x = "Ensemble spread (link scale)",
                              y = "Density",
                              color = NULL,
                              fill = NULL) +
                         theme_minimal() +
                         theme(legend.position = c(0.7, 0.75))

saveRDS(plot_ensemble_density, "saved_files/plot_ensemble_density.rds")
print(plot_ensemble_density)

# -------------------------
# 8. Base map
# -------------------------

base_map <- ggplot() +
            geom_contour(data = coordinate_grid,
                         aes(x = coordinate_east,
                             y = coordinate_north,
                             z = as.numeric(!ocean)),
                         color = "black",
                         linewidth = 0.1) +
            coord_fixed() +
            theme_void(base_size = 10)

# -------------------------
# 9. Geographic uncertainty plots
# -------------------------

plot_ensemble_geographic <- imap(
  dataset_spread,
  function(df, model_approach) {
    
    base_map +
    geom_point(data = df,
               aes(x = coordinate_east,
                   y = coordinate_north,
                   color = sigma_Ensemble),
               size = 0.2) +
    scale_color_viridis_c(option = "plasma",
                          name = "Ensemble uncertainty",
                          limits = c(0, cap),
                          oob = squish) +
    labs(title = paste(model_approach, "- Ensemble uncertainty")) +
    theme(legend.position = c(0.7, 0.75), 
          legend.box = "horizontal")
  }
)

saveRDS(plot_ensemble_geographic, "saved_files/plot_ensemble_geographic.rds")
wrap_plots(plot_ensemble_geographic, ncol = 2)

```

## 4.6 Property-lot-building decomposition
The code below decomposes predicted property values into estimated lot and building components for each modelling approach and cross-validation fold. For the direct and standard-property approaches, property values are taken from the previously fitted property models, while lot values are predicted using the corresponding lot model. For the combined approach, property, standardized-property, and lot values are generated from the same joint model by changing the sale type and building characteristics supplied at prediction.

Building value is defined as the difference between the predicted property value and predicted lot value. The same decomposition is also calculated for the standardized property. The resulting values are summarized using measures of central tendency, dispersion, selected quantiles, building shares, values per square metre, and the frequency of negative building-value estimates.

```{r Property, building and lot values, eval=TRUE, echo=TRUE, warning=FALSE}

# -------------------------
# Property, lot and building value decomposition
# -------------------------

# -------------------------
# Settings
# -------------------------

x_limits <- c(-2, 15)
density_adjust <- 1.5

building_value_limits <- c(-2, 8)
living_space_limits <- c(0, 400)
year_built_limits <- c(1900, 2024)

model_sets <- list(direct_lot_model = models_direct,
                   standard_property_model = models_standard_property,
                   combined_model = models_combined)

# -------------------------
# Helpers
# -------------------------

prepare_type <- function(data, type_value) {
  
  data %>% mutate(type = factor(type_value, levels = c("Lot sale", "Property sale")),
                  type_ordered = ordered(type, levels = c("Lot sale", "Property sale")))
}

get_property_value_vars <- function(fold_name) {
  
  property_model <- str_extract(fold_name, "Model\\d+")
  stopifnot(!is.na(property_model))
  
  list(property_value_var = paste0("property_value_", property_model),
       standard_property_value_var = paste0("standard_property_value_", property_model))
  
}

decompose_models <- function(model_set,
                             dataset_properties,
                             dataset_properties_standardized,
                             combined = FALSE) {
  
  imap(model_set, function(model, fold_name) {
    
    if (combined) {
      
      lot_value <- predict.gam(model,
                               newdata = prepare_type(dataset_properties_standardized, "Lot sale"),
                               type = "response") %>% 
                   as.vector()
      
      property_value <- predict.gam(model,
                                    newdata = prepare_type(dataset_properties, "Property sale"),
                                    type = "response") %>% 
                        as.vector()
      
      standard_property_value <- predict.gam(model,
                                             newdata = prepare_type(dataset_properties_standardized, "Property sale"),
                                             type = "response") %>% 
                                 as.vector()
      
    } else {
      
      vars <- get_property_value_vars(fold_name)
      
      lot_value <- predict.gam(model,
                               newdata = dataset_properties,
                               type = "response") %>% 
                   as.vector()
      
      property_value <- dataset_properties[[vars$property_value_var]]
      standard_property_value <- dataset_properties[[vars$standard_property_value_var]]
      
    }
    
    building_value <- property_value - lot_value
    standard_building_value <- standard_property_value - lot_value
    
    tibble(id = dataset_properties$id,
           property_value = property_value,
           standard_property_value = standard_property_value,
           lot_value = lot_value,
           building_value = building_value,
           standard_building_value = standard_building_value,
           building_share = building_value / property_value,
           standard_building_share = standard_building_value / standard_property_value,
           building_value_per_m2 = building_value / dataset_properties$living_space,
           standard_building_value_per_m2 = standard_building_value / 140,
           living_space = dataset_properties$living_space,
           year_built = dataset_properties$year_built)
    
  })
}


summarise_value <- function(x, prefix) {
  
  tibble(!!paste0(prefix, "_mean") := mean(x, na.rm = TRUE),
         !!paste0(prefix, "_median") := median(x, na.rm = TRUE),
         !!paste0(prefix, "_sd") := sd(x, na.rm = TRUE),
         !!paste0(prefix, "_mad") := mad(x, na.rm = TRUE),
         !!paste0(prefix, "_iqr") := IQR(x, na.rm = TRUE),
         !!paste0(prefix, "_q05") := quantile(x, 0.05, na.rm = TRUE),
         !!paste0(prefix, "_q10") := quantile(x, 0.10, na.rm = TRUE),
         !!paste0(prefix, "_q25") := quantile(x, 0.25, na.rm = TRUE),
         !!paste0(prefix, "_q75") := quantile(x, 0.75, na.rm = TRUE),
         !!paste0(prefix, "_q90") := quantile(x, 0.90, na.rm = TRUE),
         !!paste0(prefix, "_q95") := quantile(x, 0.95, na.rm = TRUE),
         !!paste0(prefix, "_min") := min(x, na.rm = TRUE),
         !!paste0(prefix, "_max") := max(x, na.rm = TRUE),
         !!paste0(prefix, "_n_negative") := sum(x < 0, na.rm = TRUE),
         !!paste0(prefix, "_share_negative") := mean(x < 0, na.rm = TRUE) * 100)
}

summarise_decomposition <- function(df) {
  
  bind_cols(tibble(n = nrow(df)),
            summarise_value(df$property_value, "property_value"),
            summarise_value(df$standard_property_value, "standard_property_value"),
            summarise_value(df$lot_value, "lot_value"),
            summarise_value(df$building_value, "building_value"),
            summarise_value(df$standard_building_value, "standard_building_value"),
            summarise_value(df$building_share, "building_share"),
            summarise_value(df$standard_building_share, "standard_building_share"),
            summarise_value(df$building_value_per_m2, "building_value_per_m2"),
            summarise_value(df$standard_building_value_per_m2, "standard_building_value_per_m2"))
}

density_layer <- function(df, var, label) {
  
  df_layer <- df %>%
              transmute(value = .data[[var]] / 1e6) %>%
              filter(is.finite(value),
                     value >= x_limits[1],
                     value <= x_limits[2])
  
  geom_density(data = df_layer,
               mapping = aes(x = value, color = label, fill = label),
               alpha = 0.3,
               linewidth = 0.8,
               adjust = density_adjust,
               na.rm = TRUE)
}

make_density_plot <- function(df, model_approach, fold_name) {
  
  ggplot() + density_layer(df, "property_value", "Property value") +
             density_layer(df, "standard_property_value", "Standard property value") +
             density_layer(df, "lot_value", "Lot value") +
             density_layer(df, "building_value", "Building value") +
             density_layer(df, "standard_building_value", "Standard building value") +
             coord_cartesian(xlim = x_limits) +
             scale_x_continuous(labels = label_number(suffix = "M")) +
             scale_color_discrete(name = NULL) +
             scale_fill_discrete(name = NULL) +
             labs(title = paste(model_approach, "-", fold_name),
                  x = "Value [million DKK]",
                  y = "Density") +
             theme_minimal()
}

# -------------------------
# Decomposition
# -------------------------

if (load) {
  
  property_decomposition <- readRDS("saved_files/property_decomposition.rds")
  
} else {
  
  property_decomposition <- list(
    direct_lot_model = decompose_models(
      model_set = model_sets$direct_lot_model,
      dataset_properties = dataset_properties,
      dataset_properties_standardized = dataset_properties_standardized,
      combined = FALSE
    ),
    standard_property_model = decompose_models(
      model_set = model_sets$standard_property_model,
      dataset_properties = dataset_properties,
      dataset_properties_standardized = dataset_properties_standardized,
      combined = FALSE
    ),
    combined_model = decompose_models(
      model_set = model_sets$combined_model,
      dataset_properties = dataset_properties,
      dataset_properties_standardized = dataset_properties_standardized,
      combined = TRUE
    )
  )
  
  saveRDS(property_decomposition, "saved_files/property_decomposition.rds")
}

# -------------------------
# Decomposition summary
# -------------------------

decomposition_summary <- imap_dfr(
  property_decomposition,
  function(model_set, model_approach) {
    
    imap_dfr(model_set, function(df, fold_name) {
      
      summarise_decomposition(df) %>%
        mutate(
          model_approach = model_approach,
          fold_name = fold_name,
          .before = 1
        )
    })
  }
)

saveRDS(decomposition_summary, "saved_files/decomposition_summary.rds")

# -------------------------
# Compact building-value summary
# -------------------------

building_value_summary <- decomposition_summary %>%
                          select(model_approach,
                                 fold_name,
                                 n,
                              
                                 building_value_median,
                                 building_value_iqr,
                                 building_value_mad,
                                 building_value_q10,
                                 building_value_q90,
                                 building_value_n_negative,
                                 building_value_share_negative,
                                
                                 standard_building_value_median,
                                 standard_building_value_iqr,
                                 standard_building_value_mad,
                                 standard_building_value_q10,
                                 standard_building_value_q90,
                                 standard_building_value_n_negative,
                                 standard_building_value_share_negative,
                                
                                 building_share_median,
                                 building_share_iqr,
                                 standard_building_share_median,
                                 standard_building_share_iqr,
                                
                                 building_value_per_m2_median,
                                 building_value_per_m2_iqr,
                                 standard_building_value_per_m2_median,
                                 standard_building_value_per_m2_iqr)

saveRDS(building_value_summary, "saved_files/building_value_summary.rds")

# -------------------------
# Density plots
# -------------------------

plot_value_densities <- imap(
  property_decomposition,
  function(model_set, model_approach) {
    
    imap(
      model_set,
      function(df, fold_name) {
        make_density_plot(df, model_approach, fold_name)
      }
    )
  }
)

saveRDS(plot_value_densities, "saved_files/plot_value_densities.rds")


# -------------------------
# Building diagnostics
# -------------------------

building_diagnostics_long <- imap_dfr(
  property_decomposition,
  function(model_set, model_approach) {
    
    imap_dfr(model_set, function(df, fold_name) {
      
      df %>% mutate(model_approach = model_approach,
                    fold_name = fold_name,
                    building_value_million = building_value / 1e6,
                    standard_building_value_million = standard_building_value / 1e6)
    })
  }
)

saveRDS(building_diagnostics_long, "saved_files/building_diagnostics_long.rds")

make_building_diagnostic_plot <- function(x_var, x_label, x_limits_plot) {
  
  plot_df <- building_diagnostics_long %>%
             filter(is.finite(.data[[x_var]]),
                    is.finite(building_value_million),
                    .data[[x_var]] >= x_limits_plot[1],
                    .data[[x_var]] <= x_limits_plot[2],
                    building_value_million >= building_value_limits[1],
                    building_value_million <= building_value_limits[2])
  
  ggplot(plot_df,
         aes(x = .data[[x_var]],
             y = building_value_million,
             color = model_approach,
             linetype = fold_name)) +
         geom_smooth(se = FALSE, linewidth = 1) +
         coord_cartesian(xlim = x_limits_plot,
                         ylim = building_value_limits) +
         labs(title = paste("Building value vs", x_label),
              x = x_label,
              y = "Building value [million DKK]",
              color = NULL,
              linetype = "Fold") +
         theme_minimal()
}

plot_building_value_living_space <- make_building_diagnostic_plot(x_var = "living_space",
                                                                  x_label = "Living space [m²]",
                                                                  x_limits_plot = living_space_limits)

plot_building_value_year_built <- make_building_diagnostic_plot(x_var = "year_built",
                                                                x_label = "Year built",
                                                                x_limits_plot = year_built_limits)

saveRDS(plot_building_value_living_space,"saved_files/plot_building_value_living_space.rds") 

saveRDS(plot_building_value_year_built, "saved_files/plot_building_value_year_built.rds")

print(plot_building_value_living_space)
print(plot_building_value_year_built)
```



```{r Plot distributions property building and lot values, eval=TRUE, echo=TRUE, warning=FALSE, results = 'hide', fig.width=20, fig.height=30}

# -------------------------
# Combined density figure
# -------------------------

plot_value_densities_combined <- wrap_plots(flatten(plot_value_densities), ncol = 2) +
                                 plot_layout(guides = "collect") &
                                 theme(legend.position = "bottom",
                                       legend.box = "horizontal")

saveRDS(plot_value_densities_combined, "saved_files/plot_value_densities_combined.rds")

print(plot_value_densities_combined)
```

## 4.7 Standard-house sensitivty
The code below examines how sensitive the lot-value estimates are to the choice of standard house. Several alternative specifications are considered by varying the reference dwelling’s living area and construction year while keeping the remaining building characteristics fixed.

For each specification, the lot-sale and property data are standardized to the chosen reference house, and the fold-specific property models are used to calculate corresponding standard-property values. The original lot characteristics are then restored for the lot sales so that only the assumed building characteristics differ across specifications.

The standard-property and combined lot models are subsequently re-estimated for every cross-validation fold under each alternative specification. The resulting model ensembles and standardized datasets are stored together, allowing the estimated lot values, model effects, and predictive performance to be compared across different choices of standard-house parameters.

```{r Sensitivity to standard house parameters, eval=FALSE, echo=TRUE, warning=FALSE}

# -------------------------
# Sensitivity analysis: standard-house specification
# -------------------------

# -------------------------
# 1. Settings
# -------------------------

sensitivity_dir <- "saved_files/standard_house_sensitivity"

if (!dir.exists(sensitivity_dir)) {dir.create(sensitivity_dir, recursive = TRUE)}

standard_house_specs <- list(
  
  reference_140_1970 = list(year_built = 1970,
                            year_rebuilt = 1970,
                            living_space = 140,
                            lot_size = 800,
                            maximum_coverage_percentage = 30,
                            number_of_divisions = 1,
                            number_of_bathrooms = 1,
                            house_type = "Detached",
                            roof_type = "Fiber cement",
                            wall_type = "Brick",
                            heating_type = "District heating"),
  
  small_100_1970 = list(year_built = 1970,
                        year_rebuilt = 1970,
                        living_space = 100,
                        lot_size = 800,
                        maximum_coverage_percentage = 30,
                        number_of_divisions = 1,
                        number_of_bathrooms = 1,
                        house_type = "Detached",
                        roof_type = "Fiber cement",
                        wall_type = "Brick",
                        heating_type = "District heating"),
  
  large_180_1970 = list(year_built = 1970,
                        year_rebuilt = 1970,
                        living_space = 180,
                        lot_size = 800,
                        maximum_coverage_percentage = 30,
                        number_of_divisions = 1,
                        number_of_bathrooms = 1,
                        house_type = "Detached",
                        roof_type = "Fiber cement",
                        wall_type = "Brick",
                        heating_type = "District heating"),
  
  newer_140_2000 = list(year_built = 2000,
                        year_rebuilt = 2000,
                        living_space = 140,
                        lot_size = 800,
                        maximum_coverage_percentage = 30,
                        number_of_divisions = 1,
                        number_of_bathrooms = 1,
                        house_type = "Detached",
                        roof_type = "Fiber cement",
                        wall_type = "Brick",
                        heating_type = "District heating"),
  
  older_140_1950 = list(year_built = 1950,
                        year_rebuilt = 1970,
                        living_space = 140,
                        lot_size = 800,
                        maximum_coverage_percentage = 30,
                        number_of_divisions = 1,
                        number_of_bathrooms = 1,
                        house_type = "Detached",
                        roof_type = "Fiber cement",
                        wall_type = "Brick",
                        heating_type = "District heating")
  
)

# -------------------------
# 2. Standardization helper
# -------------------------

standardize_property_spec <- function(data, reference_data, spec) {
  
  data %>% mutate(year_built = spec$year_built,
                  year_rebuilt = spec$year_rebuilt,
                  living_space = spec$living_space,
                  lot_size = spec$lot_size,
                  maximum_coverage_percentage = spec$maximum_coverage_percentage,
                  number_of_divisions = spec$number_of_divisions,
                  number_of_bathrooms = spec$number_of_bathrooms,
                  house_type = factor(spec$house_type, levels = levels(factor(reference_data$house_type))),
                  roof_type = factor(spec$roof_type, levels = levels(factor(reference_data$roof_type))),
                  wall_type = factor(spec$wall_type, levels = levels(factor(reference_data$wall_type))),
                  heating_type = factor(spec$heating_type, levels = levels(factor(reference_data$heating_type))))
  
}

restore_lot_variables <- function(standardized_lot_data, original_lot_data) {
  
  standardized_lot_data %>% mutate(lot_size = original_lot_data$lot_size,
                                   maximum_coverage_percentage = original_lot_data$maximum_coverage_percentage,
                                   number_of_divisions = original_lot_data$number_of_divisions)
  
}

prepare_sensitivity_datasets <- function(spec_name,
                                         spec,
                                         property_models,
                                         dataset_lot_sales_cv,
                                         dataset_properties,
                                         reference_data) {
  
  message("Preparing standardized datasets for ", spec_name, " ...")
  
  dataset_lot_sales_cv_standardized_spec <- standardize_property_spec(data = dataset_lot_sales_cv,
                                                                      reference_data = reference_data,
                                                                      spec = spec)
  
  dataset_properties_standardized_spec <- standardize_property_spec(data = dataset_properties,
                                                                    reference_data = reference_data,
                                                                    spec = spec)
  
  for (model_name in names(property_models)) {
    
    dataset_lot_sales_cv_standardized_spec[[paste0("standard_property_value_", model_name)]] <- as.vector(
      predict.gam(property_models[[model_name]], newdata = dataset_lot_sales_cv_standardized_spec, type = "response")
      )
    
    dataset_properties_standardized_spec[[paste0("standard_property_value_", model_name)]] <- as.vector(
      predict.gam(property_models[[model_name]], newdata = dataset_properties_standardized_spec, type = "response")
      )
  }
  
  dataset_lot_sales_cv_standardized_spec <- restore_lot_variables(standardized_lot_data = dataset_lot_sales_cv_standardized_spec,
                                                                  original_lot_data = dataset_lot_sales_cv)
  
  list(lot_sales_standardized = dataset_lot_sales_cv_standardized_spec,
       properties_standardized = dataset_properties_standardized_spec)
  
}

# -------------------------
# 3. Model-fitting helpers
# -------------------------

fit_standard_property_fold_sensitivity <- function(fold, data, k_spatial) {
  
  training_subset <- data %>% filter(validation_set != fold)
  
  standard_property_var <- paste0("standard_property_value_Model", fold)
  
  formula_standard_property <- as.formula(paste0("sales_price ~ ",
                                                 "s(", standard_property_var, ", k = 10) + ",
                                                 "s(lot_size, k = 6, pc = 800) + ",
                                                 "s(maximum_coverage_percentage, k = 3, pc = 30) + ",
                                                 "s(number_of_divisions, k = 3, pc = 1) + ",
                                                 "s(coordinate_east, coordinate_north, k = ", k_spatial,
                                                 ", pc = c(697000, 6150000))"))
  
  bam(formula_standard_property,
      data = training_subset,
      family = tw(link = log),
      method = "fREML")
  
}

fit_combined_fold_sensitivity <- function(fold,
                                          lot_data,
                                          property_data,
                                          k_spatial,
                                          k_other = 10) {
  
  lot_train <- lot_data %>% filter(validation_set != fold)
  
  property_train <- property_data %>% filter(validation_set != fold)
  
  training_subset <- bind_rows(lot_train, property_train)
  
  bam(sales_price ~
      s(coordinate_east, coordinate_north, k = k_spatial, pc = c(697000, 6150000)) +
      s(coordinate_east, coordinate_north, k = k_spatial, by = type_ordered, pc = c(697000, 6150000)) +
      s(sales_date_numeric, k = k_other, by = municipality_type, pc = 2024) +
      s(year_built, k = k_other, by = municipality_type, pc = 1970) +
      s(year_rebuilt, k = k_other, by = municipality_type, pc = 1970) +
      s(living_space, k = k_other, by = municipality_type, pc = 140) +
      s(lot_size, k = k_other, by = municipality_type, pc = 800) +
      s(coast_distance, k = k_other, pc = 1500) +
      s(lake_distance, k = k_other, pc = 500) +
      s(stream_distance, k = k_other, pc = 1000) +
      s(road_distance, k = k_other, pc = 2000) +
      s(highway_distance, k = k_other, pc = 2000) +
      s(train_station_distance, k = k_other, pc = 2000) +
      s(railway_distance, k = k_other, pc = 1000) +
      s(windturbine_distance, k = k_other, pc = 3000) +
      s(ocean_view, k = k_other, pc = 0) +
      s(lake_view, k = k_other, pc = 0) +
      s(forest_size, k = k_other, pc = 0) +
      s(number_of_bathrooms, k = 2, by = municipality_type, pc = 1) +
      house_type * municipality_type +
      roof_type * municipality_type +
      wall_type * municipality_type +
      heating_type * municipality_type +
      type,
      data = training_subset,
      chunk.size = nrow(training_subset),
      family = tw(link = log),
      method = "fREML")
  
}

fit_model_set_sensitivity <- function(fit_fun, N, ...) {
  
  models <- future_lapply(seq_len(N), function(fold) {
    message("Fitting fold ", fold, " ...")
    fit_fun(fold = fold, ...)
  })
  
  names(models) <- paste0("Model", seq_len(N))
  models
}

# -------------------------
# 4. Run sensitivity analysis
# -------------------------

if (load) {
  
  sensitivity_results <- readRDS(file.path(sensitivity_dir, "sensitivity_results.rds"))
  
} else {
  
  sensitivity_results <- imap(standard_house_specs,
                              
    function(spec, spec_name) {
      
      message("Running standard-house sensitivity: ", spec_name)
      
      datasets_spec <- prepare_sensitivity_datasets(spec_name = spec_name,
                                                    spec = spec,
                                                    property_models = models,
                                                    dataset_lot_sales_cv = dataset_lot_sales_cv,
                                                    dataset_properties = dataset_properties,
                                                    reference_data = dataset)
      
      models_standard_property_spec <- fit_model_set_sensitivity(fit_fun = fit_standard_property_fold_sensitivity,
                                                                 N = N,
                                                                 data = datasets_spec$lot_sales_standardized,
                                                                 k_spatial = k_spatial)
      
      models_combined_spec <- fit_model_set_sensitivity(fit_fun = fit_combined_fold_sensitivity,
                                                        N = N,
                                                        lot_data = datasets_spec$lot_sales_standardized,
                                                        property_data = dataset_property_sales_cv,
                                                        k_spatial = k_spatial,
                                                        k_other = k_other)
      
      list(spec = spec,
           datasets = datasets_spec,
           models_standard_property = models_standard_property_spec,
           models_combined = models_combined_spec)
      
    }
  )
  
  saveRDS(sensitivity_results, file.path(sensitivity_dir, "sensitivity_results.rds"))
  
}


```

## 4.8 Illustrative valuations
The code below performs valuations using the combined model framework on a set of simplified, fictitious properties, which are described in the main paper. 

```{r Model interpretation, eval=TRUE, echo=TRUE, warning=FALSE}

make_interpretation_results <- function(model, model_name) {
  
  model_data <- model.frame(model)
  
  # generate fictitious properties
  grid <- data.frame(coordinate_east = c(697000, 462037, 469142, 725599),
                     coordinate_north = c(6150000, 6150225, 6322298, 6185920))
  
  point_constraints <- list()
  
  for (sm in model$smooth) {if (!is.null(sm$point.con)) {point_constraints <- c(point_constraints, sm$point.con)}}
  
  vars_in_model <- all.vars(formula(model))
  
  # fill missing variables in grid
  for (var in setdiff(vars_in_model, names(grid))) {
    
    if (is.factor(model_data[[var]])) {
      
      grid[[var]] <- factor(rep(levels(model_data[[var]])[1], nrow(grid)), levels = levels(model_data[[var]]))
      
    } else {
      value <- if (var %in% names(point_constraints)) {
        point_constraints[[var]]
      } else {
        mean(model_data[[var]], na.rm = TRUE)
      }
      grid[[var]] <- rep(value, nrow(grid))
    }
  }
  
  # make lot examples 
  grid$coast_distance[2:4] <- 100
  grid$municipality_type[2:4] <- factor(c("Provincial", "Rural", "Capital"), levels = levels(model_data$municipality_type))
  grid$location <- c("Reference", "Esbjerg", "Nordjylland", "Strandvejen")
  
  #add rows for property case
  grid <- bind_rows(grid, grid)
  grid$year_built[6:8] <- 2020 #place a newer house on the properties
  grid$type[5:8] <- factor("Property sale", levels = levels(model_data$type))
  grid$type_ordered[5:8] <- ordered("Property sale", levels = levels(model_data$type))
  
  terms <- as.data.frame(predict.gam(model, grid, type = "terms"))
  
  # helper for summing possibly missing columns
  sum_terms <- function(df, cols) {
    cols_present <- intersect(cols, names(df))
    if (length(cols_present) == 0) {return(rep(0, nrow(df)))}
    rowSums(df[, cols_present, drop = FALSE])
  }
  
  adjustments <- terms %>%
                 mutate(standard_price = exp(as.vector(coef(model))[1]),
      
                        adjustment_coordinates = exp(`s(coordinate_east,coordinate_north)`),
                        
                        adjustment_type_base = exp(type),
                        
                        adjustment_type_coordinates = exp(`s(coordinate_east,coordinate_north):type_orderedProperty sale`),
      
                        # Smooths with municipality_type interactions – 
                        #might as well sum over all is it only non-zero for the relevant smooth
                         
                        adjustment_sales_date = exp(`s(sales_date_numeric):municipality_typeCapital` +
                                                    `s(sales_date_numeric):municipality_typeRural` +
                                                    `s(sales_date_numeric):municipality_typeCommuter` +
                                                    `s(sales_date_numeric):municipality_typeProvincial` +
                                                    `s(sales_date_numeric):municipality_typeLarge city`),
      
                        adjustment_year_built = exp(`s(year_built):municipality_typeCapital` +
                                                    `s(year_built):municipality_typeRural` +
                                                    `s(year_built):municipality_typeCommuter` +
                                                    `s(year_built):municipality_typeProvincial` +
                                                    `s(year_built):municipality_typeLarge city`),
      
                        adjustment_year_rebuilt = exp(`s(year_rebuilt):municipality_typeCapital` +
                                                      `s(year_rebuilt):municipality_typeRural` +
                                                      `s(year_rebuilt):municipality_typeCommuter` +
                                                      `s(year_rebuilt):municipality_typeProvincial` +
                                                      `s(year_rebuilt):municipality_typeLarge city`),
      
                        adjustment_living_space = exp(`s(living_space):municipality_typeCapital` +
                                                      `s(living_space):municipality_typeRural` +
                                                      `s(living_space):municipality_typeCommuter` +
                                                      `s(living_space):municipality_typeProvincial` +
                                                      `s(living_space):municipality_typeLarge city`),
      
                        adjustment_lot_size = exp(`s(lot_size):municipality_typeCapital` +
                                                  `s(lot_size):municipality_typeRural` +
                                                  `s(lot_size):municipality_typeCommuter` +
                                                  `s(lot_size):municipality_typeProvincial` +
                                                  `s(lot_size):municipality_typeLarge city`),
                                              
                        adjustment_coast_distance = exp(`s(coast_distance)`),
          
                        adjustment_lake_distance = exp(`s(lake_distance)`),
          
                        adjustment_stream_distance = exp(`s(stream_distance)`),
          
                        adjustment_road_distance = exp(`s(road_distance)`),
          
                        adjustment_highway_distance = exp(`s(highway_distance)`),
          
                        adjustment_train_station_distance = exp(`s(train_station_distance)`),
          
                        adjustment_railway_distance = exp(`s(railway_distance)`),
          
                        adjustment_windturbine_distance = exp(`s(windturbine_distance)`),
          
                        adjustment_ocean_view = exp(`s(ocean_view)`),
          
                        adjustment_lake_view = exp(`s(lake_view)`),
          
                        adjustment_forest_size = exp(`s(forest_size)`),
          
                        adjustment_number_of_bathrooms = exp(`s(number_of_bathrooms):municipality_typeCapital` +
                                                             `s(number_of_bathrooms):municipality_typeRural` +
                                                             `s(number_of_bathrooms):municipality_typeCommuter` +
                                                             `s(number_of_bathrooms):municipality_typeProvincial` +
                                                             `s(number_of_bathrooms):municipality_typeLarge city`),
      
                        #Factor terms with interactions
                        adjustment_municipality_type = exp(municipality_type),
                        adjustment_house_type = exp(house_type + `house_type:municipality_type`),
                        adjustment_roof_type = exp(roof_type + `municipality_type:roof_type`),
                        adjustment_wall_type = exp(wall_type + `municipality_type:wall_type`),
                        adjustment_heating_type = exp(heating_type + `municipality_type:heating_type`),
                        

                        adjustment_location = adjustment_coordinates *
                                              adjustment_sales_date *
                                              adjustment_lot_size *
                                              adjustment_coast_distance *
                                              adjustment_lake_distance *
                                              adjustment_stream_distance *
                                              adjustment_road_distance *
                                              adjustment_highway_distance *
                                              adjustment_train_station_distance *
                                              adjustment_railway_distance *
                                              adjustment_windturbine_distance *
                                              adjustment_ocean_view *
                                              adjustment_lake_view *
                                              adjustment_forest_size *
                                              adjustment_municipality_type,
      
                        adjustment_building = adjustment_type_base * 
                                              adjustment_type_coordinates * 
                                              adjustment_year_built *
                                              adjustment_year_rebuilt *
                                              adjustment_living_space *
                                              adjustment_house_type *
                                              adjustment_roof_type *
                                              adjustment_wall_type *
                                              adjustment_heating_type * 
                                              adjustment_number_of_bathrooms,
      
                        adjustment_total = adjustment_location * adjustment_building,
                        
                        predicted_price = standard_price * adjustment_total) %>%
    
                        select(standard_price,
                               adjustment_type_base,
                               adjustment_type_coordinates,
                               adjustment_municipality_type,
                               adjustment_coordinates,
                               adjustment_sales_date,
                               adjustment_year_built,
                               adjustment_year_rebuilt,
                               adjustment_living_space,
                               adjustment_lot_size,
                               adjustment_coast_distance,
                               adjustment_lake_distance,
                               adjustment_stream_distance,
                               adjustment_road_distance,
                               adjustment_highway_distance,
                               adjustment_train_station_distance,
                               adjustment_railway_distance,
                               adjustment_windturbine_distance,
                               adjustment_ocean_view,
                               adjustment_lake_view,
                               adjustment_forest_size,
                               adjustment_number_of_bathrooms,
                               adjustment_house_type,
                               adjustment_roof_type,
                               adjustment_wall_type,
                               adjustment_heating_type,
                               adjustment_location,
                               adjustment_building,
                               adjustment_total,
                               predicted_price)
  
  grid_out <- grid %>% select(coast_distance, year_built, municipality_type, location, type)
  
  bind_cols(grid_out, adjustments) %>% mutate(Model = model_name, .before = 1)
}

results_all_models <- bind_rows(
  lapply(names(models_combined), function(model_name) {
    make_interpretation_results(models_combined[[model_name]], model_name)
  })
)

saveRDS(results_all_models, "saved_files/results_interpretation_combined_models.rds")

```


# 5. Paper III: On the Accuracy of Hedonic Real Estate Models 

## 5.1 Learning curves
The code below fits a list of models to increasing fractions of the training set and assesses their validation MALE. Learning curves for each municipality type are then plotted.

```{r Learning curves, eval=FALSE, echo=TRUE, warning=FALSE, results = 'hide'}
fractions <- c(0.1, 0.2, 0.5, 1.0)
folds <- 1:N

model_grid_fraction <- expand.grid(fold = folds, fraction = fractions)

if (load) {
  models_fraction <- readRDS("saved_files/models_fraction.rds")
} else {
  
  models_fraction <- future_lapply(seq_len(nrow(model_grid_fraction)), function(idx) {
    row <- model_grid_fraction[idx, ]
    
    fit_fold(
      i = row$fold,
      data = dataset_property_sales_cv,
      k_spatial = k_spatial,
      k_other = k_other,
      fraction = row$fraction,
      checkpoint_prefix = paste0("fraction_", row$fraction)
    )
  })
  
  names(models_fraction) <- paste0(
    "Model", model_grid_fraction$fold,
    "_", model_grid_fraction$fraction
  )
  
  saveRDS(models_fraction, "saved_files/models_fraction.rds")
}


get_residuals_fraction <- function(models, dataset, grid_df) {
  
  residual_list <- future_lapply(seq_along(models), function(i) {
    
    model_name <- names(models)[i]
    model <- models[[i]]
    row <- grid_df[i, ]
    fold <- row$fold
    frac <- row$fraction
    
    predicted_sales_price <- predict.gam(model, newdata = dataset, type = "response")
    
    tibble(
      sales_price = dataset$sales_price,
      predicted_sales_price = predicted_sales_price,
      residual_sales_price = predicted_sales_price - dataset$sales_price,
      residual_sales_price_percent = (predicted_sales_price - dataset$sales_price) / dataset$sales_price * 100,
      residual_sales_price_log = log(predicted_sales_price) - log(dataset$sales_price),
      validation = dataset$validation_set == fold,
      municipality_type = dataset$municipality_type,
      Model = paste0("Model", fold),
      fraction = frac
    )
  })
  
  bind_rows(residual_list)
}

residual_df_fraction <- get_residuals_fraction(
  models = models_fraction,
  dataset = dataset_property_sales_cv,
  grid_df = model_grid_fraction
)

saveRDS(residual_df_fraction, "saved_files/residual_df_fraction.rds")

performance_fraction <- residual_df_fraction %>%
                        filter(validation) %>%
                        group_by(fraction, municipality_type) %>%
                        summarise(MALE = mean(abs(residual_sales_price_log)),
                                  .groups = "drop")

plot_learning_curves <- ggplot(performance_fraction,
                               aes(x = fraction, 
                                   y = MALE, 
                                   color = municipality_type)) +
                        geom_line(linewidth = 1) +
                        geom_point() +
                        scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
                        labs(x = "Percentage of training set",
                             y = "Validation MALE",
                             color = "Municipality type") +
                        theme_minimal()

saveRDS(plot_learning_curves, "saved_files/plot_learning_curves.rds")
print(plot_learning_curves)

```


## 5.2 Market noise and other models
The code below estimates the market noise based on properties sold twice and compares this to the error obtained using our GAM model as well as an XGboost model. 

```{r Market noise, eval=FALSE, echo=TRUE, warning=FALSE, results = 'hide'}


### Identify double sales and discard those with changes to interior between sales
dataset_sales <- dataset_sales %>%
                 group_by(id) %>%
                 mutate(n = n()) %>%
                 ungroup() %>%
                 arrange(id)

other_sales <- dataset_sales %>% filter(n != 2)
double_sales <- dataset_sales %>% filter(n == 2)

double_sales <- double_sales %>%
                group_by(id) %>%
                filter(year_rebuilt < 2014,
                       n_distinct(living_space) == 1,
                       n_distinct(lot_size) == 1,
                       n_distinct(year_built) == 1,
                       n_distinct(house_type) == 1,
                       n_distinct(number_of_floors) == 1,
                       n_distinct(number_of_rooms) == 1,
                       n_distinct(number_of_bathrooms) == 1,
                       n_distinct(roof_type) == 1,
                       n_distinct(heating_type) == 1,
                       n_distinct(wall_type) == 1) %>%
                ungroup()


### Fit models to other sales and evaluate on double sales

# 1. Indexation model

if (load){

  time_model <- readRDS("saved_files/time_model.rds")

} else{

  time_model <- gam(sales_price ~ 
                    municipality_type + 
                    s(sales_date_numeric, k = k_other, by = municipality_type, pc = 2024),
                    data = other_sales,
                    family = tw(link=log),
                    method = "REML")

  saveRDS(time_model, "saved_files/time_model.rds")

}


#index double sales
time_model_terms <- as.data.frame(predict.gam(time_model, double_sales, type = "terms"))
time_adjustment <- as.vector(rowSums(time_model_terms[, grepl("sales_date_numeric", colnames(time_model_terms)), drop = FALSE]))

double_sales <- double_sales %>% mutate(log_sales_price = log(sales_price),
                                        log_sales_price_indexed = log_sales_price - time_adjustment,
                                        sales_price_indexed = exp(log_sales_price_indexed))

#index other sales
time_model_terms <- as.data.frame(predict.gam(time_model, other_sales, type = "terms"))
time_adjustment <- as.vector(rowSums(time_model_terms[, grepl("sales_date_numeric", colnames(time_model_terms)), drop = FALSE]))
other_sales <- other_sales %>% mutate(log_sales_price = log(sales_price),
                                      log_sales_price_indexed = log_sales_price - time_adjustment,
                                      sales_price_indexed = exp(log_sales_price_indexed))

#calculate lagged values
double_sales_lagged <- double_sales %>%
                       arrange(id, sales_date) %>%
                       group_by(id) %>%
                       mutate(sales_price_diff_indexed_log = log_sales_price_indexed - lag(log_sales_price_indexed),
                              avg_price = (sales_price_indexed + lag(sales_price_indexed)) / 2,
                              sales_date_diff = sales_date_numeric - lag(sales_date_numeric)) %>%
                       ungroup()

# 2. GAM model

if (load){

  gam_model <- readRDS("saved_files/gam_model.rds")

} else{

  gam_model <- bam(sales_price_indexed ~ 
                   s(coordinate_east, coordinate_north, k = k_spatial, pc = c(697000, 6150000)) +
                   s(year_built, k = k_other, by = municipality_type, pc = 1970) +
                   s(year_rebuilt, k = k_other, by = municipality_type, pc = 1970) +
                   s(living_space, k = k_other, by = municipality_type, pc = 140) +
                   s(lot_size, k = k_other, by = municipality_type, pc = 800) +
                   s(coast_distance, k = k_other, pc = 1500) +
                   s(lake_distance, k = k_other, pc = 500) +
                   s(stream_distance, k = k_other, pc = 1000) +
                   s(road_distance, k = k_other, pc = 2000) +
                   s(highway_distance, k = k_other, pc = 2000) +
                   s(train_station_distance, k = k_other, pc = 2000) +
                   s(railway_distance, k = k_other, pc = 1000) +
                   s(windturbine_distance, k = k_other, pc = 3000) +
                   s(ocean_view, k = k_other, pc = 0) +
                   s(lake_view, k = k_other, pc = 0) +
                   s(forest_size, k = k_other, pc = 0) +
                   s(number_of_bathrooms, k = 2, by = municipality_type, pc = 1) +
                   house_type*municipality_type +
                   roof_type*municipality_type +
                   wall_type*municipality_type +
                   heating_type*municipality_type,
                   data = other_sales,
                   family = tw(link=log),
                   method = "fREML")


  saveRDS(gam_model, "saved_files/gam_model.rds")
}

double_sales <- double_sales %>% 
                mutate(predictions_gam = predict.gam(gam_model, double_sales, type = "link"),
                       residuals_gam =  log_sales_price_indexed - predictions_gam)

other_sales <- other_sales %>% 
               mutate(predictions_gam = predict.gam(gam_model, other_sales, type = "link"),
                      residuals_gam =  log_sales_price_indexed - predictions_gam)

# 3. XGboost Models (fitted to log sales prices and GAM residuals)

dummy_model_other <- dummyVars(~ roof_type +
                                 wall_type +
                                 heating_type +
                                 house_type +
                                 municipality_type + 
                                 municipality +
                                 province +
                                 region,
                               data = other_sales)

categorical_encoded_other <- predict(dummy_model_other, newdata = other_sales)

X_other <- other_sales %>%
           select(coordinate_east,
                  coordinate_north,
                  number_of_bathrooms,
                  number_of_rooms,
                  number_of_floors,
                  year_built,
                  year_rebuilt,
                  living_space,
                  lot_size,
                  coast_distance,
                  lake_distance,
                  stream_distance,
                  road_distance,
                  highway_distance,
                  train_station_distance,
                  railway_distance,
                  windturbine_distance,
                  forest_size,
                  ocean_view,
                  lake_view) %>%
            as.matrix()

X_full_other <- cbind(X_other, categorical_encoded_other)

if (load){

  xgb_model <- readRDS("saved_files/xgb_model.rds")
  xgb_model_residuals <- readRDS("saved_files/xgb_model_residuals.rds")

} else{

  #on log sales price
  dtrain <- xgb.DMatrix(data = X_full_other, label = other_sales$log_sales_price_indexed)

  xgb_model <- xgboost(
    data = dtrain,
    objective = "reg:squarederror",
    nrounds = 1000,
    max_depth = 12,
    eta = 0.1,
    subsample = 0.8,
    colsample_bytree = 0.8,
    verbose = 1
  )

  saveRDS(xgb_model, "saved_files/xgb_model.rds")
  
  #on log residuals
  dtrain <- xgb.DMatrix(data = X_full_other, label = other_sales$residuals_gam)

  xgb_model_residuals <- xgboost(
    data = dtrain,
    objective = "reg:squarederror",
    nrounds = 1000,
    max_depth = 12,
    eta = 0.1,
    subsample = 0.8,
    colsample_bytree = 0.8,
    verbose = 1
  )

  saveRDS(xgb_model_residuals, "saved_files/xgb_model_residuals.rds")

}


categorical_encoded_double <- predict(dummy_model_other, newdata = double_sales)

X_double <- double_sales %>%
            select(coordinate_east,
                   coordinate_north,
                   number_of_bathrooms,
                   number_of_rooms,
                   number_of_floors,
                   year_built,
                   year_rebuilt,
                   living_space,
                   lot_size,
                   coast_distance,
                   lake_distance,
                   stream_distance,
                   road_distance,
                   highway_distance,
                   train_station_distance,
                   railway_distance,
                   windturbine_distance,
                   forest_size,
                   ocean_view,
                   lake_view) %>%
            as.matrix()

X_full_double <- cbind(X_double, categorical_encoded_double)

missing_cols <- setdiff(colnames(X_full_other), colnames(X_full_double))

if (length(missing_cols) > 0) {
  
  X_full_double <- cbind(X_full_double,
                         matrix(0, nrow = nrow(X_full_double), ncol = length(missing_cols), dimnames = list(NULL, missing_cols)))
}

X_full_double <- X_full_double[, colnames(X_full_other), drop = FALSE]

double_sales <- double_sales %>% 
                mutate(residuals_xgb = log_sales_price_indexed - predict(xgb_model, X_full_double),
                       residuals_hybrid = residuals_gam - predict(xgb_model_residuals, X_full_double))


# 4. Noise model

if (load){

  noise_model <- readRDS("saved_files/noise_model.rds")

} else{

  noise_model <- gam(sales_price_diff_indexed_log ~ s(avg_price),
                     data = double_sales_lagged,
                     family = gaussian())


  saveRDS(noise_model, "saved_files/noise_model.rds")

}


### Make scatter plot with errors
double_sales_lagged$residuals_noise <- NA
double_sales_lagged$residuals_noise[which(!is.na(double_sales_lagged$sales_date_diff))] <- residuals(noise_model)/sqrt(2)

smooth_data <- bind_rows(double_sales_lagged %>% 
                         transmute(x = avg_price / 1e6, y = abs(residuals_noise), Model = "Market noise"),
                         double_sales %>%
                         transmute(x = sales_price / 1e6, y = abs(residuals_gam), Model = "GAM"),
                         double_sales %>%
                         transmute(x = sales_price / 1e6, y = abs(residuals_xgb), Model = "XGBoost"),
                         double_sales %>%
                         transmute(x = sales_price / 1e6, y = abs(residuals_hybrid), Model = "Hybrid model"))

plot_market_noise <- ggplot(data = double_sales_lagged,
                     aes(x = avg_price / 1e6, y = residuals_noise)) +
                     geom_point(size = 0.3) +
                     geom_smooth(data = smooth_data,
                                 aes(x = x, y = y, color = Model),
                                 method = "gam", formula = y ~ s(x),
                                 se = TRUE, 
                                 linewidth = 1) +
                     labs(x = "Average price [million DKK]",
                          y = "Log residuals",
                          color = "Model",
                          linetype = NULL) +
                     ylim(-1, 1) +
                     xlim(0, 15) +
                     theme_minimal() +
                     theme(text = element_text(size = 12),
                           legend.position = c(0.95, 0.65),
                           legend.title = element_text(size = 12),  
                           legend.text = element_text(size = 10))

saveRDS(plot_market_noise, "saved_files/plot_market_noise.rds")
print(plot_market_noise)



```


### 5.3 Number of folds for model ensemble
The code below fits the model ensemble for numbers of fold for the model ensembles and examines the ensemble spread af a function of this.

```{r Number of folds, eval=FALSE, echo=TRUE, warning=FALSE, results='hide'}
if (load) {
  models_folds <- readRDS("saved_files/models_folds.rds")
  data_folds <- readRDS("saved_files/data_folds.rds")
} else {

  N_values <- c(2,3,4,8)

  data_folds <- lapply(N_values, function(N_current) {assign_cv_folds(dataset_sales, N = N_current, seed = 123)})
  
  names(data_folds) <- paste0("N", N_values)

  models_folds <- lapply(seq_along(N_values), function(j) {
    
    N_current <- N_values[j]
    data_current <- data_folds[[j]]

    models_current <- future_lapply(1:N_current, function(i) {
      
      fit_fold(i = i,
               data = data_current,
               k_spatial = k_spatial,
               k_other = k_other,
               fraction = 1.0,
               checkpoint_prefix = paste0("N", N_current))
      })

    names(models_current) <- paste0("Model", 1:N_current)
    models_current
    
    })
  
  
  names(models_folds) <- paste0("N", N_values)

  saveRDS(models_folds, "saved_files/models_folds.rds")
  saveRDS(data_folds, "saved_files/data_folds.rds")
}

get_ensemble_spread <- function(model_list, dataset, N_value) {

  predictions <- imap(model_list, function(model, name) {
    preds <- predict.gam(model, dataset, type = "link", se.fit = FALSE)
    tibble(!!paste0("pred_", name) := as.vector(preds))
  })

  pred_matrix <- bind_cols(predictions)
  pred_cols <- grep("^pred_", names(pred_matrix), value = TRUE)

  spread_df <- bind_cols(dataset, pred_matrix) %>%
               rowwise() %>%
               mutate(sigma_Ensemble = sd(c_across(all_of(pred_cols)))) %>%
               ungroup() %>%
               mutate(N = N_value)

  spread_df
}

property_spread_folds <- bind_rows(
  lapply(names(models_folds), function(name) {
    N_current <- as.numeric(sub("^N", "", name))
    get_ensemble_spread(
      model_list = models_folds[[name]],
      dataset = dataset_properties,
      N_value = N_current
    )
  })
)

saveRDS(property_spread_folds, "saved_files/property_spread_folds.rds")

spread_summary <- property_spread_folds %>%
                  group_by(N) %>%
                  summarise(q05 = quantile(sigma_Ensemble, 0.05, na.rm = TRUE),
                            median = median(sigma_Ensemble, na.rm = TRUE),
                            q95 = quantile(sigma_Ensemble, 0.95, na.rm = TRUE),
                            mean = mean(sigma_Ensemble, na.rm = TRUE),
                            .groups = "drop")

saveRDS(spread_summary, "saved_files/spread_summary_folds.rds")

plot_folds_spread <- ggplot(spread_summary, aes(x = N)) +
                     geom_ribbon(aes(ymin = q05, ymax = q95), alpha = 0.2, fill = "grey70") +
                     geom_line(aes(y = median), linewidth = 1) +
                     geom_point(aes(y = median), size = 2) +
                     labs(x = "Number of folds",
                          y = expression(sigma[Ensemble]),
                          title = "Ensemble spread across valuation properties by number of folds")

saveRDS(plot_folds_spread, "saved_files/plot_folds_spread.rds")
print(plot_folds_spread)

```


## 5.4 Interplay between market noise and data scarcity
The code below examines the interplay between and noise and data scarcity by examining the validation MALE in the 2D space of KNN-averaged sales densities and sales prices. 

```{r Interplay, eval=FALSE, echo=TRUE, warning=FALSE, results = 'hide'}

# Parameters
K <- 100 #number of neighbors to average over
N <- 3 #number of clusters in each direction
log_scale <- TRUE #perform partitioning on log scale

# Split data
test_set <- dataset %>% filter(test_set)
training_set <- dataset %>% filter(validation_set != 1, !test_set)

# Predict log price
test_set$sales_price_predicted_log <- as.vector(
  predict.gam(models$Model1, test_set, type = "link")
)

# Find spatial neighbors
neighbors <- knn(
  data = training_set[, c("coordinate_east", "coordinate_north")],
  query = test_set[, c("coordinate_east", "coordinate_north")],
  k = K,
  radius = 0
)

distance_matrix <- neighbors$nn.dists
id_matrix <- neighbors$nn.idx

# Compute local sales densities
test_set$sales_density <- apply(distance_matrix, 1, function(dists) {
  K / (pi * (0.001 * max(dists))^2)
})

# Compute average sales price of neighbors
test_set$avg_sales_price <- vapply(
  1:nrow(test_set),
  function(j) mean(training_set$sales_price[id_matrix[j, ]]),
  numeric(1)
)

# Compute residuals in log space
test_set$resid_sales_price_log <- test_set$sales_price_predicted_log - log(test_set$sales_price)

#Find clusters
price_vec <- if (log_scale) log1p(test_set$avg_sales_price) else test_set$avg_sales_price
density_vec <- if (log_scale) log1p(test_set$sales_density) else test_set$sales_density
price_clusters <- kmeans(price_vec, centers = N)$cluster
density_clusters <- kmeans(density_vec, centers = N)$cluster
price_order <- rank(tapply(test_set$avg_sales_price, price_clusters, mean, na.rm = TRUE))
density_order <- rank(tapply(test_set$sales_density, density_clusters, mean, na.rm = TRUE))

test_set <- test_set %>%
            mutate(price_cluster = price_clusters,
                   density_cluster = density_clusters,
                   price_group = factor(price_order[price_cluster],
                   labels = c("Low price", "Medium price", "High price")),
                   density_group = factor(density_order[density_cluster],
                   labels = c("Low density", "Medium density", "High density")),
                   price_density_label = paste(price_group, density_group, sep = ", "))

plot_clusters <- ggplot(test_set, aes(x = sales_density, 
                                      y = avg_sales_price, 
                                      color = price_density_label, 
                                      shape = municipality_type)) +
                 geom_point(alpha = 0.6, size = 1.5) +
                 scale_color_brewer(palette = "Set1") +
                 scale_x_continuous(trans = "log1p") +
                 scale_y_continuous(trans = "log1p") +
                 scale_shape_discrete(name = "Municipality type") +  
                 labs(x = expression("Sales density [sales per km"^2*"]"),
                      y = "Average sales price [DKK]",
                      color = "Cluster") + 
                 theme_minimal()

saveRDS(plot_clusters, "saved_files/plot_clusters.rds")
print(plot_clusters)


#Plot clusters geographically
plot_clusters_geographically <- ggplot() +
                                # Add land/ocean contour
                                geom_contour(data = coordinate_grid,
                                             aes(x = coordinate_east, 
                                                 y = coordinate_north, 
                                                 z = as.numeric(!ocean)),
                                             color = "black", size = 0.4) +
                                # Add price-density group points
                                geom_point(data = test_set,
                                           aes(x = coordinate_east, 
                                               y = coordinate_north, 
                                               color = price_density_label),
                                           size = 0.5, alpha = 0.7) +
                                scale_color_brewer(palette = "Set1") +
                                coord_fixed() +
                                theme_void(base_size = 10) +
                                theme(legend.position = "none",
                                      plot.background = element_rect(fill = "white", color = NA),
                                      panel.background = element_rect(fill = "white", color = NA))

saveRDS(plot_clusters_geographically, "saved_files/plot_clusters_geographically.rds")
print(plot_clusters_geographically)



#Examine interplay between model performance and clusters
binned_data <- test_set %>%
               filter(!is.na(price_group), !is.na(density_group)) %>%
               group_by(price_group, density_group) %>%
               summarise(count = n(),
                         mean_abs_resid = mean(abs(resid_sales_price_log), na.rm = TRUE),
                         .groups = "drop")

plot_interplay <- ggplot(binned_data, aes(x = density_group, 
                                          y = price_group, 
                                          fill = mean_abs_resid)) +
                  geom_tile(color = "white") +
                  geom_text(aes(label = count), color = "black", size = 3) +
                  scale_fill_viridis_c(name = "MALE") +
                  labs(x = NULL,
                       y = NULL) + 
                  theme_minimal()

saveRDS(plot_interplay, "saved_files/plot_interplay.rds")
print(plot_interplay)

```

