---
title: "Datasets"
author: "August E. G. Mikkelsen"
date: "`r Sys.Date()`"
output: pdf_document
header-includes:
  - \usepackage{xcolor}
---

This notebook generates the dataset of Danish single family properties, property sales and lot sales used as input for the statistical analysis. 

```{r setup,  eval=TRUE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(stringr)
library(future)
library(purrr)
library(furrr)
library(tidyr)

unpack_element <- function(data, col, keep_empty = FALSE, ...) {
  
  data[[col]] <- lapply(data[[col]], function(x) {if (inherits(x, "list")) as.data.frame(x) else x})
  
  data %>% unnest(cols = all_of(col), keep_empty = keep_empty, ...)
  
}

```

## 1. Load raw datasets 
The final dataset is extracted from the raw datasets of properties, property sales and lot sales, which are generated by the script load_data.R. 

```{r Load datasets, eval=TRUE, echo=TRUE, warning=FALSE}
vurderingsejendomme <- readRDS("vurderingsejendomme.rds")
ejendomssalg <- readRDS("ejendomssalg.rds")
grundsalg <- readRDS("grundsalg.rds")
```


## 2. Single family houses 
The raw data contain a broad set of property types, ownership structures, buildings, units, and land parcels. This section restricts the data to a clean sample of single-family houses, single-family house sales, and single-family lot sales.

First, the data are restricted to simple property structures. For property sales and properties, only observations with exactly one valuation property, one land parcel, one unit, and one building are retained. For lot sales, only observations with exactly one valuation property and one land parcel are retained. This removes complex cases where the sale or property cannot be clearly interpreted as a single house or a single lot.

Second, bulk sales are removed. Sales with the same purchase date, transaction identifier, total purchase price, and municipality are only retained when they occur once. This reduces the risk of including transactions where several properties were sold together.

Third, the nested land-use information is unfolded. Each land parcel contains deljordstykker, and each deljordstykke contains planning information. These nested fields are expanded so that zoning and planning restrictions can be used directly in the filtering.

Finally, standard single-family house properties, standard single-family house sales, and standard single-family lot sales are selected. The filters require residential use, one single-family unit, no business area, standard residential facilities, one land parcel, no protected area, no summerhouse or unzoned area, and planning categories consistent with single-family residential use. 

```{r PCL, eval=TRUE, echo=TRUE, warning=FALSE, results = 'hide'}

#Only observations with exactly one valuation property, one land parcel, one unit, and one building
ejendomssalg <- ejendomssalg %>%
                add_count(vurderingsejendom_id_ice, df_virkningstid, name = "n_vurd") %>%
                add_count(delgrund.delgrund_ids, df_virkningstid, name = "n_delgrund") %>%
                add_count(enhed.enhed_id_ice, df_virkningstid, name = "n_enhed") %>%
                add_count(bygning.bygning_id_ice, df_virkningstid, name = "n_bygning") %>%
                filter(n_vurd == 1, n_delgrund == 1, n_enhed == 1, n_bygning == 1) %>%
                select(-n_vurd, -n_delgrund, -n_enhed, -n_bygning)

vurderingsejendomme <- vurderingsejendomme %>%
                       add_count(vurderingsejendom_id_ice, df_virkningstid, name = "n_vurd") %>%
                       add_count(delgrund.delgrund_ids, df_virkningstid, name = "n_delgrund") %>%
                       add_count(enhed.enhed_id_ice, df_virkningstid, name = "n_enhed") %>%
                       add_count(bygning.bygning_id_ice, df_virkningstid, name = "n_bygning") %>%
                       filter(n_vurd == 1, n_delgrund == 1, n_enhed == 1, n_bygning == 1) %>%
                       select(-n_vurd, -n_delgrund, -n_enhed, -n_bygning)

grundsalg <- grundsalg %>%
             add_count(vurderingsejendom_id_ice, df_virkningstid, name = "n_vurd") %>%
             add_count(delgrund.delgrund_ids, df_virkningstid, name = "n_delgrund") %>%
             filter(n_vurd == 1, n_delgrund == 1) %>%
             select(-n_vurd, -n_delgrund)


#Remove bulk sales
ejendomssalg <- ejendomssalg %>%
                group_by(salg.koebsdato, salg.anmeldelseidentifikator, salg.ialtkoebesum, vurinfo.kommunenummer) %>%
                filter(n() == 1) %>%
                ungroup()

grundsalg <- grundsalg %>%
             group_by(salg.koebsdato, salg.anmeldelseidentifikator, salg.ialtkoebesum, vurinfo.kommunenummer) %>%
             filter(n() == 1) %>%
             ungroup()


## Unfold information about each deljordstykke
ejendomssalg <- unpack_element(ejendomssalg, "delgrund.deljordstykke")

vurderingsejendomme <- unpack_element(vurderingsejendomme, "delgrund.deljordstykke")

grundsalg <- unpack_element(grundsalg, "delgrund.deljordstykke")


## Unfold planning information
ejendomssalg <- unpack_element(ejendomssalg, "deljordstykke.plandata", names_sep = "_")

vurderingsejendomme <- unpack_element(vurderingsejendomme, "deljordstykke.plandata", names_sep = "_")

grundsalg <- unpack_element(grundsalg, "deljordstykke.plandata", names_sep = "_")

#Standard single family houses
ejendomssalg_pcl <- ejendomssalg %>%
                    filter(bygning.bygningsanvendelse %in% c(120,121,122,130,131,132),
                           enhed.enhedsanvendelse %in% c(120,121,122,130,131,132),
                           salg.ejendomkategori == "EjendomKategoriEnfamiliesejendom",
                           !salg.flereejendommeindikator,
                           is.na(salg.bygningsforholdkode),
                           vurinfo.vurbenyttelseskode == "01", #Standard beboelse
                           vurinfo.antal_sfer == 1,
                           vurinfo.antal_bfer == 1,
                           is.na(bygning.ejerforholdskode), 
                           bygning.samlet_erhvervsareal == 0,
                           enhed.areal_til_erhverv == 0,
                           enhed.antal_vaerelser_til_erhverv == 0,
                           bygning.andetareal == 0,
                           enhed.boligtype == "1",
                           enhed.toiletforhold == "T",
                           enhed.badeforhold == "V",
                           enhed.koekkenforhold == "E",
                           delgrund.enheder_per_delgrund == 1,
                           delgrund.jordstykker_per_delgrund == 1,
                           delgrund.deljordstykker_per_delgrund == 1,
                           delgrund.beskyttelseslinier_andel == 0,
                           deljordstykke.areal_beskyttet == 0,
                           deljordstykke.andel_beskyttet == 0,
                           delgrund.samlet_sommerhuszoneareal == 0,
                           delgrund.samlet_udenzoneareal == 0,
                           deljordstykke.generelanvendelseskategori == 11,
                           deljordstykke.plandata_plandata.specifikanvendelseskategori %in% c("1100", "1110", "1120"))

vurderingsejendomme_pcl <- vurderingsejendomme %>%
                           filter(bygning.bygningsanvendelse %in% c(120,121,122,130,131,132),
                                  enhed.enhedsanvendelse %in% c(120,121,122,130,131,132),
                                  vurinfo.vurbenyttelseskode == "01", #Standard beboelse
                                  vurinfo.antal_sfer == 1,
                                  vurinfo.antal_bfer == 1,
                                  is.na(bygning.ejerforholdskode), 
                                  bygning.samlet_erhvervsareal == 0,
                                  enhed.areal_til_erhverv == 0,
                                  enhed.antal_vaerelser_til_erhverv == 0,
                                  bygning.andetareal == 0,
                                  enhed.boligtype == "1",
                                  enhed.toiletforhold == "T",
                                  enhed.badeforhold == "V",
                                  enhed.koekkenforhold == "E",
                                  delgrund.enheder_per_delgrund == 1,
                                  delgrund.jordstykker_per_delgrund == 1,
                                  delgrund.deljordstykker_per_delgrund == 1,
                                  delgrund.beskyttelseslinier_andel == 0,
                                  deljordstykke.areal_beskyttet == 0,
                                  deljordstykke.andel_beskyttet == 0,
                                  delgrund.samlet_sommerhuszoneareal == 0,
                                  delgrund.samlet_udenzoneareal == 0,
                                  deljordstykke.generelanvendelseskategori == 11,
                                  deljordstykke.plandata_plandata.specifikanvendelseskategori %in% c("1100", "1110", "1120"))

grundsalg_pcl <- grundsalg %>%
                 filter(salg.ejendomkategori == "EjendomKategoriUbebyggetgrund",
                        !salg.flereejendommeindikator,
                        is.na(salg.bygningsforholdkode),
                        vurinfo.vurbenyttelseskode == "09", #Ubebygget areal
                        vurinfo.antal_sfer == 1,
                        vurinfo.antal_bfer == 1,
                        delgrund.enheder_per_delgrund == 0,
                        delgrund.jordstykker_per_delgrund == 1,
                        delgrund.deljordstykker_per_delgrund == 1,
                        delgrund.beskyttelseslinier_andel == 0,
                        lengths(delgrund.alle_bygninger_paa_delgrunden) == 0,
                        deljordstykke.areal_beskyttet == 0,
                        deljordstykke.andel_beskyttet == 0,
                        delgrund.samlet_sommerhuszoneareal == 0,
                        delgrund.samlet_udenzoneareal == 0,
                        deljordstykke.generelanvendelseskategori == 11,
                        deljordstykke.plandata_plandata.specifikanvendelseskategori %in% c("1100", "1110", "1120"))

```

## 3. Data cleaning
This section applies the final cleaning steps used to construct the modelling samples.

First, observations with missing values in variables required for the subsequent models are removed. This includes coordinates, sales dates, prices, building characteristics, land area, zoning, view variables, and selected location variables.

Second, a set of consistency checks is imposed. Sales are only retained when the sale year is after the recorded construction and renovation years. The registered living area must match the building’s residential area, and the registered land area must be positive and consistent across the valuation property, land parcel, and sub-parcel records. This removes cases where area definitions differ across data sources or where road areas and parcel boundaries complicate the interpretation of the property.

Third, the sample is restricted to detached and semi-detached single-family houses and lots within reasonable value and size ranges. Property and lot sales with very low or very high prices are removed. Properties are restricted to land areas between 100 and 3,000 m², living areas between 50 and 400 m², construction year from 1900 onward, and urban or rural zoning. The categorical building and unit variables are restricted to the most common categories in the sales data, covering the dominant types of use, heating, roof material, wall material, number of rooms, bathrooms, and floors.

Fourth, sales with selected sales flags are excluded. These flags identify transactions that may not represent ordinary arm’s-length market sales or otherwise contain special circumstances that make them unsuitable for model training.

Finally, because the unfolding of planning data can create multiple rows per property, one planning category is selected for each property or sale. When several relevant planning categories are present, the priority order is 1110, then 1120, and finally 1100. The section ends with uniqueness checks to ensure that the resulting data sets contain one row per relevant property-sale, property, or lot-sale unit.


```{r Cleaning, eval=TRUE, echo=TRUE, warning=FALSE, results = 'hide'}

#Fjen NA'er for forskellige variable
ejendomssalg_pcl <- ejendomssalg_pcl %>%
                    filter(!is.na(vurinfo.adresse.koordinatoest),
                           !is.na(vurinfo.adresse.koordinatnord),
                           !is.na(salg.koebsdato),
                           !is.na(bygning.opfoerelsesaar),
                           !is.na(bygning.omtilbygningsaar),
                           !is.na(enhed.areal_til_beboelse),
                           !is.na(delgrund.registeret_areal_uden_vej),
                           !is.na(delgrund.zone),
                           !is.na(salg.ialtkoebesum),
                           !is.na(bygning.tagdaekningsmateriale),
                           !is.na(bygning.varmeinstallation),
                           !is.na(bygning.ydervaeggensmateriale),
                           !is.na(vurinfo.udsigtslaengde_hav),
                           !is.na(vurinfo.udsigtslaengde_soe),
                           !is.na(vurinfo.areal_samlet_skov))

vurderingsejendomme_pcl  <- vurderingsejendomme_pcl  %>%
                            filter(!is.na(vurinfo.adresse.koordinatoest),
                                   !is.na(vurinfo.adresse.koordinatnord),
                                   !is.na(bygning.opfoerelsesaar),
                                   !is.na(bygning.omtilbygningsaar),
                                   !is.na(enhed.areal_til_beboelse),
                                   !is.na(delgrund.registeret_areal_uden_vej),
                                   !is.na(delgrund.zone),
                                   !is.na(bygning.tagdaekningsmateriale),
                                   !is.na(bygning.varmeinstallation),
                                   !is.na(bygning.ydervaeggensmateriale),
                                   !is.na(vurinfo.udsigtslaengde_hav),
                                   !is.na(vurinfo.udsigtslaengde_soe),
                                   !is.na(vurinfo.areal_samlet_skov))

grundsalg_pcl <- grundsalg_pcl %>% filter(!is.na(vurinfo.adresse.koordinatoest),
                                          !is.na(vurinfo.adresse.koordinatnord),
                                          !is.na(salg.koebsdato),
                                          !is.na(vurinfo.registreretareal_fratrukket_vejareal),
                                          !is.na(delgrund.zone),
                                          !is.na(salg.ialtkoebesum),
                                          !is.na(vurinfo.udsigtslaengde_hav),
                                          !is.na(vurinfo.udsigtslaengde_soe),
                                          !is.na(vurinfo.areal_samlet_skov))


#Fjern forskellige abnormaliteter. Bla. ejendomme med vejarealer for at goere livet nemmere for os. 
ejendomssalg_pcl <- ejendomssalg_pcl %>%
                    filter(as.integer(format(as.Date(salg.koebsdato), "%Y")) >= bygning.opfoerelsesaar,
                           as.integer(format(as.Date(salg.koebsdato), "%Y")) >= bygning.omtilbygningsaar,
                           enhed.areal_til_beboelse == bygning.samlet_boligareal,
                           enhed.areal_til_beboelse > 0,
                           vurinfo.registreretareal_fratrukket_vejareal > 0,
                           delgrund.registeret_areal_uden_vej == vurinfo.registreretareal_fratrukket_vejareal,
                           delgrund.arealberegnet == vurinfo.registreretareal_fratrukket_vejareal,
                           deljordstykke.registreret_areal_uden_vej == vurinfo.registreretareal_fratrukket_vejareal,
                           deljordstykke.registreret_areal == vurinfo.registreretareal_fratrukket_vejareal,
                           deljordstykke.arealberegnetudenvej == vurinfo.registreretareal_fratrukket_vejareal,
                           deljordstykke.arealberegnet == vurinfo.registreretareal_fratrukket_vejareal) 

vurderingsejendomme_pcl <- vurderingsejendomme_pcl %>%
                           filter(enhed.areal_til_beboelse == bygning.samlet_boligareal,
                                  enhed.areal_til_beboelse > 0,
                                  vurinfo.registreretareal_fratrukket_vejareal > 0,
                                  delgrund.registeret_areal_uden_vej == vurinfo.registreretareal_fratrukket_vejareal,
                                  delgrund.arealberegnet == vurinfo.registreretareal_fratrukket_vejareal,
                                  deljordstykke.registreret_areal_uden_vej == vurinfo.registreretareal_fratrukket_vejareal,
                                  deljordstykke.registreret_areal == vurinfo.registreretareal_fratrukket_vejareal,
                                  deljordstykke.arealberegnetudenvej == vurinfo.registreretareal_fratrukket_vejareal,
                                  deljordstykke.arealberegnet == vurinfo.registreretareal_fratrukket_vejareal)

grundsalg_pcl <- grundsalg_pcl %>% 
                 filter(vurinfo.registreretareal_fratrukket_vejareal > 0,
                        delgrund.arealberegnet == vurinfo.registreretareal_fratrukket_vejareal,
                        deljordstykke.registreret_areal_uden_vej == vurinfo.registreretareal_fratrukket_vejareal,
                        deljordstykke.registreret_areal == vurinfo.registreretareal_fratrukket_vejareal,
                        deljordstykke.arealberegnetudenvej == vurinfo.registreretareal_fratrukket_vejareal,
                        deljordstykke.arealberegnet == vurinfo.registreretareal_fratrukket_vejareal)


#Udvælg ejendomme med standardegenskaber. For at gøre det simpelt udvælges kun de mest hyppige enhedstyper, tagtyper osv for salgene. 
round(prop.table(table(ejendomssalg_pcl$enhed.enhedsanvendelse)), 3) * 100 #120'ere og 130'ere udgør 98%, så disse vælges
round(prop.table(table(ejendomssalg_pcl$bygning.varmeinstallation)), 3) * 100 #1, 2, 5 og 7 er klart mest fremtrædende så disse vælges
round(prop.table(table(ejendomssalg_pcl$bygning.tagdaekningsmateriale)), 3) * 100 #1, 2, 3, 4 og 5 er mest fremtrædende så disse vælges
round(prop.table(table(ejendomssalg_pcl$bygning.ydervaeggensmateriale)), 3) * 100 #1, 2 og 5 er mest fremtrædende så disse vælges
round(prop.table(table(ejendomssalg_pcl$enhed.antal_vaerelser)), 3) * 100 #Vælg 2, 3, 4, 5, 6, 7 og 8 svarende til alle med en procentdel over 1
round(prop.table(table(ejendomssalg_pcl$enhed.antal_badevaerelser)), 3) * 100 #Vælg 1, 2, og 3 svarende til alle med en procentdel over 1
round(prop.table(table(ejendomssalg_pcl$bygning.antal_etager)), 3) * 100 #Vælg 1 og 2 svarende til alle med en procentdel over 1


ejendomssalg_pcl <- ejendomssalg_pcl %>%
                    filter(salg.ialtkoebesum >= 1e+5 & salg.ialtkoebesum <= 1.5e+7,
                           vurinfo.registreretareal_fratrukket_vejareal <= 3000,
                           vurinfo.registreretareal_fratrukket_vejareal >= 100,
                           enhed.areal_til_beboelse <= 400,
                           enhed.areal_til_beboelse >= 50,
                           bygning.opfoerelsesaar >= 1900,
                           delgrund.zone %in% c(1,2), #kun byzone og landzone
                           enhed.enhedsanvendelse %in% c(120,130),
                           bygning.bygningsanvendelse %in% c(120,130), 
                           bygning.varmeinstallation %in% c(1,2,5,7),
                           bygning.tagdaekningsmateriale %in% c(1,2,3,4,5),
                           bygning.ydervaeggensmateriale %in% c(1,2,5), 
                           enhed.antal_vaerelser %in% c(2,3,4,5,6,7,8,9,10),
                           enhed.antal_badevaerelser %in% c(1,2,3),
                           bygning.antal_etager %in% c(1,2))


vurderingsejendomme_pcl <- vurderingsejendomme_pcl %>%
                           filter(vurinfo.registreretareal_fratrukket_vejareal <= 3000,
                                  vurinfo.registreretareal_fratrukket_vejareal >= 100,
                                  enhed.areal_til_beboelse <= 400,
                                  enhed.areal_til_beboelse >= 50,
                                  bygning.opfoerelsesaar >= 1900,
                                  delgrund.zone %in% c(1,2), #kun byzone og landzone
                                  enhed.enhedsanvendelse %in% c(120,130),
                                  bygning.bygningsanvendelse %in% c(120,130), 
                                  bygning.varmeinstallation %in% c(1,2,5,7),
                                  bygning.tagdaekningsmateriale %in% c(1,2,3,4,5),
                                  bygning.ydervaeggensmateriale %in% c(1,2,5), 
                                  enhed.antal_vaerelser %in% c(2,3,4,5,6,7,8,9,10),
                                  enhed.antal_badevaerelser %in% c(1,2,3),
                                  bygning.antal_etager %in% c(1,2))


grundsalg_pcl <- grundsalg_pcl %>% filter(salg.ialtkoebesum >= 1e+5 & salg.ialtkoebesum <= 1.5e+7,
                                          vurinfo.registreretareal_fratrukket_vejareal <= 3000,
                                          vurinfo.registreretareal_fratrukket_vejareal >= 100,
                                          delgrund.zone %in% c(1,2)) #kun byzone og landzone
                              


#Fjern ejendomme med salgsflag 
ejendomssalg_pcl <- ejendomssalg_pcl %>% 
                    filter(!str_detect(salg_flag, "120|121|122|123|124|125|127|128|129|130|131|132|150|151|153|154|155|156|202|203|204"))

grundsalg_pcl <- grundsalg_pcl %>% 
                 filter(!str_detect(salg_flag, "120|121|122|123|124|125|127|128|129|130|131|132|150|151|153|154|155|156|202|203|204"))


#Der vælges specifikanvendelseskategori med prioritet i rækkefølgen 1110, 1120 og 1100 
ejendomssalg_pcl <- ejendomssalg_pcl %>%
                    group_by(vurderingsejendom_id_ice, salg.koebsdato) %>%
                    arrange(desc(match(deljordstykke.plandata_plandata.specifikanvendelseskategori, c("1100", "1120", "1110"))), 
                            .by_group = TRUE) %>%
                    slice(1) %>%
                    ungroup()

vurderingsejendomme_pcl <- vurderingsejendomme_pcl %>%
                           group_by(vurderingsejendom_id_ice) %>%
                           arrange(desc(match(deljordstykke.plandata_plandata.specifikanvendelseskategori, c("1100", "1120", "1110"))), 
                                   .by_group = TRUE) %>%
                           slice(1) %>%
                           ungroup()

grundsalg_pcl <- grundsalg_pcl %>%
                 group_by(vurderingsejendom_id_ice, salg.koebsdato) %>%
                 arrange(desc(match(deljordstykke.plandata_plandata.specifikanvendelseskategori, c("1100", "1120", "1110"))), 
                         .by_group = TRUE) %>%                 
                 slice(1) %>%
                 ungroup()


#Check at datasæt er unikke paa forskellige id-niveauer

##ejendomssalg
stopifnot(nrow(ejendomssalg_pcl %>% distinct(vurderingsejendom_id_ice, salg.koebsdato)) == nrow(ejendomssalg_pcl))
stopifnot(nrow(ejendomssalg_pcl %>% distinct(enhed.enhed_id_ice, salg.koebsdato)) == nrow(ejendomssalg_pcl))
stopifnot(nrow(ejendomssalg_pcl %>% distinct(delgrund.delgrund_ids, salg.koebsdato)) == nrow(ejendomssalg_pcl))
stopifnot(nrow(ejendomssalg_pcl %>% distinct(bygning.bygning_id_ice, salg.koebsdato)) == nrow(ejendomssalg_pcl))

##vurderingsejendomme
stopifnot(nrow(vurderingsejendomme_pcl %>% distinct(vurderingsejendom_id_ice)) == nrow(vurderingsejendomme_pcl))
stopifnot(nrow(vurderingsejendomme_pcl %>% distinct(enhed.enhed_id_ice)) == nrow(vurderingsejendomme_pcl))
stopifnot(nrow(vurderingsejendomme_pcl %>% distinct(delgrund.delgrund_ids)) == nrow(vurderingsejendomme_pcl))

##grundsalg
stopifnot(nrow(grundsalg_pcl %>% distinct(vurderingsejendom_id_ice, salg.koebsdato)) == nrow(grundsalg_pcl))
stopifnot(nrow(grundsalg_pcl %>% distinct(delgrund.delgrund_ids, salg.koebsdato)) == nrow(grundsalg_pcl))
stopifnot(nrow(grundsalg_pcl %>% distinct(deljordstykke.deljordstykke_id_ice, salg.koebsdato)) == nrow(grundsalg_pcl))

```

## 4. Final dataset

This section combines the cleaned property sales, valuation properties, and lot sales into one final analysis data set.

First, each observation is labelled by type: property sale, property, or lot sale. The three data sets are then combined into a single data frame. A common set of variables is selected and renamed to simpler English variable names used throughout the subsequent analysis.

The final data set contains sales information, location variables, building characteristics, land characteristics, planning variables, and the observation type. Internal property identifiers are also replaced by a compact numeric ID.

Finally, categorical variables are translated and recoded into interpretable categories. This includes zoning, municipality type, house type, roof material, heating type, and wall material. Renovation years coded as zero are set to missing values so that later imputation steps can treat them consistently.

The resulting file, dataset.rds, is the main input used in the analysis notebook.

```{r Dataset, eval=TRUE, echo=TRUE, warning=FALSE, results = 'hide'}
#join data frames of property sales and valuation properties
ejendomssalg_pcl <- ejendomssalg_pcl %>% mutate(type = "Property sale")
vurderingsejendomme_pcl <- vurderingsejendomme_pcl %>% mutate(type = "Property")
grundsalg_pcl <- grundsalg_pcl %>% mutate(type = "Lot sale")
dataset <- bind_rows(ejendomssalg_pcl, vurderingsejendomme_pcl, grundsalg_pcl)


#variables to select
variables <- c(id = "vurinfo.vurderingsejendom_id",
               sales_date = "salg.koebsdato",
               municipality = "kommune_navn",
               province = "landsdel_navn",
               region = "region_navn",
               municipality_type = "kommunegruppe",
               zone = "delgrund.zone",
               sales_price = "salg.ialtkoebesum",
               house_type = "enhed.enhedsanvendelse",
               living_space = "enhed.areal_til_beboelse",
               number_of_rooms = "enhed.antal_vaerelser",
               number_of_floors = "bygning.antal_etager",
               number_of_bathrooms = "enhed.antal_badevaerelser",
               lot_size = "vurinfo.registreretareal_fratrukket_vejareal",
               year_built = "bygning.opfoerelsesaar",
               year_rebuilt = "bygning.omtilbygningsaar",
               roof_type = "bygning.tagdaekningsmateriale",
               heating_type = "bygning.varmeinstallation",
               wall_type = "bygning.ydervaeggensmateriale",
               coordinate_east = "vurinfo.adresse.koordinatoest",
               coordinate_north = "vurinfo.adresse.koordinatnord",
               coast_distance = "vurinfo.afstand_kyst",
               lake_distance = "vurinfo.afstand_stor_soe",
               stream_distance = "vurinfo.afstand_stort_vandloeb",
               road_distance = "vurinfo.afstand_naermeste_trafikvejgennemfart_trafikfordeling",
               highway_distance = "vurinfo.afstand_motorvej_motortrafikvej",
               train_station_distance = "vurinfo.afstand_station_any",
               railway_distance = "vurinfo.afstand_jernbane_any",
               windturbine_distance = "vurinfo.afstand_vindmoelle_any",
               ocean_view = "vurinfo.udsigtslaengde_hav",
               lake_view = "vurinfo.udsigtslaengde_soe",
               forest_size = "vurinfo.areal_samlet_skov",
               maximum_coverage_percentage = "deljordstykke.plandata_plandata.maksbebygpct",
               subdivision_size = "deljordstykke.plandata_plandata.udstykstoerrelse",
               zoning_use_category = "deljordstykke.plandata_plandata.specifikanvendelseskategori",
               zoning_use_categories = "delgrund.unikke_plananvendelser",
               type = "type")

dataset <- dataset %>%
           rename(!!!setNames(variables, names(variables))) %>%
           select(names(variables))


#lav id om til standard range
id_map <- match(dataset$id, unique(dataset$id))
dataset$id <- id_map

#omdoeb zone (1=byzone,2=landzone)
dataset$zone[dataset$zone == 1] = "Urban"
dataset$zone[dataset$zone == 2] = "Rural"

#omdoeb municipality_type
dataset$municipality_type[dataset$municipality_type == "Hovedstadskommune"] = "Capital"
dataset$municipality_type[dataset$municipality_type == "Landkommune"] = "Rural"
dataset$municipality_type[dataset$municipality_type == "Oplandskommune"] = "Commuter"
dataset$municipality_type[dataset$municipality_type == "Provinsbykommune"] = "Provincial"
dataset$municipality_type[dataset$municipality_type == "Storbykommune"] = "Large city"

#omdoeb hustype
dataset$house_type[dataset$house_type == 120] = "Detached"
dataset$house_type[dataset$house_type == 130] = "Semi-detached"

#omdoeb tagtyper
dataset$roof_type[(dataset$roof_type %in% c(1,2))] = "Roofing felt"
dataset$roof_type[dataset$roof_type == 3] = "Fiber cement"
dataset$roof_type[dataset$roof_type == 4] = "Roofing tiles"
dataset$roof_type[dataset$roof_type == 5] = "Tiles"

#omdoeb varmekilder
dataset$heating_type[dataset$heating_type == 1] = "District heating"
dataset$heating_type[dataset$heating_type == 2] = "Central heating"
dataset$heating_type[dataset$heating_type == 5] = "Heat pump"
dataset$heating_type[dataset$heating_type == 7] = "Electric heating"

#omdoeb vaegtyper
dataset$wall_type[dataset$wall_type == 1] = "Brick"
dataset$wall_type[dataset$wall_type == 2] = "Lightweight concrete"
dataset$wall_type[dataset$wall_type == 5] = "Wood"


#lav ombygningsaar til NA (lidt fjollet, men for at alle imputationer virker på NA-værdier)
dataset$year_rebuilt[which(dataset$year_rebuilt==0)] = NA


#save file
saveRDS(dataset, "dataset.rds")

```

