---
title: "Supplementary Notebook"
author: "August E. G. Mikkelsen"
date: "2025-05-08"
output: pdf_document
header-includes:
  - \usepackage{xcolor}
---

```{r setup, eval=TRUE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggpubr)
library(mgcv)
library(gratia) 
library(knitr)
library(patchwork)
library(lubridate)
library(future)
library(future.apply)
library(rlang)
library(stringr)
library(furrr)
library(xgboost)
library(caret)
library(Hmisc)
library(kableExtra)
library(nabor)
library(purrr)
library(ggrastr)
library(ggridges)
library(gridExtra)
library(forcats)
library(pheatmap)
library(viridis)
library(cowplot)
library(scico)
library(scales)

# Global parameters
set.seed(123) # for reproducibility
N <- 4 # number of CV folds
k_spatial = 1000 # initial number of basis functions used in the spatial smooths
k_other = 10 # initial number of basis functions used for other smooths
load <- TRUE # determines whether or not to preload models in the various chunks
plan(multisession, workers = N) #do as much as possible in parallel
options(future.globals.maxSize = 10 * 1024^3) #increase maxSize for future to handle large models

#External files
coordinate_grid <- read.csv2("coordinates_ocean_land.csv")

#Create a saved files folder and checkpoints folder is these do not already exist
for (f in c("saved_files_lot_sales", "checkpoints_lot_sales")) {
  if (!dir.exists(f)) {
    dir.create(f, recursive = TRUE)
    message(sprintf("Created folder: %s", f))
  } else {
    message(sprintf("Folder already exists: %s", f))
  }
}
```

# Introduction
This notebook provides a walkthrough of the analysis and results described in the paper "Generalized Additive Models for Mass Appraisal: Evidence from the Danish Public Property Valuation System". 


# Dataset 

## 1. Description
The analysis is based on a dataset of single family home sales, single family lot sales and properties in Denmark from the period 01/01/2014-01/01/2024. The dataset, included as a .rds file in this repository, is loaded and inspected below.

```{r Load dataset, eval=TRUE, echo=TRUE, warning=FALSE, fig.width=20, fig.height=30}

dataset <- readRDS("dataset.rds")

table(dataset$type)

plot <- ggplot() +
        geom_contour(data = coordinate_grid,
                     aes(x = coordinate_east,
                         y = coordinate_north,
                         z = as.numeric(!ocean)),
                         color = "black",
                         size = 0.1) +
        geom_point(data = dataset,
                   aes(x = coordinate_east,
                       y = coordinate_north,
                       color = type),
                       size = 1.0) +
        coord_fixed() +
        theme_void(base_size = 10)


ggplot(dataset, aes(coordinate_east, coordinate_north)) +
  geom_contour(
    data = coordinate_grid,
    aes(
      x = coordinate_east,
      y = coordinate_north,
      z = as.numeric(!ocean)
    ),
    inherit.aes = FALSE,
    color = "black",
    linewidth = 0.1
  ) +
  geom_point(size = 0.2, alpha = 0.25, color = "steelblue") +
  coord_fixed() +
  facet_wrap(~ type) +
  theme_void(base_size = 10)

```

## 2. Variable conversions and imputation
The analysis relies on a set of variable conversions performed in order to:

- Handle missing values in the dataset
- Convert categorical variables to factors
- Convert sales dates to numeric values

The two first are straightforward while the treatment of missing values requires more consideration. The analysis presented in the paper is based on the following simply imputation scheme: 

- year_rebuilt: For this variable all NA's as well as values less than 1970 are replaced by 1970. 
- distance-variables: For these variables all NA's are replaced by the maximum value of the respective variable. 

The code for the above transformations is provided below. 

```{r Variable conversion and imputation, eval=TRUE, echo=TRUE, warning=FALSE, results= 'hide'}

#convert relevant categorical variables to factors with chosen references
dataset$municipality_type <- relevel(factor(dataset$municipality_type), ref = "Provincial")
dataset$house_type <- relevel(factor(dataset$house_type), ref = "Detached")
dataset$roof_type <- relevel(factor(dataset$roof_type), ref = "Fiber cement")
dataset$wall_type <- relevel(factor(dataset$wall_type), ref = "Brick")
dataset$heating_type <- relevel(factor(dataset$heating_type), ref = "District heating")

#create a numerical variable for sales_date
dataset <- dataset %>% mutate(sales_date = as.Date(sales_date),
                              sales_date_numeric = decimal_date(sales_date))

#impute NA-values of rebuilt year to 1970
dataset <- dataset %>%
  mutate(year_rebuilt = if_else(is.na(year_rebuilt) | year_rebuilt < 1970, 1970, year_rebuilt))

#impute NA-values of maximum_coverage_percentage to 30% 
dataset <- dataset %>% mutate(maximum_coverage_percentage = if_else(is.na(maximum_coverage_percentage), 30, maximum_coverage_percentage))

#impute NA-values of subdivision size to 700
dataset <- dataset %>% mutate(subdivision_size = if_else(is.na(subdivision_size), 700, subdivision_size))


#impute NA in distance variables to their cap values 
dataset <- dataset %>% mutate(across(c(lake_distance,
                                       stream_distance,
                                       coast_distance,
                                       road_distance,
                                       highway_distance,
                                       train_station_distance,
                                       railway_distance,
                                       windturbine_distance),
                                     ~ replace_na(., max(., na.rm = TRUE))))

#check that there are no remaining NAs in property sales data
which(is.na(subset(dataset, type == "Property sale")), arr.ind = TRUE)


```


## 3. Auxiliary variables
For the lot sales model the following auxiliary variables are calculated: 

- Gross floor area: Defined as the maximum allowed maximum coverage percentage x lot size
- Remaining floor area: Defined as lot size - gross floor area and bounded at 0. 
- Number of subdivisions: Defined as lot size / subdivision size rounded down to nearest integer and bounded at 1. 

```{r Auxiliary variables, eval=TRUE, echo=TRUE, warning=FALSE, results= 'hide'}

dataset <- dataset %>% mutate(gross_floor_area = lot_size * maximum_coverage_percentage/100,
                              remaining_floor_area = pmax(0, lot_size - gross_floor_area),
                              number_of_divisions = pmax(1, floor(lot_size / subdivision_size)))


```


## 5. Extract property types 
Extract different types of properties in the dataset for working with them separately. 


```{r Property types, eval=TRUE, echo=TRUE, warning=FALSE, results= 'hide'}
dataset_property_sales <- dataset %>% filter(type == "Property sale")
dataset_lot_sales <- dataset %>% filter(type == "Lot sale")
dataset_properties <- dataset %>% 
                      filter(type == "Property") %>% 
                      mutate(sales_date_numeric = 2024.0) #we want to value properties 1 January 2024


```

## 5. Cross-Validation
A simple 4-fold cross-validation split is implemented in the code below. 

```{r Cross-validation, eval=TRUE, echo=TRUE, warning=FALSE, results = 'hide'}

assign_cv_folds <- function(data, N, seed = 123) {
  set.seed(seed)
  data %>% mutate(validation_set = sample(rep(1:N, length.out = n())),
                  .rand = runif(n()))
}

#property sales
dataset_property_sales_cv <- assign_cv_folds(dataset_property_sales, N = N, seed = 123)

table(dataset_property_sales_cv$validation_set, useNA = "ifany")


#lot sales
dataset_lot_sales_cv <- assign_cv_folds(dataset_lot_sales, N = N, seed = 123)

table(dataset_lot_sales_cv$validation_set, useNA = "ifany")

```


# GAM Models 

## 1. Property model
The code below fits the GAM model structure described in "Generalized Additive Models for Real Estate Valuation: A Case Study of the Danish Housing Market" based on the cross validation split employed. 

```{r Model training, eval=TRUE, echo=TRUE, warning=FALSE, results='hide'}
fit_fold <- function(i, data, k_spatial, k_other = 10, fraction = 1.0,
                     checkpoint_prefix = "Model", save_checkpoint = TRUE,
                     seed = 123) {
  
  training_subset <- data %>% filter(validation_set != i)

  if (fraction < 1.0) {
    
    training_subset <- training_subset %>%
                       group_by(municipality_type) %>%
                       arrange(.rand, .by_group = TRUE) %>%
                       slice_head(prop = fraction) %>%
                       ungroup() 
  }
  
  model <- bam(sales_price ~ 
               s(coordinate_east, coordinate_north, k = k_spatial, pc = c(697000, 6150000)) +
               s(sales_date_numeric, k = k_other, by = municipality_type, pc = 2024) +
               s(year_built, k = k_other, by = municipality_type, pc = 1970) +
               s(year_rebuilt, k = k_other, by = municipality_type, pc = 1970) +
               s(living_space, k = k_other, by = municipality_type, pc = 140) +
               s(lot_size, k = k_other, by = municipality_type, pc = 800) +
               s(coast_distance, k = k_other, pc = 1500) +
               s(lake_distance, k = k_other, pc = 500) +
               s(stream_distance, k = k_other, pc = 1000) +
               s(road_distance, k = k_other, pc = 2000) +
               s(highway_distance, k = k_other, pc = 2000) +
               s(train_station_distance, k = k_other, pc = 2000) +
               s(railway_distance, k = k_other, pc = 1000) +
               s(windturbine_distance, k = k_other, pc = 3000) +
               s(ocean_view, k = k_other, pc = 0) +
               s(lake_view, k = k_other, pc = 0) +
               s(forest_size, k = k_other, pc = 0) +
               s(number_of_bathrooms, by = municipality_type, k = 2, pc = 1) +
               house_type * municipality_type +
               roof_type * municipality_type +
               wall_type * municipality_type +
               heating_type * municipality_type,
               data = training_subset,
               chunk.size = nrow(training_subset),
               family = tw(link = log),
               method = "fREML")
  
  if (save_checkpoint) {
    filename <- sprintf("%s_fold%d_k%d_frac%.2f.rds", checkpoint_prefix, i, k_spatial, fraction)
    saveRDS(model, file = file.path("checkpoints_lot_sales", filename))
  }
  
  model
}

fit_model_ensemble <- function(data, N, k_spatial, k_other = 10, fraction = 1.0, checkpoint_prefix = "Model") {
  
  models <- future_lapply(1:N, function(i) {
    fit_fold(
      i = i,
      data = data,
      k_spatial = k_spatial,
      k_other = k_other,
      fraction = fraction,
      checkpoint_prefix = checkpoint_prefix
    )
  })
  names(models) <- paste0("Model", 1:N)
  models
}

if (load) {
  models <- readRDS("saved_files_lot_sales/models.rds")
} else {
  models <- fit_model_ensemble(
    data = dataset_property_sales_cv,
    N = N,
    k_spatial = k_spatial,
    k_other = k_other,
    fraction = 1.0,
    checkpoint_prefix = "main"
  )
  saveRDS(models, "saved_files_lot_sales/models.rds")
}

```



## 5. Standardization
The lot sales and properties are augmented with a fictitious single family house corresponding to a set of standard building characteristics. Their standardized property value is subsequently calculated. 

```{r Standardization, eval=TRUE, echo=TRUE, warning=FALSE, results= 'hide'}

#augment lot sales
dataset_lot_sales_cv_standardized <- dataset_lot_sales_cv
dataset_lot_sales_cv_standardized$year_built <- 1970
dataset_lot_sales_cv_standardized$year_rebuilt <- 1970
dataset_lot_sales_cv_standardized$living_space <- 140
dataset_lot_sales_cv_standardized$lot_size <- 800
dataset_lot_sales_cv_standardized$number_of_bathrooms <- 1
dataset_lot_sales_cv_standardized$house_type <- factor("Detached", levels = levels(as.factor(dataset$house_type)))
dataset_lot_sales_cv_standardized$roof_type <- factor("Fiber cement", levels = levels(as.factor(dataset$roof_type)))
dataset_lot_sales_cv_standardized$wall_type <- factor("Brick", levels = levels(as.factor(dataset$wall_type)))
dataset_lot_sales_cv_standardized$heating_type <- factor("District heating", levels = levels(as.factor(dataset$heating_type)))

#augment properties
dataset_properties$year_built <- 1970
dataset_properties$year_rebuilt <- 1970
dataset_properties$living_space <- 140
dataset_properties$lot_size <- 800
dataset_properties$number_of_bathrooms <- 1
dataset_properties$house_type <- factor("Detached", levels = levels(as.factor(dataset$house_type)))
dataset_properties$roof_type <- factor("Fiber cement", levels = levels(as.factor(dataset$roof_type)))
dataset_properties$wall_type <- factor("Brick", levels = levels(as.factor(dataset$wall_type)))
dataset_properties$heating_type <- factor("District heating", levels = levels(as.factor(dataset$heating_type)))

#calculate standarized prices for each model 

for (model_name in names(models)) {
  
  dataset_lot_sales_cv[[paste0("sales_price_standardized_", model_name)]] <- as.vector(predict.gam(models[[model_name]], dataset_lot_sales_cv_standardized, type = "response"))
  dataset_properties[[paste0("sales_price_standardized_", model_name)]] <- as.vector(predict.gam(models[[model_name]], dataset_properties, type = "response"))

}



```


## 6. Lot models

```{r Lot models, eval=TRUE, echo=TRUE, warning=FALSE, results= 'hide'}

fit_fold_model <- function(formula, data, fold) {
  
  training_subset <- data %>% filter(validation_set != fold)
  
  bam(formula,
      data = training_subset,
      family = tw(link = log),
      method = "fREML")
}

fit_model_set <- function(formula, data, N) {
  
  models <- future_lapply(seq_len(N), function(i) {
    fit_fold_model(
      formula = formula,
      data = data,
      fold = i
    )
  })
  
  names(models) <- paste0("Model", seq_len(N))
  
  models
}


standard_vars <- paste0("sales_price_standardized_Model", 1:N)

base_terms <- list(none = NULL,
                   lot_size = "s(lot_size, k = 6, pc = 800)",
                   divisions = "s(number_of_divisions, k = 3, pc = 1)",
                   coverage = "s(maximum_coverage_percentage, k = 3, pc = 30)",
                   lot_size_divisions = c("s(lot_size, k = 6, pc = 800)", "s(number_of_divisions, k = 3, pc = 1)"),
                   lot_size_coverage = c("s(lot_size, k = 6, pc = 800)", "s(maximum_coverage_percentage, k = 3, pc = 30)"),
                   lot_size_divisions_coverage = c("s(lot_size, k = 6, pc = 800)","s(number_of_divisions, k = 3, pc = 1)", "s(maximum_coverage_percentage, k = 3, pc = 30)"),
                   te_lot_coverage = "te(lot_size, maximum_coverage_percentage, k = c(6, 3))",
                   te_lot_divisions = "te(lot_size, number_of_divisions, k = c(6, 3))",
                   te_lot_coverage = "te(lot_size, maximum_coverage_percentage, number_of_divisions, k = c(6, 3, 3))",
                   gross_floor_area = "s(gross_floor_area, k = 6, pc = 240)",
                   gross_remaining_area = c("s(gross_floor_area, k = 6, pc = 240)", "s(remaining_floor_area, k = 6, pc = 560)"),
                   te_gross_remaining = "te(gross_floor_area, remaining_floor_area, k = c(6, 6))")

coord_terms <- list(no_coords = NULL, coords = "s(coordinate_east, coordinate_north, k = 10)")

make_formula <- function(std_var, extra_terms = NULL, coord_term = NULL) {
  
  rhs <- c(paste0("s(", std_var, ", k = 10)"), extra_terms, coord_term)
  
  as.formula(paste("sales_price ~", paste(rhs, collapse = " + ")))
  
}

model_grid <- expand.grid(std_var = standard_vars,
                          base_name = names(base_terms),
                          coord_name = names(coord_terms),
                          stringsAsFactors = FALSE)

model_formulas <- pmap(
  model_grid,
  function(std_var, base_name, coord_name) {
    make_formula(
      std_var = std_var,
      extra_terms = base_terms[[base_name]],
      coord_term = coord_terms[[coord_name]]
    )
  }
)

names(model_formulas) <- paste0(
  "model_",
  seq_along(model_formulas),
  "_",
  model_grid$std_var,
  "_",
  model_grid$base_name,
  "_",
  model_grid$coord_name
)
models <- lapply(names(model_formulas), function(model_name) {
  message("Fitting ", model_name, " ...")
  fit_model_set(
    formula = model_formulas[[model_name]],
    data = dataset_lot_sales_cv,
    N = N
  )
})

names(models) <- names(model_formulas)

saveRDS(models, "~/vurdst-avm-extension/saved_files_lot_sales/models_lot_sales.rds")


```




#### 7. Smooth effect of standard prices 

```{r Smooth effects, eval=TRUE, echo=TRUE, warning=FALSE, results = 'hide', fig.width=20, fig.height=30}

extract_std_var <- function(model_name) {str_extract(model_name, "sales_price_standardized_Model[1-4]")}

extract_standard_price_smooths <- function(models, x_min = 0, x_max = 1e7, n = 200) {
  
    imap_dfr(models, function(model_set, model_name) {
    
      std_var <- str_extract(model_name, "sales_price_standardized_Model[1-4]")
    
      imap_dfr(model_set, function(mod, lot_split_name) {
      
      new_data <- tibble(!!std_var := seq(x_min, x_max, length.out = n),
                         lot_size = 800,
                         number_of_divisions = 1,
                         maximum_coverage_percentage = 30,
                         gross_floor_area = 240,
                         remaining_floor_area = 560,
                         coordinate_east = 697000,
                         coordinate_north = 6150000)
      
      df <- tryCatch(smooth_estimates(
                     mod,
                     smooth = paste0("s(", std_var, ")"),
                     data = new_data,
                     partial_match = TRUE),
            error = function(e) {
            message(sprintf("Skipping %s / %s: %s", model_name, lot_split_name, e$message))
            return(NULL)})
      
      if (is.null(df)) return(tibble::tibble())
      
      intercept <- coef(mod)[["(Intercept)"]]
      
      df %>% mutate(model_name = model_name,
                    property_model = stringr::str_extract(std_var, "Model[1-4]"),
                    lot_split = sub("^Model", "Lot split ", lot_split_name),
                    standardized_price = .data[[std_var]],
                    partial_effect = .estimate,
                    price_effect = exp(intercept + .estimate),
                    price_lower = exp(intercept + .estimate - 1.96 * .se),
                    price_upper = exp(intercept + .estimate + 1.96 * .se))
    })
  })
}

combined_df <- extract_standard_price_smooths(models)


#Plot on link scale


plot_standardized_smooth <- ggplot(combined_df,
                                   aes(x = standardized_price,
                                       y = .estimate,
                                       color = lot_split,
                                       linetype = property_model,
                                       group = interaction(model_name, lot_split))) +
                            geom_ribbon(data = filter(combined_df, property_model == "Model1", lot_split == "Lot split 1"),
                                        aes(x = standardized_price,
                                            ymin = .estimate - 1.96 * .se,
                                            ymax = .estimate + 1.96 * .se),
                                        inherit.aes = FALSE,
                                        alpha = 0.25,
                                        fill = "grey60") +
                            geom_line(linewidth = 0.8, alpha = 0.8) +
                            scale_x_continuous(limits = c(0, 1e7),
                                               labels = function(x) paste0(x / 1e6, "M")) +
                            labs(x = "Standard property value [kr.]",
                                 y = "Partial effect",
                                 color = "Lot split",
                                 linetype = "Property split") +
                            theme_minimal()

plot_standardized_smooth


# response scale

plot_standardized_smooth_response <- ggplot(combined_df,
                                            aes(x = standardized_price,
                                                y = price_effect,
                                                color = lot_split,
                                                linetype = property_model,
                                                group = interaction(model_name, lot_split))) +
                                     geom_abline(slope = 1,
                                                 intercept = 0,
                                                 color = "black",
                                                 linetype = "dashed",
                                                 linewidth = 0.8) +
                                     geom_ribbon(data = filter(combined_df, property_model == "Model1", lot_split == "Lot split 1"),
                                                 aes(x = standardized_price,
                                                     ymin = price_lower,
                                                     ymax = price_upper),
                                                 inherit.aes = FALSE,
                                                 alpha = 0.25,
                                                 fill = "grey60") +
                                     geom_line(linewidth = 0.8, alpha = 0.8) +
                                     scale_x_continuous(limits = c(0, 0.75*1e7),
                                                        labels = function(x) paste0(x / 1e6, "M")) +
                                     scale_y_continuous(limits = c(0, 0.75*1e7),
                                                        labels = function(x) paste0(round(x / 1e6, 1), "M")) +
                                     labs(x = "Standard property value [kr.]",
                                          y = "Standard lot value [kr.]",
                                          color = "Lot split",
                                          linetype = "Property split") +
                                     theme_minimal()

plot_standardized_smooth_response

```

#### 7. Other 1D smooth effects 

```{r Other 1D smooth effects, eval=TRUE, echo=TRUE, warning=FALSE, results = 'hide', fig.width=20, fig.height=30}

reference_values <- list(sales_price_standardized_Model1 = 5e6,
                         sales_price_standardized_Model2 = 5e6,
                         sales_price_standardized_Model3 = 5e6,
                         sales_price_standardized_Model4 = 5e6,
                         lot_size = 800,
                         number_of_divisions = 1,
                         maximum_coverage_percentage = 30,
                         gross_floor_area = 240,
                         remaining_floor_area = 560,
                         coordinate_east = 697000,
                         coordinate_north = 6150000)

smooth_ranges <- list(lot_size = c(0, 3000),
                      number_of_divisions = c(1, 6),
                      maximum_coverage_percentage = c(0, 100),
                      gross_floor_area = c(0, 1500),
                      remaining_floor_area = c(0, 3000))


extract_1d_smooths <- function(models, smooth_vars, reference_values, smooth_ranges, n = 200) {
  
  imap_dfr(models, function(model_set, model_name) {
    
    std_var <- str_extract(model_name, "sales_price_standardized_Model[1-4]")
    
    imap_dfr(model_set, function(mod, lot_split_name) {
      
      available_smooths <- vapply(mod$smooth, `[[`, character(1), "label")
      
      map_dfr(smooth_vars, function(smooth_var) {
        
        smooth_label <- paste0("s(", smooth_var, ")")
        
        if (!smooth_label %in% available_smooths) {
          return(tibble())
        }
        
        x_seq <- seq(
          smooth_ranges[[smooth_var]][1],
          smooth_ranges[[smooth_var]][2],
          length.out = n
        )
        
        new_data <- as_tibble(reference_values)
        new_data <- new_data[rep(1, n), ]
        new_data[[std_var]] <- reference_values[[std_var]]
        new_data[[smooth_var]] <- x_seq
        
        df <- tryCatch(
          smooth_estimates(
            mod,
            smooth = smooth_label,
            data = new_data
          ),
          error = function(e) {
            message(sprintf(
              "Skipping %s / %s / %s: %s",
              model_name, lot_split_name, smooth_var, e$message
            ))
            return(NULL)
          }
        )
        
        if (is.null(df)) return(tibble())
        

        df %>%
          mutate(
            model_name = model_name,
            property_model = str_extract(std_var, "Model[1-4]"),
            lot_split = sub("^Model", "Lot split ", lot_split_name),
            smooth_var = smooth_var,
            x_value = .data[[smooth_var]],
            partial_effect = .estimate)
      })
    })
  })
}


smooth_vars_1d <- c("lot_size",
                    "number_of_divisions",
                    "maximum_coverage_percentage",
                    "gross_floor_area",
                    "remaining_floor_area")

combined_1d_df <- extract_1d_smooths(models = models,
                                     smooth_vars = smooth_vars_1d,
                                     reference_values = reference_values,
                                     smooth_ranges = smooth_ranges,
                                     n = 200)


plot_1d_smooths <- ggplot(combined_1d_df,
                          aes(x = x_value,
                              y = partial_effect,
                              color = lot_split,
                              linetype = property_model,
                              group = interaction(model_name, lot_split))) +
                          geom_ribbon(data = filter(combined_1d_df, property_model == "Model1", lot_split == "Lot split 1"),
                          aes(x = x_value,
                              ymin = partial_effect - 1.96 * .se,
                              ymax = partial_effect + 1.96 * .se),
                          inherit.aes = FALSE,
                          alpha = 0.25,
                          fill = "grey60") +
                          geom_line(linewidth = 0.8, alpha = 0.8) +
                          facet_wrap(~ smooth_var, scales = "free_x") +
                          labs(x = NULL,
                               y = "Partial effect",
                               color = "Lot split",
                               linetype = "Property split") +
                          theme_minimal()

plot_1d_smooths
```




#### 9. Performance

```{r Performance, eval=TRUE, echo=TRUE, warning=FALSE, results = 'hide', fig.width=20, fig.height=30}


get_residuals <- function(model_set, dataset) {
  
  residual_list <- future_lapply(names(model_set), function(fold_model_name) {
    
    model <- model_set[[fold_model_name]]
    fold <- as.integer(str_extract(fold_model_name, "\\d+"))
    
    predicted_sales_price <- as.vector(
      predict.gam(
        model,
        newdata = dataset,
        type = "response"
      )
    )
    
    sales_price <- dataset$sales_price
    
    tibble(
      sales_price = sales_price,
      sales_date_numeric = dataset$sales_date_numeric,
      predicted_sales_price = predicted_sales_price,
      residual_sales_price = predicted_sales_price - sales_price,
      residual_sales_price_percent = (predicted_sales_price - sales_price) / sales_price * 100,
      residual_sales_price_log = log(predicted_sales_price) - log(sales_price),
      ratio = predicted_sales_price / sales_price,
      validation = dataset$validation_set == fold,
      id = dataset$id,
      municipality_type = dataset$municipality_type,
      fold_model = fold_model_name
    )
  })
  
  bind_rows(residual_list)
}

summarise_performance <- function(model_set, dataset, label) {
  
  residual_df <- get_residuals(model_set, dataset)
  
  residual_df %>%
    mutate(
      dataset_type = if_else(validation, "Validation", "Training"),
      Model = label
    ) %>%
    group_by(Model, dataset_type) %>%
    summarise(
      municipality_type = "Countrywide",
      MALE = mean(abs(residual_sales_price_log), na.rm = TRUE),
      MAPE = mean(abs(residual_sales_price_percent), na.rm = TRUE),
      MAE = mean(abs(residual_sales_price), na.rm = TRUE),
      PM20 = mean(abs(residual_sales_price_percent) <= 20, na.rm = TRUE) * 100,
      COD = mean(abs(ratio - median(ratio, na.rm = TRUE)), na.rm = TRUE) /
        median(ratio, na.rm = TRUE) * 100,
      PRD = mean(ratio, na.rm = TRUE) /
        (sum(predicted_sales_price, na.rm = TRUE) / sum(sales_price, na.rm = TRUE)),
      .groups = "drop"
    )
}


performance_countrywide <- bind_rows(
  lapply(names(models), function(label) {
    summarise_performance(
      model_set = models[[label]],
      dataset = dataset_lot_sales_cv,
      label = label
    )
  })
)


saveRDS(performance_countrywide, "saved_files_lot_sales/performance_countrywide.rds")
```
