---
title: "Supplementary Notebook"
author: "August E. G. Mikkelsen"
date: "2025-05-08"
output: pdf_document
header-includes:
  - \usepackage{xcolor}
---

```{r setup, eval=TRUE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggpubr)
library(mgcv)
library(gratia) 
library(knitr)
library(patchwork)
library(lubridate)
library(future)
library(future.apply)
library(rlang)
library(stringr)
library(furrr)
library(xgboost)
library(caret)
library(Hmisc)
library(kableExtra)
library(nabor)
library(purrr)
library(ggrastr)
library(ggridges)
library(gridExtra)
library(forcats)
library(pheatmap)
library(viridis)
library(cowplot)
library(scico)
library(scales)

# Global parameters
set.seed(123) # for reproducibility
N <- 4 # number of CV folds
k_spatial = 500 # initial number of basis functions used in the spatial smooths
k_other = 10 # initial number of basis functions used for other smooths
load <- FALSE # determines whether or not to preload models in the various chunks
plan(multisession, workers = N) #do as much as possible in parallel
options(future.globals.maxSize = 10 * 1024^3) #increase maxSize for future to handle large models

#External files
coordinate_grid <- read.csv2("coordinates_ocean_land.csv")

#Create a saved files folder and checkpoints folder is these do not already exist
for (f in c("saved_files", "checkpoints")) {
  if (!dir.exists(f)) {
    dir.create(f, recursive = TRUE)
    message(sprintf("Created folder: %s", f))
  } else {
    message(sprintf("Folder already exists: %s", f))
  }
}
```

# Introduction
This notebook provides a walkthrough of the analysis and results described in the paper "Generalized Additive Models for Mass Appraisal: Evidence from the Danish Public Property Valuation System". 


# Dataset 

## 1. Description
The analysis is based on a dataset of single family home sales, single family lot sales and properties in Denmark from the period 01/01/2014-01/01/2024. The dataset, included as a .rds file in this repository, is loaded and inspected below.

```{r Load dataset, eval=TRUE, echo=TRUE, warning=FALSE, fig.width=20, fig.height=30}

dataset <- readRDS("dataset.rds")
str(dataset)

```

## 2. Variable conversions and imputation
The analysis relies on a set of variable conversions performed in order to:

- Handle missing values in the dataset
- Convert categorical variables to factors
- Convert sales dates to numeric values

The two first are straightforward while the treatment of missing values requires more consideration. The analysis presented in the paper is based on the following simply imputation scheme: 

- year_rebuilt: For this variable all NA's as well as values less than 1970 are replaced by 1970. 
- distance-variables: For these variables all NA's are replaced by the maximum value of the respective variable. 

The code for the above transformations is provided below. 

```{r Variable conversion and imputation, eval=TRUE, echo=TRUE, warning=FALSE, results= 'hide'}

#convert relevant categorical variables to factors with chosen references
dataset$municipality_type <- relevel(factor(dataset$municipality_type), ref = "Provincial")
dataset$house_type <- relevel(factor(dataset$house_type), ref = "Detached")
dataset$roof_type <- relevel(factor(dataset$roof_type), ref = "Fiber cement")
dataset$wall_type <- relevel(factor(dataset$wall_type), ref = "Brick")
dataset$heating_type <- relevel(factor(dataset$heating_type), ref = "District heating")

#create a numerical variable for sales_date
dataset <- dataset %>% mutate(sales_date = as.Date(sales_date),
                              sales_date_numeric = decimal_date(sales_date))

#impute NA-values
dataset <- dataset %>%
  mutate(year_rebuilt = if_else(is.na(year_rebuilt) | year_rebuilt < 1970, 1970, year_rebuilt))

dataset <- dataset %>% mutate(across(c(lake_distance,
                                       stream_distance,
                                       coast_distance,
                                       road_distance,
                                       highway_distance,
                                       train_station_distance,
                                       railway_distance,
                                       windturbine_distance),
                                     ~ replace_na(., max(., na.rm = TRUE))))

#check that there are no remaining NAs in sales data
which(is.na(subset(dataset, type == "Property sale")), arr.ind = TRUE)
which(is.na(subset(dataset, type == "Lot sale")), arr.ind = TRUE)

```

## 3. Overview of property types
A simple 4-fold cross-validation split is implemented in the code below. 

```{r Overview of property types, eval=TRUE, echo=TRUE, warning=FALSE, results = 'hide'}
#The three types included in the dataset. This notebook only uses Property and Property Sales
table(dataset$type)



```


## 4. Cross-Validation
A simple 4-fold cross-validation split is implemented in the code below. 

```{r Cross-validation, eval=TRUE, echo=TRUE, warning=FALSE, results = 'hide'}
# pick out sales and properties
dataset_sales <- dataset %>% filter(type == "Property sale")
dataset_properties <- dataset %>% 
                      filter(type == "Property") %>% 
                      mutate(sales_date_numeric = 2024.0) #we want to value properties 1 January 2024

assign_cv_folds <- function(data, N, seed = 123) {
  set.seed(seed)
  data %>% mutate(validation_set = sample(rep(1:N, length.out = n())),
                  .rand = runif(n()))
}

dataset_sales_cv <- assign_cv_folds(dataset_sales, N = N, seed = 123)

table(dataset_sales_cv$validation_set, useNA = "ifany")

```


# GAM Models 

## 1. Model training
The code below fits the GAM model structure described in "Generalized Additive Models for Real Estate Valuation: A Case Study of the Danish Housing Market" based on the cross validation split employed. 

```{r Model training, eval=TRUE, echo=TRUE, warning=FALSE, results='hide'}
fit_fold <- function(i, data, k_spatial, k_other = 10, fraction = 1.0,
                     checkpoint_prefix = "Model", save_checkpoint = TRUE,
                     seed = 123) {
  
  training_subset <- data %>% filter(validation_set != i)

  if (fraction < 1.0) {
    
    training_subset <- training_subset %>%
                       group_by(municipality_type) %>%
                       arrange(.rand, .by_group = TRUE) %>%
                       slice_head(prop = fraction) %>%
                       ungroup() 
  }
  
  model <- bam(sales_price ~ 
               s(coordinate_east, coordinate_north, k = k_spatial, pc = c(697000, 6150000)) +
               s(sales_date_numeric, k = k_other, by = municipality_type, pc = 2024) +
               s(year_built, k = k_other, by = municipality_type, pc = 1970) +
               s(year_rebuilt, k = k_other, by = municipality_type, pc = 1970) +
               s(living_space, k = k_other, by = municipality_type, pc = 140) +
               s(lot_size, k = k_other, by = municipality_type, pc = 800) +
               s(coast_distance, k = k_other, pc = 1500) +
               s(lake_distance, k = k_other, pc = 500) +
               s(stream_distance, k = k_other, pc = 1000) +
               s(road_distance, k = k_other, pc = 2000) +
               s(highway_distance, k = k_other, pc = 2000) +
               s(train_station_distance, k = k_other, pc = 2000) +
               s(railway_distance, k = k_other, pc = 1000) +
               s(windturbine_distance, k = k_other, pc = 3000) +
               s(ocean_view, k = k_other, pc = 0) +
               s(lake_view, k = k_other, pc = 0) +
               s(forest_size, k = k_other, pc = 0) +
               s(number_of_bathrooms, by = municipality_type, k = 2, pc = 1) +
               house_type * municipality_type +
               roof_type * municipality_type +
               wall_type * municipality_type +
               heating_type * municipality_type,
               data = training_subset,
               chunk.size = nrow(training_subset),
               family = tw(link = log),
               method = "fREML")
  
  if (save_checkpoint) {
    filename <- sprintf("%s_fold%d_k%d_frac%.2f.rds", checkpoint_prefix, i, k_spatial, fraction)
    saveRDS(model, file = file.path("checkpoints", filename))
  }
  
  model
}

fit_model_ensemble <- function(data, N, k_spatial, k_other = 10, fraction = 1.0, checkpoint_prefix = "Model") {
  
  models <- future_lapply(1:N, function(i) {
    fit_fold(
      i = i,
      data = data,
      k_spatial = k_spatial,
      k_other = k_other,
      fraction = fraction,
      checkpoint_prefix = checkpoint_prefix
    )
  })
  names(models) <- paste0("Model", 1:N)
  models
}

if (load) {
  models <- readRDS("saved_files/models.rds")
} else {
  models <- fit_model_ensemble(
    data = dataset_sales_cv,
    N = N,
    k_spatial = k_spatial,
    k_other = k_other,
    fraction = 1.0,
    checkpoint_prefix = "main"
  )
  saveRDS(models, "saved_files/models.rds")
}

```

