---
title: "Supplementary Notebook"
author: "August E. G. Mikkelsen"
date: "2025-05-08"
output: pdf_document
header-includes:
  - \usepackage{xcolor}
---

```{r setup, eval=TRUE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggpubr)
library(mgcv)
library(gratia) 
library(knitr)
library(patchwork)
library(lubridate)
library(future)
library(future.apply)
library(rlang)
library(stringr)
library(furrr)
library(xgboost)
library(caret)
library(Hmisc)
library(kableExtra)
library(nabor)
library(purrr)
library(ggrastr)
library(ggridges)
library(gridExtra)
library(forcats)
library(pheatmap)
library(viridis)
library(cowplot)
library(scico)
library(scales)

# Global parameters
set.seed(123) # for reproducibility
N <- 4 # number of CV folds
k_spatial = 2000 # initial number of basis functions used in the spatial smooths
k_other = 10 # initial number of basis functions used for other smooths
load <- FALSE # determines whether or not to preload models in the various chunks
plan(multisession, workers = N) #do as much as possible in parallel
options(future.globals.maxSize = 10 * 1024^3) #increase maxSize for future to handle large models

#External files
coordinate_grid <- read.csv2("coordinates_ocean_land.csv")

#Create a saved files folder and checkpoints folder is these do not already exist
for (f in c("saved_files_lot_sales", "checkpoints_lot_sales")) {
  if (!dir.exists(f)) {
    dir.create(f, recursive = TRUE)
    message(sprintf("Created folder: %s", f))
  } else {
    message(sprintf("Folder already exists: %s", f))
  }
}
```

# Introduction
This notebook provides a walkthrough of the analysis and results described in the paper "Generalized Additive Models for Mass Appraisal: Evidence from the Danish Public Property Valuation System". 


# Dataset 

## 1. Description
The analysis is based on a dataset of single family home sales, single family lot sales and properties in Denmark from the period 01/01/2014-01/01/2024. The dataset, included as a .rds file in this repository, is loaded and inspected below.

```{r Load dataset, eval=TRUE, echo=TRUE, warning=FALSE, fig.width=20, fig.height=30}

dataset <- readRDS("dataset.rds")

table(dataset$type)

#Distribution of sales prices
plot_sales_price_distribution <- ggplot(dataset %>% filter(type %in% c("Lot sale", "Property sale")),
                                 aes(x = sales_price / 1e6,
                                     fill = type,
                                     color = type)) +
                                 geom_density(linewidth = 1,
                                              alpha = 0.3,
                                              adjust = 1.5,
                                              na.rm = TRUE) +
                                 coord_cartesian(xlim = c(0, 15)) +
                                 labs(title = "Sales price distribution",
                                      x = "Sales price [million DKK]",
                                      y = "Density",
                                      fill = NULL,
                                      color = NULL) +
                                 theme(legend.position = c(0.8, 0.8)) + 
                                 theme_minimal()

plot_sales_price_distribution

#Geographic distribution of properties
plot <- ggplot() +
        geom_contour(data = coordinate_grid,
                     aes(x = coordinate_east,
                         y = coordinate_north,
                         z = as.numeric(!ocean)),
                         color = "black",
                         size = 0.1) +
        geom_point(data = dataset,
                   aes(x = coordinate_east,
                       y = coordinate_north,
                       color = type),
                       size = 1.0) +
        coord_fixed() +
        theme_void(base_size = 10)


plot_geographic <- ggplot(dataset, aes(coordinate_east, coordinate_north)) +
                   geom_contour(data = coordinate_grid,
                                aes(x = coordinate_east,
                                    y = coordinate_north,
                                    z = as.numeric(!ocean)),
                                inherit.aes = FALSE,
                                color = "black",
                                linewidth = 0.1) +
                   geom_point(size = 0.2, alpha = 0.25, color = "steelblue") +
                   coord_fixed() +
                   facet_wrap(~ type) +
                   theme_void(base_size = 10)

plot_geographic

```

## 2. Variable conversions and imputation
The analysis relies on a set of variable conversions performed in order to:

- Handle missing values in the dataset
- Convert categorical variables to factors
- Convert sales dates to numeric values

The two first are straightforward while the treatment of missing values requires more consideration. The analysis presented in the paper is based on the following simply imputation scheme: 

- year_rebuilt: For this variable all NA's as well as values less than 1970 are replaced by 1970. 
- distance-variables: For these variables all NA's are replaced by the maximum value of the respective variable. 

The code for the above transformations is provided below. 

```{r Variable conversion and imputation, eval=TRUE, echo=TRUE, warning=FALSE, results= 'hide'}

#convert relevant categorical variables to factors with chosen references
dataset$municipality_type <- relevel(factor(dataset$municipality_type), ref = "Provincial")
dataset$house_type <- relevel(factor(dataset$house_type), ref = "Detached")
dataset$roof_type <- relevel(factor(dataset$roof_type), ref = "Fiber cement")
dataset$wall_type <- relevel(factor(dataset$wall_type), ref = "Brick")
dataset$heating_type <- relevel(factor(dataset$heating_type), ref = "District heating")

#create a numerical variable for sales_date
dataset <- dataset %>% mutate(sales_date = as.Date(sales_date),
                              sales_date_numeric = decimal_date(sales_date))

#impute NA-values of rebuilt year to 1970
dataset <- dataset %>%
  mutate(year_rebuilt = if_else(is.na(year_rebuilt) | year_rebuilt < 1970, 1970, year_rebuilt))

#impute NA-values of maximum_coverage_percentage to 30% 
dataset <- dataset %>% mutate(maximum_coverage_percentage = if_else(is.na(maximum_coverage_percentage), 30, maximum_coverage_percentage))

#impute NA-values of subdivision size to 700
dataset <- dataset %>% mutate(subdivision_size = if_else(is.na(subdivision_size), 700, subdivision_size))


#impute NA in distance variables to their cap values 
dataset <- dataset %>% mutate(across(c(lake_distance,
                                       stream_distance,
                                       coast_distance,
                                       road_distance,
                                       highway_distance,
                                       train_station_distance,
                                       railway_distance,
                                       windturbine_distance),
                                     ~ replace_na(., max(., na.rm = TRUE))))

#check that there are no remaining NAs in property sales data
which(is.na(subset(dataset, type == "Property sale")), arr.ind = TRUE)


```


## 3. Auxiliary variables
For the lot sales model the following auxiliary variables are calculated: 

- Gross floor area: Defined as the maximum allowed maximum coverage percentage x lot size
- Remaining floor area: Defined as lot size - gross floor area and bounded at 0. 
- Unused floor area: Defined as gross floor area - living space and bounded at 0.
- Number of subdivisions: Defined as lot size / subdivision size rounded down to nearest integer and bounded at 1. 

```{r Auxiliary variables, eval=TRUE, echo=TRUE, warning=FALSE, results= 'hide'}

dataset <- dataset %>% mutate(gross_floor_area = lot_size * maximum_coverage_percentage/100,
                              remaining_floor_area = pmax(0, lot_size - gross_floor_area),
                              unused_floor_area = pmax(0, gross_floor_area - living_space),
                              number_of_divisions = pmax(1, floor(lot_size / subdivision_size)))


```


## 4. Extract property types 
Extract different types of properties in the dataset for working with them separately. 


```{r Property types, eval=TRUE, echo=TRUE, warning=FALSE, results= 'hide'}
dataset_property_sales <- dataset %>% filter(type == "Property sale")
dataset_lot_sales <- dataset %>% filter(type == "Lot sale")
dataset_properties <- dataset %>% 
                      filter(type == "Property") %>% 
                      mutate(sales_date_numeric = 2024.0) #we want to value properties 1 January 2024


```

## 5. Cross-Validation
A simple 4-fold cross-validation split is implemented in the code below. 

```{r Cross-validation, eval=TRUE, echo=TRUE, warning=FALSE, results = 'hide'}

assign_cv_folds <- function(data, N, seed = 123) {
  set.seed(seed)
  data %>% mutate(validation_set = sample(rep(1:N, length.out = n())),
                  .rand = runif(n()))
}

#property sales
dataset_property_sales_cv <- assign_cv_folds(dataset_property_sales, N = N, seed = 123)

table(dataset_property_sales_cv$validation_set, useNA = "ifany")


#lot sales
dataset_lot_sales_cv <- assign_cv_folds(dataset_lot_sales, N = N, seed = 123)

table(dataset_lot_sales_cv$validation_set, useNA = "ifany")

```


# GAM Models 

## 1. Property model
The code below fits the GAM model structure described in "Generalized Additive Models for Real Estate Valuation: A Case Study of the Danish Housing Market" based on the cross validation split employed. 

```{r Model training, eval=TRUE, echo=TRUE, warning=FALSE, results='hide'}
fit_fold <- function(i, data, k_spatial, k_other = 10, fraction = 1.0,
                     checkpoint_prefix = "Model", save_checkpoint = TRUE,
                     seed = 123) {
  
  training_subset <- data %>% filter(validation_set != i)

  if (fraction < 1.0) {
    
    training_subset <- training_subset %>%
                       group_by(municipality_type) %>%
                       arrange(.rand, .by_group = TRUE) %>%
                       slice_head(prop = fraction) %>%
                       ungroup() 
  }
  
  model <- bam(sales_price ~ 
               s(coordinate_east, coordinate_north, k = k_spatial, pc = c(697000, 6150000)) +
               s(sales_date_numeric, k = k_other, by = municipality_type, pc = 2024) +
               s(year_built, k = k_other, by = municipality_type, pc = 1970) +
               s(year_rebuilt, k = k_other, by = municipality_type, pc = 1970) +
               s(living_space, k = k_other, by = municipality_type, pc = 140) +
               s(lot_size, k = k_other, by = municipality_type, pc = 800) +
               s(coast_distance, k = k_other, pc = 1500) +
               s(lake_distance, k = k_other, pc = 500) +
               s(stream_distance, k = k_other, pc = 1000) +
               s(road_distance, k = k_other, pc = 2000) +
               s(highway_distance, k = k_other, pc = 2000) +
               s(train_station_distance, k = k_other, pc = 2000) +
               s(railway_distance, k = k_other, pc = 1000) +
               s(windturbine_distance, k = k_other, pc = 3000) +
               s(ocean_view, k = k_other, pc = 0) +
               s(lake_view, k = k_other, pc = 0) +
               s(forest_size, k = k_other, pc = 0) +
               s(number_of_bathrooms, by = municipality_type, k = 2, pc = 1) +
               house_type * municipality_type +
               roof_type * municipality_type +
               wall_type * municipality_type +
               heating_type * municipality_type,
               data = training_subset,
               chunk.size = nrow(training_subset),
               family = tw(link = log),
               method = "fREML")
  
  if (save_checkpoint) {
    filename <- sprintf("%s_fold%d_k%d_frac%.2f.rds", checkpoint_prefix, i, k_spatial, fraction)
    saveRDS(model, file = file.path("checkpoints_lot_sales", filename))
  }
  
  model
}

fit_model_ensemble <- function(data, N, k_spatial, k_other = 10, fraction = 1.0, checkpoint_prefix = "Model") {
  
  models <- future_lapply(1:N, function(i) {
    fit_fold(
      i = i,
      data = data,
      k_spatial = k_spatial,
      k_other = k_other,
      fraction = fraction,
      checkpoint_prefix = checkpoint_prefix
    )
  })
  names(models) <- paste0("Model", 1:N)
  models
}

if (load) {
  models <- readRDS("saved_files_lot_sales/models.rds")
} else {
  models <- fit_model_ensemble(
    data = dataset_property_sales_cv,
    N = N,
    k_spatial = k_spatial,
    k_other = k_other,
    fraction = 1.0,
    checkpoint_prefix = "main"
  )
  saveRDS(models, "saved_files_lot_sales/models.rds")
}

```

#### Smooth effects 

```{r Other smooth effects, eval=TRUE, echo=TRUE, warning=FALSE, results = 'hide', fig.width=20, fig.height=30}

smooth_variables <- c(
  sales_date_numeric = "Sales date",
  year_built = "Year built",
  year_rebuilt = "Year rebuilt",
  living_space = "Living space [m²]",
  lot_size = "Lot size [m²]",
  unused_floor_area = "Unused floor area [m²]",
  number_of_divisions = "Number of divisions",
  coast_distance = "Distance to coast [m]",
  lake_distance = "Distance to lake [m]",
  stream_distance = "Distance to stream [m]",
  road_distance = "Distance to road [m]",
  highway_distance = "Distance to highway [m]",
  train_station_distance = "Distance to train station [m]",
  railway_distance = "Distance to railway [m]",
  windturbine_distance = "Distance to wind turbine [m]",
  ocean_view = "Ocean view [m]",
  lake_view = "Lake view [m]",
  forest_size = "Forest size [ha]",
  number_of_bathrooms = "Number of bathrooms")

plot_smooths <- list()

for (var in names(smooth_variables)) {

  smooth_dfs <- lapply(names(models), function(model_name) {
    mod <- models[[model_name]]
    df <- tryCatch(
      smooth_estimates(mod, smooth = paste0("s(", var, ")"), partial_match = TRUE),
      error = function(e) {
        message(sprintf("Skipping '%s' in %s: %s", var, model_name, e$message))
        return(NULL)
      })
    if (!is.null(df)) {
      df$Model <- sub("^Model", "Model ", model_name)
    }
    df
  })

  smooth_dfs <- Filter(Negate(is.null), smooth_dfs)  # remove NULLs
  if (length(smooth_dfs) == 0) next

  combined_df <- bind_rows(smooth_dfs)

  has_by <- any(!is.na(combined_df$.by))
  plot_title <- paste("Partial effect of", gsub("_", " ", var))

  if (has_by) {

    by_var_name <- unique(na.omit(combined_df$.by))[[1]]

    p <- ggplot(combined_df, aes(x = .data[[var]], y = .estimate,
                                 color = .data[[by_var_name]],
                                 linetype = Model, 
                                 group = interaction(Model, .data[[by_var_name]]))) +
        geom_ribbon(data = subset(combined_df, Model == "Model 1"),
                    aes(ymin = .estimate - 1.96*.se, ymax = .estimate + 1.96*.se),
                    alpha = 0.5, 
                    color = NA) +
        geom_line(linewidth = 1) +
        labs(title = plot_title,
             x = smooth_variables[[var]],
             y = "Partial effect") +
        scale_color_brewer(palette = "Dark2", name = "Municipality type") +
        scale_linetype_manual(values = c("solid", 
                                         "dashed", 
                                         "dotted", 
                                         "dotdash", 
                                         "twodash", 
                                         "longdash")) +
        guides(fill = guide_legend(override.aes = list(linetype = 0)))

  } else {

    p <- ggplot(combined_df, aes(x = .data[[var]], y = .estimate, linetype = Model)) +
         geom_ribbon(data = subset(combined_df, Model == "Model 1"),
                     aes(ymin = .estimate - 1.96*.se, ymax = .estimate + 1.96*.se),
                     alpha = 0.5, 
                     color = NA) +
         geom_line(linewidth = 1, color = "black") +
         labs(title = plot_title,
              x = smooth_variables[[var]],
              y = "Partial effect") +
         scale_linetype_manual(values = c("solid", 
                                          "dashed", 
                                          "dotted", 
                                          "dotdash", 
                                          "twodash", 
                                          "longdash"))
  }
  
  p <- p + theme(legend.position = "none")

  plot_smooths[[var]] <- p
}


#manual axis specifications of some ill-behaved smooths to make plots easier to understand
plot_smooths$stream_distance <- plot_smooths$stream_distance + ylim(-0.03,0.1)
plot_smooths$lake_view <- plot_smooths$lake_view + xlim(0.0,10000) + ylim(0, 0.6)
plot_smooths$forest_size <- plot_smooths$forest_size + xlim(0.0,40) + ylim(-0.02, 0.1)

wrap_plots(plot_smooths, ncol = 2) + 
           plot_layout(guides = "collect") &
           theme(legend.position = "bottom",
                 legend.text = element_text(size = 12),
                 legend.title = element_text(size = 13, face = "bold"),
                 legend.box = "horizontal",
                 legend.justification = "center")
```


## Model performance

```{r Model performance, eval=TRUE, echo=TRUE, warning=FALSE}

get_residuals <- function(models, dataset, fold_from_name = TRUE) {
  
  residual_list <- future_lapply(seq_along(models), function(i) {
    
    model_name <- names(models)[i]
    model <- models[[i]]
    
    if (fold_from_name) {
      fold <- as.integer(sub("^Model(\\d+).*$", "\\1", model_name))
    } else {
      fold <- i
    }
    
    predicted_sales_price <- predict.gam(model, newdata = dataset, type = "response")
    
    tibble(sales_price = dataset$sales_price,
           sales_date_numeric = dataset$sales_date_numeric,
           predicted_sales_price = predicted_sales_price,
           residual_sales_price = predicted_sales_price - dataset$sales_price,
           residual_sales_price_percent = (predicted_sales_price - dataset$sales_price) / dataset$sales_price * 100,
           residual_sales_price_log = log(predicted_sales_price) - log(dataset$sales_price),
           ratio = predicted_sales_price / dataset$sales_price,
           validation = dataset$validation_set == fold,
           id = dataset$id,
           municipality_type = dataset$municipality_type,
           Model = model_name)
    
  })
  
  bind_rows(residual_list)
}

residual_df <- get_residuals(models, dataset_property_sales_cv)
saveRDS(residual_df, "saved_files_lot_sales/residual_df.rds")

performance_countrywide <- residual_df %>%
                           mutate(dataset_type = if_else(validation, "Validation", "Training")) %>%
                           group_by(dataset_type) %>%
                           summarise(municipality_type = "Countrywide",
                                     MALE = mean(abs(residual_sales_price_log)),
                                     MAPE = mean(abs(residual_sales_price_percent)),
                                     MAE = mean(abs(residual_sales_price)),
                                     PM20 = mean(abs(residual_sales_price_percent) <= 20) * 100,
                                     COD = mean(abs(ratio - median(ratio))) / median(ratio) * 100,
                                     PRD = mean(ratio) / (sum(predicted_sales_price) / sum(sales_price)),
                                     .groups = "drop")

performance_municipality_type <- residual_df %>%
                                 mutate(dataset_type = if_else(validation, "Validation", "Training")) %>%
                                 group_by(dataset_type, municipality_type) %>%
                                 summarise(MALE = mean(abs(residual_sales_price_log)),
                                           MAPE = mean(abs(residual_sales_price_percent)),
                                           MAE = mean(abs(residual_sales_price)),
                                           PM20 = mean(abs(residual_sales_price_percent) <= 20) * 100,
                                           COD = mean(abs(ratio - median(ratio))) / median(ratio) * 100,
                                           .groups = "drop")

performance <- bind_rows(performance_countrywide, performance_municipality_type)

municipality_levels <- c("Countrywide", sort(setdiff(unique(performance$municipality_type), "Countrywide")))

performance_sorted <- performance %>%
                      mutate(dataset_type = factor(dataset_type, levels = c("Training", "Validation")),
                             municipality_type = factor(municipality_type, levels = municipality_levels),
                             MAE = round(MAE,0),
                             PRD = ifelse(is.na(PRD), "", round(PRD, 2))) %>%
                      arrange(municipality_type, dataset_type)  %>% 
                      rename(Type = dataset_type,
                             `Municipality type` = municipality_type,
                             `MAPE [%]` = MAPE,
                             `MAE [DKK]` = MAE,
                             `PM20 [%]` = PM20,
                             `COD [%]` = COD)

#standard table
#knitr::kable(performance_sorted, digits = 2, caption = "Model Performance by Municipality Type")

#nicer table
kable(performance_sorted %>% 
select(-`Municipality type`), 
       digits = 2, 
       caption = "Model Performance by Municipality Type", 
       booktabs = TRUE) %>%
kable_styling(full_width = FALSE, position = "center") %>%
group_rows("Countrywide", 1, 2) %>%
group_rows("Capital", 3, 4) %>%
group_rows("Commuter", 5, 6) %>%
group_rows("Large city", 7, 8) %>%
group_rows("Provincial", 9, 10) %>%
group_rows("Rural", 11, 12)

tbl_latex <- kable(performance_sorted %>% select(-`Municipality type`),
                   format = "latex", booktabs = TRUE, digits = 2,
                   caption = "Model performance by municipality type.",
                   label = "perf_by_muni") %>%
             kable_styling(full_width = FALSE, 
                           position = "center", 
                           latex_options = c("hold_position")) %>%
             group_rows("Countrywide", 1, 2) %>%
             group_rows("Capital", 3, 4) %>%
             group_rows("Commuter", 5, 6) %>%
             group_rows("Large city", 7, 8) %>%
             group_rows("Provincial", 9, 10) %>%
             group_rows("Rural", 11, 12)

```




#### 5. Standardization
The lot sales and properties are augmented with a fictitious single family house corresponding to a set of standard building characteristics. Their standardized property value is subsequently calculated. 

```{r Standardization, eval=TRUE, echo=TRUE, warning=FALSE, results= 'hide'}


standardize_property <- function(data, reference_data) {
  
    data %>% mutate(year_built = 1970,
                    year_rebuilt = 1970,
                    living_space = 140,
                    lot_size = 800,
                    maximum_coverage_percentage = 30,
                    unused_floor_area = 0,
                    number_of_divisions = 1,
                    number_of_bathrooms = 1,
                    house_type = factor("Detached", levels = levels(factor(reference_data$house_type))),
                    roof_type = factor("Fiber cement", levels = levels(factor(reference_data$roof_type))),
                    wall_type = factor("Brick", levels = levels(factor(reference_data$wall_type))),
                    heating_type = factor("District heating", levels = levels(factor(reference_data$heating_type))))
  }


if (load) {
  
  dataset_properties <- readRDS("saved_files_lot_sales/dataset_properties_standardized.rds")
  dataset_lot_sales_cv <- readRDS("saved_files_lot_sales/dataset_lot_sales_cv_standardized.rds")
  

} else {
  
  #standardize
  dataset_lot_sales_cv_standardized <- standardize_property(dataset_lot_sales_cv, reference_data = dataset)
  dataset_properties_standardized <- standardize_property(dataset_properties, reference_data = dataset)
  
  #calculate standarized prices for each model 
  for (model_name in names(models)) {
    
    dataset_lot_sales_cv[[paste0("standard_property_value_", model_name)]] <- as.vector(predict.gam(models[[model_name]], dataset_lot_sales_cv_standardized, type = "response"))
    dataset_properties[[paste0("standard_property_value_", model_name)]] <- as.vector(predict.gam(models[[model_name]], dataset_properties_standardized, type = "response"))
    dataset_properties[[paste0("property_value_", model_name)]] <- as.vector(predict.gam(models[[model_name]], dataset_properties, type = "response"))
  
  }

  #save
  saveRDS(dataset_lot_sales_cv, "saved_files_lot_sales/dataset_lot_sales_cv_standardized.rds")
  saveRDS(dataset_properties, "saved_files_lot_sales/dataset_properties_standardized.rds")
  
}




```


#### 6. Lot models

```{r Lot models, eval=TRUE, echo=TRUE, warning=FALSE, results= 'hide'}


# -------------------------
# 1. Model fitting functions
# -------------------------

fit_fold_model <- function(formula, data, fold) {
  
  training_subset <- data %>% filter(validation_set != fold)
  
  bam(formula,
      data = training_subset,
      family = tw(link = log),
      method = "fREML")
  
}

fit_model_set <- function(formula, data, N) {
  
  models <- future_lapply(seq_len(N), function(i) {
    
    fit_fold_model(formula = formula,
                   data = data,
                   fold = i)
    
  })
  
  names(models) <- paste0("Model", seq_len(N))
  
  models
  
}

# -------------------------
# 2. Formula helpers
# -------------------------

make_formula_from_terms <- function(terms) {as.formula(paste("sales_price ~", paste(terms, collapse = " + ")))}

make_standardized_formula <- function(std_var, extra_terms = NULL, coord_term = NULL) {
  
  rhs <- c(paste0("s(", std_var, ", k = 10)"), extra_terms, coord_term)
  
  make_formula_from_terms(rhs)
}

# -------------------------
# 3. Standardized-price model grid
# -------------------------

standard_vars <- paste0("standard_property_value_Model", seq_len(N))

base_terms <- list(
  
  none = NULL,
  
  lot_size_coverage_divisions= c("s(lot_size, k = 6, pc = 800)", "s(maximum_coverage_percentage, k = 3, pc = 30)", "s(number_of_divisions, k = 3, pc = 1)"),
  
  gross_remaining_area_divisions = c("s(gross_floor_area, k = 6, pc = 240)", "s(remaining_floor_area, k = 6, pc = 560)",  "s(number_of_divisions, k = 3, pc = 1)")
  
)

coord_terms <- list(no_coords = NULL, coords = "s(coordinate_east, coordinate_north, k = 500, pc = c(697000, 6150000))")

model_grid <- expand.grid(std_var = standard_vars, base_name = names(base_terms), coord_name = names(coord_terms), stringsAsFactors = FALSE)

model_formulas <- pmap(model_grid,
                       
  function(std_var, base_name, coord_name) {
    make_standardized_formula(
      std_var = std_var,
      extra_terms = base_terms[[base_name]],
      coord_term = coord_terms[[coord_name]]
    )
  }
)

names(model_formulas) <- paste0(
  "model_",
  seq_along(model_formulas),
  "_",
  model_grid$std_var,
  "_",
  model_grid$base_name,
  "_",
  model_grid$coord_name
)

# -------------------------
# 4. Direct lot model terms
# -------------------------

direct_lot_terms <- c("s(coordinate_east, coordinate_north, k = 300, pc = c(697000, 6150000))",
                      "s(sales_date_numeric, k = 5, by = municipality_type, pc = 2024)",
                      "s(lot_size, k = 5, by = municipality_type, pc = 800)",
                      "s(coast_distance, k = 5, pc = 1500)",
                      "s(lake_distance, k = 5, pc = 500)",
                      "s(stream_distance, k = 5, pc = 1000)",
                      "s(road_distance, k = 5, pc = 2000)",
                      "s(highway_distance, k = 5, pc = 2000)",
                      "s(train_station_distance, k = 5, pc = 2000)",
                      "s(railway_distance, k = 5, pc = 1000)",
                      "s(windturbine_distance, k = 5, pc = 3000)",
                      "s(ocean_view, k = 5, pc = 0)",
                      "s(lake_view, k = 5, pc = 0)",
                      "s(forest_size, k = 5, pc = 0)")

# -------------------------
# 5. Extra model families
# -------------------------

extra_formulas <- list()

# 5.1 Direct lot model, no standardized price
extra_formulas[["direct_lot_model"]] <- make_formula_from_terms(direct_lot_terms)


# 5.2 Standardized price by municipality type + coordinates + lot size + coverage + divisions
#for (std_var in standard_vars) {
  
#  extra_formulas[[paste0("standard_price_municipality_plus_all_", std_var)]] <-
#    make_formula_from_terms(c(
#      "municipality_type",
#      paste0("s(", std_var, ", k = 10, by = municipality_type)"),
#      "s(coordinate_east, coordinate_north, k = 300, pc = c(697000, 6150000))",
#      "s(lot_size, k = 6, pc = 800)",
#      "s(number_of_divisions, k = 3, pc = 1)",
#      "s(maximum_coverage_percentage, k = 3, pc = 30)"

#    ))
#}


# Combine all formulas
model_formulas <- c(model_formulas, extra_formulas)


if (load) {
  
  models <- readRDS("saved_files_lot_sales/models_lot_sales.rds")

} else {

  models <- lapply(names(model_formulas), function(model_name) {
  
  message("Fitting ", model_name, " ...")
  
  fit_model_set(formula = model_formulas[[model_name]], data = dataset_lot_sales_cv, N = N)})

  names(models) <- names(model_formulas)

  saveRDS(models, "saved_files_lot_sales/models_lot_sales.rds")
  
}


```




#### 7. Smooth effect of standard prices 

```{r Smooth effects, eval=TRUE, echo=TRUE, warning=FALSE, results = 'hide', fig.width=20, fig.height=30}

# -------------------------
# 1. Extract standardized-price smooths
# -------------------------

extract_standard_price_smooths <- function(models, x_min = 0, x_max = 1e7, n = 200) {
  
  imap_dfr(models, function(model_set, model_name) {
    
    std_var <- str_extract(model_name, "standard_property_value_Model[1-4]")
    
    if (is.na(std_var)) return(tibble())
    
    imap_dfr(model_set, function(mod, lot_split_name) {
      
      new_data <- tibble(!!std_var := seq(x_min, x_max, length.out = n),
                         lot_size = 800,
                         number_of_divisions = 1,
                         maximum_coverage_percentage = 30,
                         gross_floor_area = 240,
                         remaining_floor_area = 560,
                         coordinate_east = 697000,
                         coordinate_north = 6150000)
      
      df <- tryCatch(
        smooth_estimates(
          mod,
          smooth = paste0("s(", std_var, ")"),
          data = new_data,
          partial_match = TRUE
        ),
        error = function(e) {
          message(sprintf("Skipping %s / %s: %s", model_name, lot_split_name, e$message))
          return(NULL)
        }
      )
      
      if (is.null(df)) return(tibble())
      
      intercept <- coef(mod)[["(Intercept)"]]
      
      df %>% mutate(model_name = model_name,
                    property_model = str_extract(std_var, "Model[1-4]"),
                    lot_split = sub("^Model", "Lot split ", lot_split_name),
                    specification = model_name %>%
                    str_remove("^model_[0-9]+_") %>%
                    str_remove("standard_property_value_Model[1-4]_"),
                    standardized_price = .data[[std_var]],
                    partial_effect = .estimate,
                    partial_lower = .estimate - 1.96 * .se,
                    partial_upper = .estimate + 1.96 * .se,
                    price_effect = exp(intercept + .estimate),
                    price_lower = exp(intercept + .estimate - 1.96 * .se),
                    price_upper = exp(intercept + .estimate + 1.96 * .se)
        )
    })
  })
}

combined_df <- extract_standard_price_smooths(models)

# -------------------------
# 2. Confidence interval subset
# -------------------------

ci_df <- combined_df %>% filter(property_model == "Model1", lot_split == "Lot split 1")

# -------------------------
# 3. Link-scale plot
# -------------------------

plot_standardized_smooth <- ggplot(combined_df,
                            aes(x = standardized_price,
                                y = partial_effect,
                                color = lot_split,
                                linetype = property_model,
                                group = interaction(property_model, lot_split))) +
                            geom_ribbon(data = ci_df,
                                        aes(x = standardized_price,
                                            ymin = partial_lower,
                                            ymax = partial_upper),
                                        inherit.aes = FALSE,
                                        alpha = 0.25,
                                        fill = "grey60") +
                            geom_line(linewidth = 0.8, alpha = 0.8) +
                            facet_wrap(~ specification, scales = "free_y") +
                            scale_x_continuous(limits = c(0, 1e7), labels = function(x) paste0(x / 1e6, "M")) +
                            labs(x = "Standard property value [kr.]",
                                 y = "Partial effect",
                                 color = "Lot split",
                                 linetype = "Property split") +
                            theme_minimal()

plot_standardized_smooth

# -------------------------
# 4. Response-scale plot
# -------------------------

plot_standardized_smooth_response <- ggplot(combined_df,
                                     aes(x = standardized_price,
                                         y = price_effect,
                                         color = lot_split,
                                         linetype = property_model,
                                         group = interaction(property_model, lot_split))) +
                                     geom_abline(slope = 1,
                                                 intercept = 0,
                                                 color = "black",
                                                 linetype = "dashed",
                                                 linewidth = 0.8) +
                                     geom_ribbon(data = ci_df,
                                                 aes(x = standardized_price,
                                                     ymin = price_lower,
                                                     ymax = price_upper),
                                                 inherit.aes = FALSE,
                                                 alpha = 0.25,
                                                 fill = "grey60") +
                                     geom_line(linewidth = 0.8, alpha = 0.8) +
                                     facet_wrap(~ specification, scales = "free_y") +
                                     scale_x_continuous(limits = c(0, 0.75 * 1e7), labels = function(x) paste0(x / 1e6, "M")) +
                                     scale_y_continuous(labels = function(x) paste0(round(x / 1e6, 1), "M")) +
                                     coord_cartesian(xlim = c(0, 0.75 * 1e7),
                                                     ylim = c(0, 0.75 * 1e7)) +
                                     labs(x = "Standard property value [kr.]",
                                          y = "Standard lot value [kr.]",
                                          color = "Lot split",
                                         linetype = "Property split") +
                                     theme_minimal()

plot_standardized_smooth_response

```

#### 8. Other 1D smooth effects 

```{r Other 1D smooth effects, eval=TRUE, echo=TRUE, warning=FALSE, results = 'hide', fig.width=20, fig.height=30}

# -------------------------
# 1. Reference values/ranges
# -------------------------

reference_values <- list(standard_property_value_Model1 = 5e6,
                         standard_property_value_Model2 = 5e6,
                         standard_property_value_Model3 = 5e6,
                         standard_property_value_Model4 = 5e6,
                         lot_size = 800,
                         number_of_divisions = 1,
                         maximum_coverage_percentage = 30,
                         gross_floor_area = 240,
                         remaining_floor_area = 560,
                         coordinate_east = 697000,
                         coordinate_north = 6150000)

smooth_ranges <- list(lot_size = c(0, 3000),
                      number_of_divisions = c(1, 4),
                      maximum_coverage_percentage = c(20, 50),
                      gross_floor_area = c(0, 1000),
                      remaining_floor_area = c(0, 2000))

smooth_vars_1d <- c("lot_size",
                    "number_of_divisions",
                    "maximum_coverage_percentage",
                    "gross_floor_area",
                    "remaining_floor_area")

# -------------------------
# 2. Extract 1D smooths
# -------------------------

extract_1d_smooths <- function(models, smooth_vars, reference_values, smooth_ranges, n = 200) {
  
  imap_dfr(models, function(model_set, model_name) {
    
    std_var <- str_extract(model_name, "standard_property_value_Model[1-4]")
    
    if (is.na(std_var)) return(tibble())
    
    imap_dfr(model_set, function(mod, lot_split_name) {
      
      available_smooths <- vapply(mod$smooth, `[[`, character(1), "label")
      
      map_dfr(smooth_vars, function(smooth_var) {
        
        smooth_label <- paste0("s(", smooth_var, ")")
        
        if (!smooth_label %in% available_smooths) {
          return(tibble())
        }
        
        x_seq <- seq(smooth_ranges[[smooth_var]][1],
                     smooth_ranges[[smooth_var]][2],
                     length.out = n)
        
        new_data <- as_tibble(reference_values)
        new_data <- new_data[rep(1, n), ]
        new_data[[std_var]] <- reference_values[[std_var]]
        new_data[[smooth_var]] <- x_seq
        
        df <- tryCatch(
          smooth_estimates(
            mod,
            smooth = smooth_label,
            data = new_data
          ),
          error = function(e) {
            message(sprintf(
              "Skipping %s / %s / %s: %s",
              model_name, lot_split_name, smooth_var, e$message
            ))
            return(NULL)
          }
        )
        
        if (is.null(df)) return(tibble())
        
        df %>% mutate(model_name = model_name,
                      specification = model_name %>% str_remove("^model_[0-9]+_") %>% str_remove("standard_property_value_Model[1-4]_"),
                      property_model = str_extract(std_var, "Model[1-4]"),
                      lot_split = sub("^Model", "Lot split ", lot_split_name),
                      smooth_var = smooth_var,
                      x_value = .data[[smooth_var]],
                      partial_effect = .estimate,
                      partial_lower = .estimate - 1.96 * .se,
                      partial_upper = .estimate + 1.96 * .se)
      })
    })
  })
}

combined_1d_df <- extract_1d_smooths(
  models = models,
  smooth_vars = smooth_vars_1d,
  reference_values = reference_values,
  smooth_ranges = smooth_ranges,
  n = 200
)

# -------------------------
# 3. Plot list: one plot per smooth variable
# -------------------------

plot_1d_smooths_list <- split(combined_1d_df, combined_1d_df$smooth_var) %>%
  imap(function(df_var, var_name) {
    
    ci_df <- df_var %>% filter(property_model == "Model1", lot_split == "Lot split 1")
    
    ggplot(df_var,
           aes(x = x_value,
               y = partial_effect,
               color = lot_split,
               linetype = property_model,
               group = interaction(property_model, lot_split))) +
    geom_ribbon(data = ci_df,
                aes(x = x_value,
                    ymin = partial_lower,
                    ymax = partial_upper),
                inherit.aes = FALSE,
                alpha = 0.25,
                fill = "grey60") +
    geom_line(linewidth = 0.8, alpha = 0.8) +
    facet_wrap(~ specification, scales = "free_y") +
    labs(title = var_name,
         x = var_name,
         y = "Partial effect",
         color = "Lot split",
         linetype = "Property split") +
      theme_minimal()
  })

# -------------------------
# 4. View examples
# -------------------------

plot_1d_smooths_list$lot_size
plot_1d_smooths_list$number_of_divisions
plot_1d_smooths_list$maximum_coverage_percentage
plot_1d_smooths_list$gross_floor_area
plot_1d_smooths_list$remaining_floor_area
```


#### 9. Spatial smooths 

```{r Spatial smooths, eval=TRUE, echo=TRUE, warning=FALSE, results = 'hide', fig.width=20, fig.height=30}

# -------------------------
# 1. Helpers
# -------------------------

extract_spatial_grid <- function(mod, coordinate_grid, reference_values) {
  
  smooth_labels <- vapply(mod$smooth, `[[`, character(1), "label")
  
  spatial_label <- smooth_labels[str_detect(smooth_labels, "coordinate_east") & str_detect(smooth_labels, "coordinate_north")]
  
  if (length(spatial_label) == 0) return(NULL)
  
  spatial_label <- spatial_label[1]
  
  grid <- coordinate_grid
  
  for (nm in names(reference_values)) {
    if (!nm %in% names(grid)) {grid[[nm]] <- reference_values[[nm]]}
  }
  
  model_data <- model.frame(mod)
  
  for (var in setdiff(all.vars(formula(mod)), names(grid))) {
    x <- model_data[[var]]
    
    if (is.factor(x)) {
      grid[[var]] <- factor(rep(levels(x)[1], nrow(grid)), levels = levels(x))
    } else {
      grid[[var]] <- 0
    }
  }
  
  terms <- predict.gam(mod,
                       newdata = grid,
                       type = "terms")
  
  df_terms <- as.data.frame(terms)
  
  grid$effect_coordinates <- df_terms[[spatial_label]]
  
  if ("ocean" %in% names(grid)) {
    grid$effect_coordinates[grid$ocean] <- NA
  }
  
  grid
}

plot_spatial_effect <- function(grid, title, effect_limits) {
  
  ggplot(grid) +
  geom_tile(aes(x = coordinate_east,
                y = coordinate_north,
                fill = effect_coordinates)) +
  scale_fill_scico(palette = "vik",
                   name = "Partial effect",
                   na.value = "white",
                   limits = effect_limits) +
  geom_contour(aes(x = coordinate_east,
                   y = coordinate_north,
                   z = as.numeric(!ocean)),
               color = "black",
               linewidth = 0.4) +
  coord_fixed() +
  labs(title = title, x = "", y = "") +
  theme_void() +
  theme(text = element_text(size = 12),
        legend.position = c(0.85, 0.75),
        legend.title = element_text(size = 9, margin = margin(b = 5)),
        legend.text = element_text(size = 9),
        legend.title.align = 0.5)
}

# -------------------------
# 2. Extract spatial grids
# -------------------------

reference_values <- list(standard_property_value_Model1 = 5e6,
                         standard_property_value_Model2 = 5e6,
                         standard_property_value_Model3 = 5e6,
                         standard_property_value_Model4 = 5e6,
                         lot_size = 800,
                         number_of_divisions = 1,
                         maximum_coverage_percentage = 30,
                         gross_floor_area = 240,
                         remaining_floor_area = 560)

spatial_grids <- imap(models, function(model_set, model_name) {
  
                 grids <- imap(model_set, function(mod, lot_split_name) {
                    
                    grid <- extract_spatial_grid(mod = mod,
                                                 coordinate_grid = coordinate_grid,
                                                 reference_values = reference_values)
                    grid
                    
                  })
                  
                grids <- compact(grids)
                  
                if (length(grids) == 0) return(NULL)
                  
                grids}) %>%
                compact()

# -------------------------
# 3. Global effect limits
# -------------------------

all_effects <- unlist(map(spatial_grids, function(model_set) {
                      map(model_set, ~ .x$effect_coordinates)}),
                      use.names = FALSE)

max_abs_effect <- max(abs(all_effects), na.rm = TRUE)
effect_limits <- c(-max_abs_effect, max_abs_effect)

# -------------------------
# 4. Make nested plot list
# -------------------------

plot_spatial_smooths <- imap(spatial_grids, function(model_set, model_name) {
  
  imap(model_set, function(grid, lot_split_name) {
    
    plot_spatial_effect(grid = grid,
                        title = paste(model_name, "/", lot_split_name),
                        effect_limits = effect_limits)
    
  })
})

saveRDS(plot_spatial_smooths, "saved_files_lot_sales/plot_spatial_smooths.rds")

#example plots
wrap_plots(plot_spatial_smooths$model_25_standard_property_value_Model1_none_coords, ncol = 4)
```


#### 9. Performance

```{r Performance, eval=TRUE, echo=TRUE, warning=FALSE, results = 'hide', fig.width=20, fig.height=30}

# -------------------------
# 1. Helpers
# -------------------------

extract_property_model <- function(model_name) {
  out <- str_extract(model_name, "standard_property_value_Model[1-4]") %>%
    str_extract("Model[1-4]")
  
  ifelse(is.na(out), "Direct", out)
}

extract_model_specification <- function(model_name) {
  model_name %>%
    str_remove("^model_[0-9]+_") %>%
    str_remove("_?standard_property_value_Model[1-4]?") %>%
    str_replace_all("__+", "_") %>%
    str_remove("_$")
}

# -------------------------
# 2. Residuals
# -------------------------

get_residuals_one_model_set <- function(model_set, dataset, model_label) {
  
  property_model <- extract_property_model(model_label)
  model_specification <- extract_model_specification(model_label)
  
  future_lapply(names(model_set), function(fold_model_name) {
    
    model <- model_set[[fold_model_name]]
    fold <- as.integer(str_extract(fold_model_name, "\\d+"))
    
    pred <- as.vector(predict.gam(model, newdata = dataset, type = "response"))
    
    obs <- dataset$sales_price
    
    tibble(Model = model_label,
           model_specification = model_specification,
           property_model = property_model,
           fold_model = fold_model_name,
           fold = fold,
           dataset_type = if_else(dataset$validation_set == fold, "Validation", "Training"),
           id = dataset$id,
           municipality_type = dataset$municipality_type,
           sales_price = obs,
           predicted_sales_price = pred,
           residual_sales_price = pred - obs,
           residual_sales_price_percent = (pred - obs) / obs * 100,
           residual_sales_price_log = log(pred) - log(obs),
           ratio = pred / obs)}) %>%
    bind_rows()
}

residual_df <- bind_rows(
  lapply(names(models), function(model_label) {
    message("Predicting residuals for ", model_label, " ...")
    
    get_residuals_one_model_set(
      model_set = models[[model_label]],
      dataset = dataset_lot_sales_cv,
      model_label = model_label
    )
  })
)

saveRDS(residual_df, "saved_files_lot_sales/residual_df_lot_sales.rds")

# -------------------------
# 3. Metric function
# -------------------------

summarise_performance_from_residuals <- function(residual_df, group_vars) {
  
  residual_df %>%
  group_by(across(all_of(group_vars))) %>%
  summarise(municipality_type = "Countrywide",
            MALE = mean(abs(residual_sales_price_log), na.rm = TRUE),
            MAPE = mean(abs(residual_sales_price_percent), na.rm = TRUE),
            MAE = mean(abs(residual_sales_price), na.rm = TRUE),
            PM20 = mean(abs(residual_sales_price_percent) <= 20, na.rm = TRUE) * 100,
            COD = mean(abs(ratio - median(ratio, na.rm = TRUE)), na.rm = TRUE) /
            median(ratio, na.rm = TRUE) * 100,
            PRD = mean(ratio, na.rm = TRUE) / (sum(predicted_sales_price, na.rm = TRUE) / sum(sales_price, na.rm = TRUE)),
            .groups = "drop")
}

# -------------------------
# 4. Exact model performance
# -------------------------

performance_exact <- summarise_performance_from_residuals(residual_df, group_vars = c("Model", "model_specification", "property_model", "dataset_type"))


# -------------------------
# 5. Average metrics over property models
# -------------------------

performance_by_specification <- performance_exact %>%
                                group_by(model_specification, dataset_type) %>%
                                summarise(across(c(MALE, MAPE, MAE, PM20, COD, PRD), ~ mean(.x, na.rm = TRUE)), .groups = "drop")

```


#### 10. Building and lot values

```{r Building and lot values, eval=TRUE, echo=TRUE, warning=FALSE, results = 'hide', fig.width=20, fig.height=30}

if (load) {
  
  property_decomposition <- readRDS("saved_files_lot_sales/property_decomposition.rds")

} else {

  property_decomposition <- list()

  for (model_name in names(models)) {
  
    property_model <- str_extract(model_name, "Model[1-4]")
  
    property_value_var <- paste0("property_value_", property_model)
    standard_property_value_var <- paste0("standard_property_value_", property_model)
  
    if (is.na(property_model)){
    
      property_value_var <- "property_value_Model1"
      standard_property_value_var <- "standard_property_value_Model1"
    
      }
  
    property_decomposition[[model_name]] <- list()
  
    for (lot_split_name in names(models[[model_name]])) {
      
      lot_value <- as.vector(predict.gam(models[[model_name]][[lot_split_name]], newdata = dataset_properties, type = "response"))
    
      property_decomposition[[model_name]][[lot_split_name]] <- tibble(id = dataset_properties$id,
                                                                       property_value = dataset_properties[[property_value_var]],
                                                                       standard_property_value = dataset_properties[[standard_property_value_var]],
                                                                       lot_value = lot_value,
                                                                       building_value = dataset_properties[[property_value_var]] - lot_value,
                                                                       standard_building_value = dataset_properties[[standard_property_value_var]] - lot_value)
    }
  }

  saveRDS(property_decomposition, "saved_files_lot_sales/property_decomposition.rds")
  
}


# -------------------------
# 1. Settings
# -------------------------

x_limits <- c(-2, 15)
density_adjust <- 1.5

# -------------------------
# 2. Helper: filtered density layer
# -------------------------

density_layer <- function(df, var, label, x_limits, density_adjust = 1.5) {
  
  if (!var %in% names(df)) return(NULL)
  
  df_layer <- df %>% transmute(value = .data[[var]] / 1e6) %>%
                     filter(!is.na(value),
                            value >= x_limits[1],
                            value <= x_limits[2])
  
  geom_density(data = df_layer,
               aes(x = value,
                   color = label,
                   fill = label),
               alpha = 0.3,
               linewidth = 0.8,
               adjust = density_adjust,
               na.rm = TRUE)
  
}

# -------------------------
# 3. Helper: mean text
# -------------------------

make_mean_text <- function(df) {
  
  vars <- c(property_value = "Property",
            standard_property_value = "Std property",
            lot_value = "Lot",
            building_value = "Building",
            standard_building_value = "Std building")
  
  vars <- vars[names(vars) %in% names(df)]
  
  mean_labels <- map2_chr(
    names(vars),
    vars,
    function(var, label) {
      mean_value <- mean(df[[var]], na.rm = TRUE) / 1e6
      paste0(label, ": ", round(mean_value, 2), "M")
    }
  )
  
  paste(mean_labels, collapse = "\n")
}

# -------------------------
# 4. Create density plots
# -------------------------

plot_value_densities <- imap(property_decomposition,
  
  function(model_set, model_name) {
    
    imap(model_set, function(df, lot_split_name) {
      
      mean_text <- make_mean_text(df)
      
      ggplot() +
      density_layer(df, 
                    "property_value",
                    "Property value",
                    x_limits,
                    density_adjust) +
      density_layer(df, 
                    "standard_property_value",
                    "Standard property value",
                    x_limits,
                    density_adjust) +
      density_layer(df, 
                    "lot_value",
                    "Lot value",
                    x_limits,
                    density_adjust) +
      density_layer(df, 
                    "building_value",
                    "Building value",
                    x_limits,
                    density_adjust) +
        density_layer(df, 
                      "standard_building_value",
                      "Standard building value",
                      x_limits,
                      density_adjust) +
        coord_cartesian(xlim = x_limits) +
        scale_x_continuous(labels = label_number(suffix = "M")) +
        scale_color_discrete(name = NULL) +
        scale_fill_discrete(name = NULL) +
        labs(
          title = paste(model_name, "-", lot_split_name),
          x = "Value [million DKK]",
          y = "Density"
        ) +
        annotate(
          "text",
          x = x_limits[1] + 0.55 * diff(x_limits),
          y = Inf,
          label = mean_text,
          hjust = 0,
          vjust = 1.1,
          size = 3.5
        ) +
        theme_minimal()
    })
  }
)

# -------------------------
# 5. Show plots
# -------------------------

wrap_plots(flatten(plot_value_densities), ncol = 4)

```


#### 11. Model stability

```{r Stability, eval=TRUE, echo=TRUE, warning=FALSE, results = 'hide', fig.width=20, fig.height=30}
dataset_properties <- dataset_properties[1:10000,]


# -------------------------
# 1. Helper: extract model specification
# -------------------------

extract_model_specification <- function(model_name) {
  
  model_name %>% 
  str_remove("^model_[0-9]+_") %>%
  str_remove("_?standard_property_value_Model[1-4]?") %>%
  str_replace_all("__+", "_") %>%
  str_remove("_$")
}

# -------------------------
# 2. Group models by specification
# -------------------------

models_by_specification <- imap(
  
  models,
  
  function(model_set, model_name) {
    
    model_specification <- extract_model_specification(model_name)
    
    imap(model_set, function(model, fold_name) {
      
      list(
        model = model,
        model_name = model_name,
        fold_name = fold_name,
        model_specification = model_specification
      )
    })
  }
) %>%
flatten() %>%
split(map_chr(., "model_specification"))

# -------------------------
# 3. Spread function
# -------------------------

spread_model_set <- function(model_set, dataset) {
  
  predictions <- imap(model_set, function(model_info, i) {
    
    preds <- predict.gam(model_info$model,
                         newdata = dataset,
                         type = "link")
    
    tibble(!!paste0(
        "pred_",
        model_info$model_name,
        "_",
        model_info$fold_name
      ) := as.vector(preds)
    )
  })
  
  pred_matrix <- bind_cols(predictions)
  pred_mat <- as.matrix(pred_matrix)
  
  bind_cols(dataset, pred_matrix) %>%
  mutate(mean_Ensemble = rowMeans(pred_mat, na.rm = TRUE),
         sigma_Ensemble = apply(pred_mat, 1, sd, na.rm = TRUE))
}

# -------------------------
# 4. Predict for each specification
# -------------------------

dataset_spread <- lapply(
  names(models_by_specification),
  
  function(model_specification) {
    
    message("Predicting for ", model_specification, " ...")
    
    spread_model_set(model_set = models_by_specification[[model_specification]],
                     dataset = dataset_properties)
  }
)

names(dataset_spread) <- names(models_by_specification)


cap <- 0.2

# -------------------------
# 5. Density data
# -------------------------

spread_long <- imap_dfr(dataset_spread, function(df, model_specification) {
  
  df %>% transmute(model_specification = model_specification,
                   sigma_Ensemble = sigma_Ensemble)
  
})

# -------------------------
# 6. Density plot
# -------------------------

plot_ensemble_density <- ggplot(spread_long,
                         aes(x = sigma_Ensemble,
                             color = model_specification,
                             fill = model_specification)) +
                         geom_density(alpha = 0.15, linewidth = 0.8, na.rm = TRUE) +
                         coord_cartesian(xlim = c(0, cap)) +
                         labs(x = "Ensemble spread (link scale)",
                              y = "Density",
                              color = NULL,
                              fill = NULL) +
                         theme_minimal()

plot_ensemble_density

# -------------------------
# 7. Base map
# -------------------------

base_map <- ggplot() +
            geom_contour(data = coordinate_grid,
            aes(x = coordinate_east,
                y = coordinate_north,
                z = as.numeric(!ocean)),
            color = "black",
            linewidth = 0.1) +
            coord_fixed() +
            theme_void(base_size = 10)

# -------------------------
# 8. One geographic plot per model specification
# -------------------------

plot_ensemble_geographic <- imap(dataset_spread, function(df, model_specification) {
  
  base_map + geom_point(data = df,
                        aes(x = coordinate_east,
                            y = coordinate_north,
                            color = sigma_Ensemble),
                        size = 0.2) +
             scale_color_viridis_c(option = "plasma",
                                   name = "Ensemble uncertainty",
                                   limits = c(0, cap),
                                   oob = squish) +
             labs(title = paste(model_specification, "- Ensemble uncertainty")) +
             theme(legend.position = c(0.7, 0.75),
                   legend.box = "horizontal")
})


```




#### 11. Combined property and lot model 

```{r Combined model, eval=TRUE, echo=TRUE, warning=FALSE, results = 'hide', fig.width=20, fig.height=30}

#Create combined sales set with standardized properties on lot sales

standardize_property <- function(data, reference_data) {
  
    data %>% mutate(year_built = 1970,
                    year_rebuilt = 1970,
                    living_space = 140,
                    maximum_coverage_percentage = 30,
                    unused_floor_area = 0,
                    number_of_divisions = 1,
                    number_of_bathrooms = 1,
                    house_type = factor("Detached", levels = levels(factor(reference_data$house_type))),
                    roof_type = factor("Fiber cement", levels = levels(factor(reference_data$roof_type))),
                    wall_type = factor("Brick", levels = levels(factor(reference_data$wall_type))),
                    heating_type = factor("District heating", levels = levels(factor(reference_data$heating_type))))
  }

dataset_lot_sales <- standardize_property(dataset_lot_sales, dataset)
dataset_sales <- bind_rows(dataset_property_sales, dataset_lot_sales)
dataset_sales$type <- factor(dataset_sales$type, levels = c("Property sale", "Lot sale"))

#assign folds
assign_cv_folds <- function(data, N, seed = 123) {
  set.seed(seed)
  data %>% mutate(validation_set = sample(rep(1:N, length.out = n())),
                  .rand = runif(n()))
}

dataset_sales_cv <- assign_cv_folds(dataset_sales, N = N, seed = 123)

table(dataset_sales_cv$validation_set, useNA = "ifany")


#fit models
fit_fold <- function(i, data, k_spatial, k_other = 10, fraction = 1.0,
                     checkpoint_prefix = "Model", save_checkpoint = TRUE,
                     seed = 123) {
  
  training_subset <- data %>% filter(validation_set != i)

  if (fraction < 1.0) {
    
    training_subset <- training_subset %>%
                       group_by(municipality_type) %>%
                       arrange(.rand, .by_group = TRUE) %>%
                       slice_head(prop = fraction) %>%
                       ungroup() 
  }
  
  model <- bam(sales_price ~ 
               s(coordinate_east, coordinate_north, k = k_spatial, pc = c(697000, 6150000)) +
               s(sales_date_numeric, k = k_other, by = municipality_type, pc = 2024) +
               s(year_built, k = k_other, by = municipality_type, pc = 1970) +
               s(year_rebuilt, k = k_other, by = municipality_type, pc = 1970) +
               s(living_space, k = k_other, by = municipality_type, pc = 140) +
               s(lot_size, k = k_other, by = municipality_type, pc = 800) +
               s(coast_distance, k = k_other, pc = 1500) +
               s(lake_distance, k = k_other, pc = 500) +
               s(stream_distance, k = k_other, pc = 1000) +
               s(road_distance, k = k_other, pc = 2000) +
               s(highway_distance, k = k_other, pc = 2000) +
               s(train_station_distance, k = k_other, pc = 2000) +
               s(railway_distance, k = k_other, pc = 1000) +
               s(windturbine_distance, k = k_other, pc = 3000) +
               s(ocean_view, k = k_other, pc = 0) +
               s(lake_view, k = k_other, pc = 0) +
               s(forest_size, k = k_other, pc = 0) +
               s(number_of_bathrooms, by = type, k = 2, pc = 1) +
               house_type * municipality_type +
               roof_type * municipality_type +
               wall_type * municipality_type +
               heating_type * municipality_type +
               type * municipality_type,
               data = training_subset,
               chunk.size = nrow(training_subset),
               family = tw(link = log),
               method = "fREML")
  
  if (save_checkpoint) {
    filename <- sprintf("%s_fold%d_k%d_frac%.2f.rds", checkpoint_prefix, i, k_spatial, fraction)
    saveRDS(model, file = file.path("checkpoints_lot_sales", filename))
  }
  
  model
}

fit_model_ensemble <- function(data, N, k_spatial, k_other = 10, fraction = 1.0, checkpoint_prefix = "Model") {
  
  models <- future_lapply(1:N, function(i) {
    fit_fold(
      i = i,
      data = data,
      k_spatial = k_spatial,
      k_other = k_other,
      fraction = fraction,
      checkpoint_prefix = checkpoint_prefix
    )
  })
  names(models) <- paste0("Model", 1:N)
  models
}

if (load) {
  models <- readRDS("saved_files_lot_sales/models_combined.rds")
} else {
  models <- fit_model_ensemble(
    data = dataset_sales_cv,
    N = N,
    k_spatial = k_spatial,
    k_other = k_other,
    fraction = 1.0,
    checkpoint_prefix = "main"
  )
  saveRDS(models, "saved_files_lot_sales/models_combined.rds")
}

```


```{r Spatial smooth 1, eval=TRUE, echo=TRUE, warning=FALSE, results = 'hide', fig.width=20, fig.height=30}


# 1a) Predict once per model and store results
prediction_results <- future_lapply(seq_along(models), function(i) {
  model <- models[[i]]
  model_data <- model.frame(model)
  grid <- coordinate_grid

  for (var in setdiff(all.vars(formula(model)), names(grid))) {
    x <- model_data[[var]]
    if (is.factor(x)) {
      grid[[var]] <- factor(rep(levels(x)[1], nrow(grid)), levels = levels(x))
    } else {
      grid[[var]] <- 0
    }
  }

  terms <- predict.gam(model, grid, type = "terms", se.fit = TRUE)
  df_predictions <- as.data.frame(terms$fit)
  df_se <- as.data.frame(terms$se.fit)

  effect <- df_predictions$`s(coordinate_east,coordinate_north)`
  se <- df_se$`s(coordinate_east,coordinate_north)`
  effect[grid$ocean] <- NA
  se[grid$ocean] <- NA

  grid$effect_coordinates <- effect
  grid$se_coordinates <- se

  list(model_name = names(models)[i],
       grid = grid)
})


# 1b) Compute SD across models (per grid cell) ---
effects_mat <- do.call(cbind, lapply(prediction_results, function(x) x$grid$effect_coordinates))
base_grid   <- prediction_results[[1]]$grid

se_across_models <- apply(effects_mat, 1, sd, na.rm = TRUE)
se_across_models[base_grid$ocean] <- NA

# 2. Global ranges (normalized across everything) ---
# Effects: symmetric across *all model* effects
all_effects <- unlist(lapply(prediction_results, function(x) x$grid$effect_coordinates))
max_abs_effect <- max(abs(all_effects), na.rm = TRUE)
effect_limits  <- c(-max_abs_effect, max_abs_effect)

# SEs: include per-model SEs AND the across-models SD
all_ses_models <- unlist(lapply(prediction_results, function(x) x$grid$se_coordinates))
se_limits <- c(0, max(c(all_ses_models, se_across_models), na.rm = TRUE))  # SEs are >= 0

# 3. Plotting function (adds land/ocean outline) ---
plot_model_terms_from_grid <- function(grid, model_name, effect_range, se_range) {

  plot_effect <- ggplot(data = grid) +
                 geom_tile(aes(coordinate_east, coordinate_north, fill = effect_coordinates)) +
                 labs(x = "", y = "") +
                 scale_fill_scico(palette = "vik",
                                  name = paste("Partial effect", sub("^Model", "Model ", model_name)),
                                  na.value = "white",
                                  limits = effect_range) +
                 geom_contour(aes(x = coordinate_east, y = coordinate_north, z = as.numeric(!ocean)),
                              color = "black", 
                              size = 0.4) +
                 coord_fixed() +
                 theme_void() +
                 theme(text = element_text(size = 12),
                       legend.position = c(0.85, 0.75),
                       legend.title = element_text(size = 9, margin = margin(b = 5)),
                       legend.text = element_text(size = 9),
                       legend.title.align = 0.5)
  
  model_label <- sub("^Model", "Model ", model_name)

  plot_se <- ggplot(data = grid) +
             geom_tile(aes(coordinate_east, coordinate_north, fill = se_coordinates)) +
             labs(x = "", y = "") +
             scale_fill_scico(
             palette = "vik",
             name = bquote(sigma[.(model_label)]),
             na.value = "white",
             limits = se_range) +
             geom_contour(aes(x = coordinate_east, y = coordinate_north, z = as.numeric(!ocean)),
                          color = "black", 
                          size = 0.4) +
             coord_fixed() +
             theme_void() +
             theme(text = element_text(size = 12),
                   legend.position = c(0.85, 0.75),
                   legend.title = element_text(size = 9, margin = margin(b = 5)),
                   legend.text = element_text(size = 9),
                   legend.title.align = 0.5)

  list(effect = plot_effect, se = plot_se)
}

# 4. Build plots for each model (effect + SE) ---
plot_lists <- lapply(prediction_results, function(result) {
  plot_model_terms_from_grid(
    result$grid, result$model_name,
    effect_range = effect_limits,
    se_range     = se_limits
  )
})

plot_spatial_smooths <- unlist(plot_lists, recursive = FALSE)

# 5. ONE extra plot: SD of effects across models (uses same se_limits) ---
base_grid2 <- base_grid
base_grid2$se_across_models <- se_across_models  # <- bake it into the data

plot_sd_across <- ggplot(data = base_grid2) +
                  geom_tile(aes(coordinate_east, coordinate_north, fill = se_across_models)) +
                  labs(x = "", y = "") +
                  scale_fill_scico(palette = "vik",
                                   name = expression(sigma[Ensemble]),
                                   na.value = "white",
                                   limits = se_limits) +
                  geom_contour(aes(x = coordinate_east, y = coordinate_north, z = as.numeric(!ocean)),
                               color = "black", 
                               size = 0.4) +
                  coord_fixed() +
                  theme_void() +
                  theme(text = element_text(size = 12),
                        legend.position = c(0.85, 0.75),
                        legend.title = element_text(size = 9, margin = margin(b = 5)),
                        legend.text = element_text(size = 9),
                        legend.title.align = 0.5)

plot_spatial_smooths <- c(plot_spatial_smooths, list(plot_sd_across))
wrap_plots(plot_spatial_smooths, ncol = 3)

```



```{r Other smooth effects 1, eval=TRUE, echo=TRUE, warning=FALSE, results = 'hide', fig.width=20, fig.height=30}

smooth_variables <- c(
  sales_date_numeric = "Sales date",
  year_built = "Year built",
  year_rebuilt = "Year rebuilt",
  living_space = "Living space [m²]",
  lot_size = "Lot size [m²]",
  coast_distance = "Distance to coast [m]",
  lake_distance = "Distance to lake [m]",
  stream_distance = "Distance to stream [m]",
  road_distance = "Distance to road [m]",
  highway_distance = "Distance to highway [m]",
  train_station_distance = "Distance to train station [m]",
  railway_distance = "Distance to railway [m]",
  windturbine_distance = "Distance to wind turbine [m]",
  ocean_view = "Ocean view [m]",
  lake_view = "Lake view [m]",
  forest_size = "Forest size [ha]",
  number_of_bathrooms = "Number of bathrooms")

plot_smooths <- list()

for (var in names(smooth_variables)) {

  smooth_dfs <- lapply(names(models), function(model_name) {
    mod <- models[[model_name]]
    df <- tryCatch(
      smooth_estimates(mod, smooth = paste0("s(", var, ")"), partial_match = TRUE),
      error = function(e) {
        message(sprintf("Skipping '%s' in %s: %s", var, model_name, e$message))
        return(NULL)
      })
    if (!is.null(df)) {
      df$Model <- sub("^Model", "Model ", model_name)
    }
    df
  })

  smooth_dfs <- Filter(Negate(is.null), smooth_dfs)  # remove NULLs
  if (length(smooth_dfs) == 0) next

  combined_df <- bind_rows(smooth_dfs)

  has_by <- any(!is.na(combined_df$.by))
  plot_title <- paste("Partial effect of", gsub("_", " ", var))

  if (has_by) {

    by_var_name <- unique(na.omit(combined_df$.by))[[1]]

    p <- ggplot(combined_df, aes(x = .data[[var]], y = .estimate,
                                 color = .data[[by_var_name]],
                                 linetype = Model, 
                                 group = interaction(Model, .data[[by_var_name]]))) +
         geom_ribbon(data = subset(combined_df, Model == "Model 1"),
                     aes(ymin = .estimate - 1.96*.se, ymax = .estimate + 1.96*.se),
                     alpha = 0.5, 
                     color = NA) +
         geom_line(linewidth = 1) +
         labs(title = plot_title,
              x = smooth_variables[[var]],
              y = "Partial effect") +
         scale_color_brewer(palette = "Dark2", name = "Municipality type") +
         scale_linetype_manual(values = c("solid", 
                                          "dashed", 
                                          "dotted", 
                                          "dotdash", 
                                          "twodash", 
                                          "longdash")) +
         guides(fill = guide_legend(override.aes = list(linetype = 0)))

  } else {

    p <- ggplot(combined_df, aes(x = .data[[var]], y = .estimate, linetype = Model)) +
         geom_ribbon(data = subset(combined_df, Model == "Model 1"),
                     aes(ymin = .estimate - 1.96*.se, ymax = .estimate + 1.96*.se),
                     alpha = 0.5, 
                     color = NA) +
         geom_line(linewidth = 1, color = "black") +
         labs(title = plot_title,
              x = smooth_variables[[var]],
              y = "Partial effect") +
         scale_linetype_manual(values = c("solid", 
                                          "dashed", 
                                          "dotted", 
                                          "dotdash", 
                                          "twodash", 
                                          "longdash"))
  }
  
  p <- p + theme(legend.position = "none")

  plot_smooths[[var]] <- p
}


#manual axis specifications of some ill-behaved smooths to make plots easier to understand
plot_smooths$stream_distance <- plot_smooths$stream_distance + ylim(-0.03,0.1)
plot_smooths$lake_view <- plot_smooths$lake_view + xlim(0.0,10000) + ylim(0, 0.6)
plot_smooths$forest_size <- plot_smooths$forest_size + xlim(0.0,40) + ylim(-0.02, 0.1)

wrap_plots(plot_smooths, ncol = 2) + 
           plot_layout(guides = "collect") &
           theme(legend.position = "bottom",
                 legend.text = element_text(size = 12),
                 legend.title = element_text(size = 13, face = "bold"),
                 legend.box = "horizontal",
                 legend.justification = "center")
```



```{r Factor effects 1, eval=TRUE, echo=TRUE, warning=FALSE, results = 'hide', fig.width=20, fig.height=20}
factor_variables <- c(
  house_type = "House type",
  type = "Type",
  municipality_type = "Municipality type",
  roof_type = "Roof type",
  wall_type = "Wall type",
  heating_type = "Heating type")

make_termwise_df <- function(model, model_data, variable, interaction_by = "municipality_type") {

  v_vals <- unique(model_data[[variable]])
  if (is.null(interaction_by)) {
    df <- data.frame(setNames(list(v_vals), variable), stringsAsFactors = FALSE)
  } else {
    m_vals <- unique(model_data[[interaction_by]])
    grid_list <- setNames(list(v_vals, m_vals), c(variable, interaction_by))
    df <- expand.grid(grid_list, stringsAsFactors = FALSE)
    df[[interaction_by]] <- factor(df[[interaction_by]], 
                                   levels = levels(model_data[[interaction_by]]))
  }
  for (var in setdiff(all.vars(formula(model)), names(df))) {
    x <- model_data[[var]]
    if (is.factor(x)) {
      df[[var]] <- factor(rep(levels(x)[1], nrow(df)), levels = levels(x))
    } else {
      df[[var]] <- 0
    }
  }
  return(df)
}

combine_all_term_effects <- function(df, term, interaction_by = "municipality_type") {
  pattern <- paste0("^", term, "$|^", term, ":" ,"|:", term, "$")
  cols_to_sum <- grep(pattern, names(df), value = TRUE)
  df[[paste0("effect_", term)]] <- rowSums(as.matrix(df[cols_to_sum]))
  return(df)
}

combine_all_term_se <- function(df, term, interaction_by = "municipality_type") {
  pattern <- paste0("^", term, "$|^", term, ":" ,"|:", term, "$")
  cols_to_sum <- grep(pattern, names(df), value = TRUE)
  df[[paste0("se_", term)]] <- sqrt(rowSums(as.matrix(df[cols_to_sum])^2))
  return(df)
}

plot_term_effect <- function(df, term, interaction_by = "municipality_type") {
  effect_col <- paste0("effect_", term)
  lower_col <- paste0("lower_", effect_col)
  upper_col <- paste0("upper_", effect_col)
  df_sub <- df %>% filter(term == !!term)
  plot_title <- paste("Partial effect of", gsub("_", " ", term))

  p <- ggplot(df_sub, aes(x = .data[[term]], 
                          y = .data[[effect_col]], 
                          shape = Model,
                          group = interaction(Model, if (identical(term, interaction_by) || 
                                                        !(interaction_by %in% names(df_sub))) NULL 
                                                     else .data[[interaction_by]]))) +
      geom_pointrange(data = subset(df_sub, Model == "Model 1"),
                      aes(ymin = .data[[lower_col]], ymax = .data[[upper_col]]),
                      position = position_dodge(width = 0.5),
                      size = 0.8) +
      geom_point(data = subset(df_sub, Model != "Model 1"),
                 position = position_dodge(width = 0.5),
                 size = 2.5) +
      coord_flip() +
      labs(x = NULL, y = "Partial effect", title = plot_title, shape = "Model") +
      theme(axis.text.y = element_text(angle = 45, hjust = 1))

  if (interaction_by %in% names(df_sub)) {
      color_var <- if (term == interaction_by) term else interaction_by
      p <- p + aes(color = .data[[color_var]]) +
      scale_color_brewer(palette = "Dark2", name = "Municipality type")
  }

  return(p)
}

factor_dfs <- list()

for (model_name in names(models)) {

  model <- models[[model_name]]
  model_data <- model.frame(model)
  relevant_factors <- factor_variables[names(factor_variables) %in% all.vars(formula(model))]

  for (var in names(relevant_factors)) {

    interaction <- if (var == "municipality_type") NULL else "municipality_type"
    df <- make_termwise_df(model, model_data, var, interaction_by = interaction)
    df$term <- var

    terms <- predict.gam(model, df, type = "terms", se.fit = TRUE)
    df_predictions <- as.data.frame(terms$fit)
    df_se <- as.data.frame(terms$se.fit)

    combined_predictions <- combine_all_term_effects(df_predictions, 
                                                     var, 
                                                     interaction_by = interaction) %>% 
                            select(starts_with("effect_"))
    
    combined_se <- combine_all_term_se(df_se, 
                                       var, 
                                       interaction_by = interaction) %>% 
                   select(starts_with("se_"))

    df_final <- df %>% select(all_of(c(var, interaction)), term) %>%
                       bind_cols(combined_predictions) %>%
                       bind_cols(combined_se) %>%
                       mutate(across(starts_with("effect_"),
                                     ~ . - 1.96 * get(str_replace(cur_column(), 
                                     "effect_", 
                                     "se_")),
                                     .names = "lower_{.col}")) %>%
                       mutate(across(starts_with("effect_"),
                                     ~ . + 1.96 * get(str_replace(cur_column(), "effect_", "se_")),
                                     .names = "upper_{.col}"))

    df_final$Model <- sub("^Model", "Model ", model_name)
    factor_dfs[[length(factor_dfs) + 1]] <- df_final
  }
}

df_all_factors <- bind_rows(factor_dfs)
plot_factors <- lapply(names(factor_variables), 
                       function(term) {plot_term_effect(df_all_factors, term)})
names(plot_factors) <- names(factor_variables)


wrap_plots(plot_factors, ncol = 2) +
           plot_layout(guides = "collect") &
           theme(legend.position = "bottom",
                 legend.text = element_text(size = 12),
                 legend.title = element_text(size = 13, face = "bold"),
                 legend.box = "horizontal",
                 legend.justification = "center")
```



```{r Model performance 1, eval=TRUE, echo=TRUE, warning=FALSE}

get_residuals <- function(models, dataset, fold_from_name = TRUE) {
  
  residual_list <- future_lapply(seq_along(models), function(i) {
    
    model_name <- names(models)[i]
    model <- models[[i]]
    
    if (fold_from_name) {
      fold <- as.integer(sub("^Model(\\d+).*$", "\\1", model_name))
    } else {
      fold <- i
    }
    
    predicted_sales_price <- predict.gam(model, newdata = dataset, type = "response")
    
    tibble(sales_price = dataset$sales_price,
           sales_date_numeric = dataset$sales_date_numeric,
           predicted_sales_price = predicted_sales_price,
           residual_sales_price = predicted_sales_price - dataset$sales_price,
           residual_sales_price_percent = (predicted_sales_price - dataset$sales_price) / dataset$sales_price * 100,
           residual_sales_price_log = log(predicted_sales_price) - log(dataset$sales_price),
           ratio = predicted_sales_price / dataset$sales_price,
           validation = dataset$validation_set == fold,
           id = dataset$id,
           type = dataset$type,
           municipality_type = dataset$municipality_type,
           Model = model_name)
    
  })
  
  bind_rows(residual_list)
}

residual_df <- get_residuals(models, dataset_sales_cv)
saveRDS(residual_df, "saved_files_lot_sales//residual_df_combined.rds")

performance_countrywide <- residual_df %>%
                           mutate(dataset_type = if_else(validation, "Validation", "Training")) %>%
                           group_by(dataset_type, type) %>%
                           summarise(municipality_type = "Countrywide",
                                     MALE = mean(abs(residual_sales_price_log)),
                                     MAPE = mean(abs(residual_sales_price_percent)),
                                     MAE = mean(abs(residual_sales_price)),
                                     PM20 = mean(abs(residual_sales_price_percent) <= 20) * 100,
                                     COD = mean(abs(ratio - median(ratio))) / median(ratio) * 100,
                                     PRD = mean(ratio) / (sum(predicted_sales_price) / sum(sales_price)),
                                     .groups = "drop")

performance_municipality_type <- residual_df %>%
                                 mutate(dataset_type = if_else(validation, "Validation", "Training")) %>%
                                 group_by(dataset_type, municipality_type) %>%
                                 summarise(MALE = mean(abs(residual_sales_price_log)),
                                           MAPE = mean(abs(residual_sales_price_percent)),
                                           MAE = mean(abs(residual_sales_price)),
                                           PM20 = mean(abs(residual_sales_price_percent) <= 20) * 100,
                                           COD = mean(abs(ratio - median(ratio))) / median(ratio) * 100,
                                           .groups = "drop")

performance <- bind_rows(performance_countrywide, performance_municipality_type)

municipality_levels <- c("Countrywide", sort(setdiff(unique(performance$municipality_type), "Countrywide")))

performance_sorted <- performance %>%
                      mutate(dataset_type = factor(dataset_type, levels = c("Training", "Validation")),
                             municipality_type = factor(municipality_type, levels = municipality_levels),
                             MAE = round(MAE,0),
                             PRD = ifelse(is.na(PRD), "", round(PRD, 2))) %>%
                      arrange(municipality_type, dataset_type)  %>% 
                      rename(Type = dataset_type,
                             `Municipality type` = municipality_type,
                             `MAPE [%]` = MAPE,
                             `MAE [DKK]` = MAE,
                             `PM20 [%]` = PM20,
                             `COD [%]` = COD)

#standard table
#knitr::kable(performance_sorted, digits = 2, caption = "Model Performance by Municipality Type")

#nicer table
kable(performance_sorted %>% 
select(-`Municipality type`), 
       digits = 2, 
       caption = "Model Performance by Municipality Type", 
       booktabs = TRUE) %>%
kable_styling(full_width = FALSE, position = "center") %>%
group_rows("Countrywide", 1, 2) %>%
group_rows("Capital", 3, 4) %>%
group_rows("Commuter", 5, 6) %>%
group_rows("Large city", 7, 8) %>%
group_rows("Provincial", 9, 10) %>%
group_rows("Rural", 11, 12)

tbl_latex <- kable(performance_sorted %>% select(-`Municipality type`),
                   format = "latex", booktabs = TRUE, digits = 2,
                   caption = "Model performance by municipality type.",
                   label = "perf_by_muni") %>%
             kable_styling(full_width = FALSE, 
                           position = "center", 
                           latex_options = c("hold_position")) %>%
             group_rows("Countrywide", 1, 2) %>%
             group_rows("Capital", 3, 4) %>%
             group_rows("Commuter", 5, 6) %>%
             group_rows("Large city", 7, 8) %>%
             group_rows("Provincial", 9, 10) %>%
             group_rows("Rural", 11, 12)


```




#### 12. Combined property and lot model augmented 

```{r Combined model augmented, eval=TRUE, echo=TRUE, warning=FALSE, results = 'hide', fig.width=20, fig.height=30}
standard_building_value <- 1e+6


#Create combined sales set with standardized properties on lot sales

standardize_property <- function(data, reference_data) {
  
    data %>% mutate(year_built = 1970,
                    year_rebuilt = 1970,
                    living_space = 140,
                    maximum_coverage_percentage = 30,
                    unused_floor_area = 0,
                    number_of_divisions = 1,
                    number_of_bathrooms = 1,
                    house_type = factor("Detached", levels = levels(factor(reference_data$house_type))),
                    roof_type = factor("Fiber cement", levels = levels(factor(reference_data$roof_type))),
                    wall_type = factor("Brick", levels = levels(factor(reference_data$wall_type))),
                    heating_type = factor("District heating", levels = levels(factor(reference_data$heating_type))))
  }

dataset_lot_sales <- standardize_property(dataset_lot_sales, dataset)

standard_building_value <- 1e6

dataset_lot_sales <- dataset_lot_sales %>%
                     mutate(sales_price_observed = sales_price,
                            standard_building_value_added = standard_building_value,
                            sales_price = sales_price + standard_building_value_added)

dataset_property_sales <- dataset_property_sales %>%
                          mutate(sales_price_observed = sales_price,
                                 standard_building_value_added = 0)

dataset_sales <- bind_rows(dataset_property_sales, dataset_lot_sales)
dataset_sales$type <- factor(dataset_sales$type, levels = c("Property sale", "Lot sale"))

#assign folds
assign_cv_folds <- function(data, N, seed = 123) {
  set.seed(seed)
  data %>% mutate(validation_set = sample(rep(1:N, length.out = n())),
                  .rand = runif(n()))
}

dataset_sales_cv <- assign_cv_folds(dataset_sales, N = N, seed = 123)

table(dataset_sales_cv$validation_set, useNA = "ifany")


#fit models
fit_fold <- function(i, data, k_spatial, k_other = 10, fraction = 1.0,
                     checkpoint_prefix = "Model", save_checkpoint = TRUE,
                     seed = 123) {
  
  training_subset <- data %>% filter(validation_set != i)

  if (fraction < 1.0) {
    
    training_subset <- training_subset %>%
                       group_by(municipality_type) %>%
                       arrange(.rand, .by_group = TRUE) %>%
                       slice_head(prop = fraction) %>%
                       ungroup() 
  }
  
  model <- bam(sales_price ~ 
               s(coordinate_east, coordinate_north, k = k_spatial, pc = c(697000, 6150000)) +
               s(sales_date_numeric, k = k_other, by = municipality_type, pc = 2024) +
               s(year_built, k = k_other, by = municipality_type, pc = 1970) +
               s(year_rebuilt, k = k_other, by = municipality_type, pc = 1970) +
               s(living_space, k = k_other, by = municipality_type, pc = 140) +
               s(lot_size, k = k_other, by = municipality_type, pc = 800) +
               s(coast_distance, k = k_other, pc = 1500) +
               s(lake_distance, k = k_other, pc = 500) +
               s(stream_distance, k = k_other, pc = 1000) +
               s(road_distance, k = k_other, pc = 2000) +
               s(highway_distance, k = k_other, pc = 2000) +
               s(train_station_distance, k = k_other, pc = 2000) +
               s(railway_distance, k = k_other, pc = 1000) +
               s(windturbine_distance, k = k_other, pc = 3000) +
               s(ocean_view, k = k_other, pc = 0) +
               s(lake_view, k = k_other, pc = 0) +
               s(forest_size, k = k_other, pc = 0) +
               s(number_of_bathrooms, by = type, k = 2, pc = 1) +
               house_type * municipality_type +
               roof_type * municipality_type +
               wall_type * municipality_type +
               heating_type * municipality_type +
               type * municipality_type,
               data = training_subset,
               chunk.size = nrow(training_subset),
               family = tw(link = log),
               method = "fREML")
  
  if (save_checkpoint) {
    filename <- sprintf("%s_fold%d_k%d_frac%.2f.rds", checkpoint_prefix, i, k_spatial, fraction)
    saveRDS(model, file = file.path("checkpoints_lot_sales", filename))
  }
  
  model
}

fit_model_ensemble <- function(data, N, k_spatial, k_other = 10, fraction = 1.0, checkpoint_prefix = "Model") {
  
  models <- future_lapply(1:N, function(i) {
    fit_fold(
      i = i,
      data = data,
      k_spatial = k_spatial,
      k_other = k_other,
      fraction = fraction,
      checkpoint_prefix = checkpoint_prefix
    )
  })
  names(models) <- paste0("Model", 1:N)
  models
}

if (load) {
  models <- readRDS("saved_files_lot_sales/models_combined_augmented.rds")
} else {
  models <- fit_model_ensemble(
    data = dataset_sales_cv,
    N = N,
    k_spatial = k_spatial,
    k_other = k_other,
    fraction = 1.0,
    checkpoint_prefix = "main"
  )
  saveRDS(models, "saved_files_lot_sales/models_combined_augmented.rds")
}

```



```{r Model performance 2, eval=TRUE, echo=TRUE, warning=FALSE}

get_residuals <- function(models, dataset, fold_from_name = TRUE) {
  
  residual_list <- future_lapply(seq_along(models), function(i) {
    
    model_name <- names(models)[i]
    model <- models[[i]]
    
    fold <- if (fold_from_name) {
      as.integer(sub("^Model(\\d+).*$", "\\1", model_name))
    } else {
      i
    }
    
    predicted_model_price <- as.vector(
      predict.gam(model, newdata = dataset, type = "response")
    )
    
    predicted_observed_price <- predicted_model_price - dataset$standard_building_value_added
    
    observed_price <- dataset$sales_price_observed
    
    tibble(sales_price = observed_price,
           sales_price_model = dataset$sales_price,
           predicted_sales_price = predicted_observed_price,
           predicted_model_price = predicted_model_price,
           residual_sales_price = predicted_observed_price - observed_price,
           residual_sales_price_percent = (predicted_observed_price - observed_price) / observed_price * 100,
           residual_sales_price_log = log(predicted_observed_price) - log(observed_price),
           ratio = predicted_observed_price / observed_price,
           validation = dataset$validation_set == fold,
           id = dataset$id,
           type = dataset$type,
           municipality_type = dataset$municipality_type,
           Model = model_name)
  })
  
  bind_rows(residual_list)
}

residual_df <- get_residuals(models, dataset_sales_cv)
saveRDS(residual_df, "saved_files_lot_sales//residual_df_combined_augmented.rds")

performance_countrywide <- residual_df %>%
                           mutate(dataset_type = if_else(validation, "Validation", "Training")) %>%
                           group_by(dataset_type, type) %>%
                           summarise(municipality_type = "Countrywide",
                                     MALE = mean(abs(residual_sales_price_log)),
                                     MAPE = mean(abs(residual_sales_price_percent)),
                                     MAE = mean(abs(residual_sales_price)),
                                     PM20 = mean(abs(residual_sales_price_percent) <= 20) * 100,
                                     COD = mean(abs(ratio - median(ratio))) / median(ratio) * 100,
                                     PRD = mean(ratio) / (sum(predicted_sales_price) / sum(sales_price)),
                                     .groups = "drop")

performance_municipality_type <- residual_df %>%
                                 mutate(dataset_type = if_else(validation, "Validation", "Training")) %>%
                                 group_by(dataset_type, municipality_type) %>%
                                 summarise(MALE = mean(abs(residual_sales_price_log)),
                                           MAPE = mean(abs(residual_sales_price_percent)),
                                           MAE = mean(abs(residual_sales_price)),
                                           PM20 = mean(abs(residual_sales_price_percent) <= 20) * 100,
                                           COD = mean(abs(ratio - median(ratio))) / median(ratio) * 100,
                                           .groups = "drop")

performance <- bind_rows(performance_countrywide, performance_municipality_type)

municipality_levels <- c("Countrywide", sort(setdiff(unique(performance$municipality_type), "Countrywide")))

performance_sorted <- performance %>%
                      mutate(dataset_type = factor(dataset_type, levels = c("Training", "Validation")),
                             municipality_type = factor(municipality_type, levels = municipality_levels),
                             MAE = round(MAE,0),
                             PRD = ifelse(is.na(PRD), "", round(PRD, 2))) %>%
                      arrange(municipality_type, dataset_type)  %>% 
                      rename(Type = dataset_type,
                             `Municipality type` = municipality_type,
                             `MAPE [%]` = MAPE,
                             `MAE [DKK]` = MAE,
                             `PM20 [%]` = PM20,
                             `COD [%]` = COD)

#standard table
#knitr::kable(performance_sorted, digits = 2, caption = "Model Performance by Municipality Type")

#nicer table
kable(performance_sorted %>% 
select(-`Municipality type`), 
       digits = 2, 
       caption = "Model Performance by Municipality Type", 
       booktabs = TRUE) %>%
kable_styling(full_width = FALSE, position = "center") %>%
group_rows("Countrywide", 1, 2) %>%
group_rows("Capital", 3, 4) %>%
group_rows("Commuter", 5, 6) %>%
group_rows("Large city", 7, 8) %>%
group_rows("Provincial", 9, 10) %>%
group_rows("Rural", 11, 12)

tbl_latex <- kable(performance_sorted %>% select(-`Municipality type`),
                   format = "latex", booktabs = TRUE, digits = 2,
                   caption = "Model performance by municipality type.",
                   label = "perf_by_muni") %>%
             kable_styling(full_width = FALSE, 
                           position = "center", 
                           latex_options = c("hold_position")) %>%
             group_rows("Countrywide", 1, 2) %>%
             group_rows("Capital", 3, 4) %>%
             group_rows("Commuter", 5, 6) %>%
             group_rows("Large city", 7, 8) %>%
             group_rows("Provincial", 9, 10) %>%
             group_rows("Rural", 11, 12)


```


