---
title: "Supplementary Notebook"
author: "August E. G. Mikkelsen"
date: "2025-05-08"
output: pdf_document
header-includes:
  - \usepackage{xcolor}
---

```{r setup, eval=TRUE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggpubr)
library(mgcv)
library(gratia) 
library(knitr)
library(patchwork)
library(lubridate)
library(future)
library(future.apply)
library(rlang)
library(stringr)
library(furrr)
library(xgboost)
library(caret)
library(Hmisc)
library(kableExtra)
library(nabor)
library(purrr)
library(ggrastr)
library(ggridges)
library(gridExtra)
library(forcats)
library(pheatmap)
library(viridis)
library(cowplot)
library(scico)
library(scales)

# Global parameters
set.seed(123) # for reproducibility
N <- 4 # number of CV folds
k_spatial = 500 # initial number of basis functions used in the spatial smooths
k_other = 10 # initial number of basis functions used for other smooths
load <- FALSE # determines whether or not to preload models in the various chunks
plan(multisession, workers = N) #do as much as possible in parallel
options(future.globals.maxSize = 10 * 1024^3) #increase maxSize for future to handle large models

#External files
coordinate_grid <- read.csv2("coordinates_ocean_land.csv")

#Create a saved files folder and checkpoints folder is these do not already exist
for (f in c("saved_files", "checkpoints")) {
  if (!dir.exists(f)) {
    dir.create(f, recursive = TRUE)
    message(sprintf("Created folder: %s", f))
  } else {
    message(sprintf("Folder already exists: %s", f))
  }
}
```

# Introduction
This notebook provides a walkthrough of the analysis and results described in the paper "Generalized Additive Models for Mass Appraisal: Evidence from the Danish Public Property Valuation System". 


# Dataset 

## 1. Description
The analysis is based on a dataset of single family home sales, single family lot sales and properties in Denmark from the period 01/01/2014-01/01/2024. The dataset, included as a .rds file in this repository, is loaded and inspected below and in the subsequent cells histograms of selected variables are produced. 

```{r Load dataset, eval=TRUE, echo=TRUE, warning=FALSE, fig.width=20, fig.height=30}

dataset <- readRDS("dataset.rds")
str(dataset)
#The three types included in the dataset. This notebook only uses Property and Property Sales
table(dataset$type)

```

```{r Continuous variables, eval=TRUE, echo=TRUE, warning=FALSE, fig.width=20, fig.height=30}

get_label <- function(var, labels_vec) {
  lbl <- tryCatch(labels_vec[[var]], error = function(e) NULL)
  if (is.null(lbl) || is.na(lbl)) var else lbl
}

plot_one <- function(df, var, label) {
  
  x <- df[[var]]
  
  if (label == "Sales date") {
    x_date <- as.Date(x)
    x <- decimal_date(x_date)
  }
  
  na_count <- sum(is.na(x))
  zero_count <- sum(x == 0, na.rm = TRUE)
  na_label <- paste0("Number of NA-values: ", format(na_count, big.mark = ",", scientific = FALSE))
  zero_label <- paste0("Number of zero-values: ", format(zero_count, big.mark = ",", scientific = FALSE))

  if (is.numeric(x)) {
    
    df2 <- tibble(x = x) %>% filter(is.finite(x) & x != 0)

    rng <- range(df2$x, na.rm = TRUE)
    
    rng_label <- paste0("Range: [", format(rng[1], big.mark = "", scientific = FALSE, digits = 2, trim = TRUE), " , ",
                                    format(rng[2], big.mark = "", scientific = FALSE, digits = 2, trim = TRUE), "]")

    x_pos = (0.75*min(x, na.rm = TRUE) + 0.25*max(x, na.rm = TRUE))
    
    ggplot(df2, aes(x = x)) +
      geom_histogram(bins = 30, fill = "#3182bd", color = "white") +
      labs(x = label, y = NULL) +
      annotate("text", x = x_pos, y = Inf, label = na_label, hjust = 0, vjust = 0, size = 4.5, color = "grey25", fontface = "bold") +
      annotate("text", x = x_pos, y = Inf, label = zero_label, hjust = 0, vjust = 1.5, size = 4.5, color = "grey25", fontface = "bold") +
      annotate("text", x = x_pos, y = Inf, label = rng_label, hjust = 0, vjust = 3, size = 4.5, color = "grey25", fontface = "bold") +
      coord_cartesian(clip = "off") +
      theme(text = element_text(size = 12),
            plot.margin = margin(t = 28, r = 12, b = 5, l = 5),
            panel.grid.minor = element_blank(),
            panel.grid.major.x = element_blank())

  } else {
    
    x2 <- fct_lump_n(factor(x), n = 30, other_level = "Other")
    
    df_counts <- tibble(x = x2) %>%
                 filter(!is.na(x)) %>%
                 count(x, name = "n")
    
    x_pos = (nlevels(df_counts$x) + 1) / 2
    
    ggplot(df_counts, aes(x = x, y = n)) +
      geom_col(aes(fill = x), color = "white", show.legend = FALSE) +
      scale_fill_brewer(palette = "Set3") +
      geom_text(aes(label = comma(n)), vjust = -0.2, size = 3.3, color = "grey15") +
      labs(x = NULL, y = NULL) +
      ggtitle(label) + 
      scale_y_continuous(expand = expansion(mult = c(0.02, 0.12))) +
      coord_cartesian(clip = "off") +
      theme(plot.margin = margin(t = 20, r = 12, b = 5, l = 5),
            axis.text.x = element_text(angle = 45, hjust = 1),
            panel.grid.minor = element_blank(),
            panel.grid.major.x = element_blank())}
}


variables_continuous <- c(
  sales_price = "Sales price [DKK]",
  year_built = "Year built",
  year_rebuilt = "Year rebuilt",
  living_space = "Living space [m²]",
  lot_size = "Lot size [m²]",
  coast_distance = "Distance to coast [m]",
  lake_distance = "Distance to lake [m]",
  stream_distance = "Distance to stream [m]",
  road_distance = "Distance to road [m]",
  highway_distance = "Distance to highway [m]",
  train_station_distance = "Distance to train station [m]",
  railway_distance = "Distance to railway [m]",
  windturbine_distance = "Distance to wind turbine [m]",
  ocean_view = "Ocean view [m]",
  lake_view = "Lake view [m]",
  forest_size = "Forest size [ha]",
  number_of_bathrooms = "Number of bathrooms",
  number_of_rooms = "Number of rooms",
  number_of_floors = "Number of floors",
  sales_date = "Sales date")

vars_cont <- intersect(names(variables_continuous), names(dataset))

histograms_cont <- map(vars_cont, ~ plot_one(dataset, .x, get_label(.x, variables_continuous)))

saveRDS(histograms_cont, "saved_files/histograms_continuous.rds")
wrap_plots(histograms_cont, ncol = 3)
```

```{r Categorical variables, eval=TRUE, echo=TRUE, warning=FALSE, fig.width=20, fig.height=30}

variables_categorical <- c(
  house_type = "House type",
  roof_type = "Roof type",
  wall_type = "Wall type",
  heating_type = "Heating type",
  municipality_type = "Municipality type",
  province = "Province",
  region = "Region",
  zone = "Zone"
)

vars_cat <- intersect(names(variables_categorical), names(dataset))

histograms_cat <- map(vars_cat, ~ plot_one(dataset, .x, get_label(.x, variables_categorical)))

saveRDS(histograms_cat, "saved_files/histograms_categorical.rds")

wrap_plots(histograms_cat, ncol = 3)

```


## 2. Variable conversions and imputation
The analysis relies on a set of variable conversions performed in order to:

- Handle missing values in the dataset
- Convert categorical variables to factors
- Convert sales dates to numeric values

The two first are straightforward while the treatment of missing values requires more consideration. The analysis presented in the paper is based on the following simply imputation scheme: 

- year_rebuilt: For this variable all NA's as well as values less than 1970 are replaced by 1970. 
- distance-variables: For these variables all NA's are replaced by the maximum value of the respective variable. 

The code for the above transformations is provided below. 

```{r Variable conversion and imputation, eval=TRUE, echo=TRUE, warning=FALSE, results= 'hide'}

#convert relevant categorical variables to factors with chosen references
dataset$municipality_type <- relevel(factor(dataset$municipality_type), ref = "Provincial")
dataset$house_type <- relevel(factor(dataset$house_type), ref = "Detached")
dataset$roof_type <- relevel(factor(dataset$roof_type), ref = "Fiber cement")
dataset$wall_type <- relevel(factor(dataset$wall_type), ref = "Brick")
dataset$heating_type <- relevel(factor(dataset$heating_type), ref = "District heating")

#create a numerical variable for sales_date
dataset <- dataset %>% mutate(sales_date = as.Date(sales_date),
                              sales_date_numeric = decimal_date(sales_date))

#impute NA-values
dataset <- dataset %>%
  mutate(year_rebuilt = if_else(is.na(year_rebuilt) | year_rebuilt < 1970, 1970, year_rebuilt))

dataset <- dataset %>% mutate(across(c(lake_distance,
                                       stream_distance,
                                       coast_distance,
                                       road_distance,
                                       highway_distance,
                                       train_station_distance,
                                       railway_distance,
                                       windturbine_distance),
                                     ~ replace_na(., max(., na.rm = TRUE))))

#check that there are no remaining NAs in sales data
which(is.na(subset(dataset, type == "Property sale")), arr.ind = TRUE)

```

## 3. Cross-Validation 
A simple 4-fold cross-validation split is implemented in the code below. 

```{r Test set and cross-validation, eval=TRUE, echo=TRUE, warning=FALSE, results = 'hide'}
# pick out sales and properties
dataset_sales <- dataset %>% filter(type == "Property sale")
dataset_properties <- dataset %>% 
                      filter(type == "Property") %>% 
                      mutate(sales_date_numeric = 2024.0) #we want to value properties 1 January 2024

assign_cv_folds <- function(data, N, seed = 123) {
  set.seed(seed)
  data %>% mutate(validation_set = sample(rep(1:N, length.out = n())),
                  .rand = runif(n()))
}

dataset_sales_cv <- assign_cv_folds(dataset_sales, N = N, seed = 123)

table(dataset_sales_cv$validation_set, useNA = "ifany")

```
# GAM Models 

## 1. Model training
The code below fits the GAM model structure described in "Generalized Additive Models for Real Estate Valuation: A Case Study of the Danish Housing Market" based on the cross validation split employed. 

```{r Model training, eval=TRUE, echo=TRUE, warning=FALSE, results='hide'}
fit_fold <- function(i, data, k_spatial, k_other = 10, fraction = 1.0,
                     checkpoint_prefix = "Model", save_checkpoint = TRUE,
                     seed = 123) {
  
  training_subset <- data %>% filter(validation_set != i)

  if (fraction < 1.0) {
    
    training_subset <- training_subset %>%
                       group_by(municipality_type) %>%
                       arrange(.rand, .by_group = TRUE) %>%
                       slice_head(prop = fraction) %>%
                       ungroup() 
  }
  
  model <- bam(sales_price ~ 
               s(coordinate_east, coordinate_north, k = k_spatial, pc = c(697000, 6150000)) +
               s(sales_date_numeric, k = k_other, by = municipality_type, pc = 2024) +
               s(year_built, k = k_other, by = municipality_type, pc = 1970) +
               s(year_rebuilt, k = k_other, by = municipality_type, pc = 1970) +
               s(living_space, k = k_other, by = municipality_type, pc = 140) +
               s(lot_size, k = k_other, by = municipality_type, pc = 800) +
               s(coast_distance, k = k_other, pc = 1500) +
               s(lake_distance, k = k_other, pc = 500) +
               s(stream_distance, k = k_other, pc = 1000) +
               s(road_distance, k = k_other, pc = 2000) +
               s(highway_distance, k = k_other, pc = 2000) +
               s(train_station_distance, k = k_other, pc = 2000) +
               s(railway_distance, k = k_other, pc = 1000) +
               s(windturbine_distance, k = k_other, pc = 3000) +
               s(ocean_view, k = k_other, pc = 0) +
               s(lake_view, k = k_other, pc = 0) +
               s(forest_size, k = k_other, pc = 0) +
               s(number_of_bathrooms, by = municipality_type, k = 2, pc = 1) +
               house_type * municipality_type +
               roof_type * municipality_type +
               wall_type * municipality_type +
               heating_type * municipality_type,
               data = training_subset,
               chunk.size = nrow(training_subset),
               family = tw(link = log),
               method = "fREML")
  
  if (save_checkpoint) {
    filename <- sprintf("%s_fold%d_k%d_frac%.2f.rds", checkpoint_prefix, i, k_spatial, fraction)
    saveRDS(model, file = file.path("checkpoints", filename))
  }
  
  model
}

fit_model_ensemble <- function(data, N, k_spatial, k_other = 10, fraction = 1.0, checkpoint_prefix = "Model") {
  
  models <- future_lapply(1:N, function(i) {
    fit_fold(
      i = i,
      data = data,
      k_spatial = k_spatial,
      k_other = k_other,
      fraction = fraction,
      checkpoint_prefix = checkpoint_prefix
    )
  })
  names(models) <- paste0("Model", 1:N)
  models
}

if (load) {
  models <- readRDS("saved_files/models.rds")
} else {
  models <- fit_model_ensemble(
    data = dataset_sales_cv,
    N = N,
    k_spatial = k_spatial,
    k_other = k_other,
    fraction = 1.0,
    checkpoint_prefix = "main"
  )
  saveRDS(models, "saved_files/models.rds")
}

```


## 2. Partial effects

#### Spatial smooth 
The code below extracts the spatial smooth for each model and plots it alongside the log-scale standard error. The plotting relies on a premade spatial grid of Denmark, where ocean and land are marked. 

```{r Spatial smooth, eval=TRUE, echo=TRUE, warning=FALSE, results = 'hide', fig.width=20, fig.height=30}


# 1a) Predict once per model and store results
prediction_results <- future_lapply(seq_along(models), function(i) {
  model <- models[[i]]
  model_data <- model.frame(model)
  grid <- coordinate_grid

  for (var in setdiff(all.vars(formula(model)), names(grid))) {
    x <- model_data[[var]]
    if (is.factor(x)) {
      grid[[var]] <- factor(rep(levels(x)[1], nrow(grid)), levels = levels(x))
    } else {
      grid[[var]] <- 0
    }
  }

  terms <- predict.gam(model, grid, type = "terms", se.fit = TRUE)
  df_predictions <- as.data.frame(terms$fit)
  df_se <- as.data.frame(terms$se.fit)

  effect <- df_predictions$`s(coordinate_east,coordinate_north)`
  se <- df_se$`s(coordinate_east,coordinate_north)`
  effect[grid$ocean] <- NA
  se[grid$ocean] <- NA

  grid$effect_coordinates <- effect
  grid$se_coordinates <- se

  list(model_name = names(models)[i],
       grid = grid)
})


# 1b) Compute SD across models (per grid cell) ---
effects_mat <- do.call(cbind, lapply(prediction_results, function(x) x$grid$effect_coordinates))
base_grid   <- prediction_results[[1]]$grid

se_across_models <- apply(effects_mat, 1, sd, na.rm = TRUE)
se_across_models[base_grid$ocean] <- NA

# 2. Global ranges (normalized across everything) ---
# Effects: symmetric across *all model* effects
all_effects <- unlist(lapply(prediction_results, function(x) x$grid$effect_coordinates))
max_abs_effect <- max(abs(all_effects), na.rm = TRUE)
effect_limits  <- c(-max_abs_effect, max_abs_effect)

# SEs: include per-model SEs AND the across-models SD
all_ses_models <- unlist(lapply(prediction_results, function(x) x$grid$se_coordinates))
se_limits <- c(0, max(c(all_ses_models, se_across_models), na.rm = TRUE))  # SEs are >= 0

# 3. Plotting function (adds land/ocean outline) ---
plot_model_terms_from_grid <- function(grid, model_name, effect_range, se_range) {

  plot_effect <- ggplot(data = grid) +
    geom_tile(aes(coordinate_east, coordinate_north, fill = effect_coordinates)) +
    labs(x = "", y = "") +
    scale_fill_scico(
      palette = "vik",
      name = paste("Partial effect", sub("^Model", "Model ", model_name)),
      na.value = "white",
      limits = effect_range
    ) +
    geom_contour(
      aes(x = coordinate_east, y = coordinate_north, z = as.numeric(!ocean)),
      color = "black", size = 0.4
    ) +
    coord_fixed() +
    theme_void() +
    theme(
      text = element_text(size = 12),
      legend.position = c(0.85, 0.75),
      legend.title = element_text(size = 9, margin = margin(b = 5)),
      legend.text = element_text(size = 9),
      legend.title.align = 0.5
    )
  
  model_label <- sub("^Model", "Model ", model_name)

  plot_se <- ggplot(data = grid) +
    geom_tile(aes(coordinate_east, coordinate_north, fill = se_coordinates)) +
    labs(x = "", y = "") +
    scale_fill_scico(
    palette = "vik",
    name = bquote(sigma[.(model_label)]),
    na.value = "white",
    limits = se_range) +
    geom_contour(aes(x = coordinate_east, y = coordinate_north, z = as.numeric(!ocean)),
                 color = "black", 
                 size = 0.4) +
    coord_fixed() +
    theme_void() +
    theme(
      text = element_text(size = 12),
      legend.position = c(0.85, 0.75),
      legend.title = element_text(size = 9, margin = margin(b = 5)),
      legend.text = element_text(size = 9),
      legend.title.align = 0.5)

  list(effect = plot_effect, se = plot_se)
}

# 4. Build plots for each model (effect + SE) ---
plot_lists <- lapply(prediction_results, function(result) {
  plot_model_terms_from_grid(
    result$grid, result$model_name,
    effect_range = effect_limits,
    se_range     = se_limits
  )
})

plot_spatial_smooths <- unlist(plot_lists, recursive = FALSE)

# 5. ONE extra plot: SD of effects across models (uses same se_limits) ---
base_grid2 <- base_grid
base_grid2$se_across_models <- se_across_models  # <- bake it into the data

plot_sd_across <- ggplot(data = base_grid2) +
  geom_tile(aes(coordinate_east, coordinate_north, fill = se_across_models)) +
  labs(x = "", y = "") +
  scale_fill_scico(palette = "vik",
                   name = expression(sigma[Ensemble]),
                   na.value = "white",
                   limits = se_limits) +
  geom_contour(
    aes(x = coordinate_east, y = coordinate_north, z = as.numeric(!ocean)),
    color = "black", size = 0.4
  ) +
  coord_fixed() +
  theme_void() +
  theme(
    text = element_text(size = 12),
    legend.position = c(0.85, 0.75),
    legend.title = element_text(size = 9, margin = margin(b = 5)),
    legend.text = element_text(size = 9),
    legend.title.align = 0.5
  )

plot_spatial_smooths <- c(plot_spatial_smooths, list(plot_sd_across))
saveRDS(plot_spatial_smooths, "saved_files/plot_spatial_smooths.rds")
wrap_plots(plot_spatial_smooths, ncol = 3)

```


#### Other smooth effects 
The code below extracts and plots the remaining smooth effects of the trained GAM-models. A smooth corresponding to each trained model is plotted, but for readability confidence intervals are only displayed for Model 1. 

```{r Other smooth effects, eval=TRUE, echo=TRUE, warning=FALSE, results = 'hide', fig.width=20, fig.height=30}

smooth_variables <- c(
  sales_date_numeric = "Sales date",
  year_built = "Year built",
  year_rebuilt = "Year rebuilt",
  living_space = "Living space [m²]",
  lot_size = "Lot size [m²]",
  coast_distance = "Distance to coast [m]",
  lake_distance = "Distance to lake [m]",
  stream_distance = "Distance to stream [m]",
  road_distance = "Distance to road [m]",
  highway_distance = "Distance to highway [m]",
  train_station_distance = "Distance to train station [m]",
  railway_distance = "Distance to railway [m]",
  windturbine_distance = "Distance to wind turbine [m]",
  ocean_view = "Ocean view [m]",
  lake_view = "Lake view [m]",
  forest_size = "Forest size [ha]",
  number_of_bathrooms = "Number of bathrooms")

plot_smooths <- list()

for (var in names(smooth_variables)) {

  smooth_dfs <- lapply(names(models), function(model_name) {
    mod <- models[[model_name]]
    df <- tryCatch(
      smooth_estimates(mod, smooth = paste0("s(", var, ")"), partial_match = TRUE),
      error = function(e) {
        message(sprintf("Skipping '%s' in %s: %s", var, model_name, e$message))
        return(NULL)
      })
    if (!is.null(df)) {
      df$Model <- sub("^Model", "Model ", model_name)
    }
    df
  })

  smooth_dfs <- Filter(Negate(is.null), smooth_dfs)  # remove NULLs
  if (length(smooth_dfs) == 0) next

  combined_df <- bind_rows(smooth_dfs)

  has_by <- any(!is.na(combined_df$.by))
  plot_title <- paste("Partial effect of", gsub("_", " ", var))

  if (has_by) {

    by_var_name <- unique(na.omit(combined_df$.by))[[1]]

    p <- ggplot(combined_df, aes(x = .data[[var]], y = .estimate,
                                 color = .data[[by_var_name]],
                                 linetype = Model, 
                                 group = interaction(Model, .data[[by_var_name]]))) +
        geom_ribbon(data = subset(combined_df, Model == "Model 1"),
                    aes(ymin = .estimate - 1.96*.se, ymax = .estimate + 1.96*.se),
                    alpha = 0.5, 
                    color = NA) +
        geom_line(linewidth = 1) +
        labs(title = plot_title,
             x = smooth_variables[[var]],
             y = "Partial effect") +
        scale_color_brewer(palette = "Dark2", name = "Municipality type") +
        scale_linetype_manual(values = c("solid", 
                                         "dashed", 
                                         "dotted", 
                                         "dotdash", 
                                         "twodash", 
                                         "longdash")) +
        guides(fill = guide_legend(override.aes = list(linetype = 0)))

  } else {

    p <- ggplot(combined_df, aes(x = .data[[var]], y = .estimate, linetype = Model)) +
         geom_ribbon(data = subset(combined_df, Model == "Model 1"),
                     aes(ymin = .estimate - 1.96*.se, ymax = .estimate + 1.96*.se),
                     alpha = 0.5, 
                     color = NA) +
         geom_line(linewidth = 1, color = "black") +
         labs(title = plot_title,
              x = smooth_variables[[var]],
              y = "Partial effect") +
         scale_linetype_manual(values = c("solid", 
                                          "dashed", 
                                          "dotted", 
                                          "dotdash", 
                                          "twodash", 
                                          "longdash"))
  }
  
  p <- p + theme(legend.position = "none")

  plot_smooths[[var]] <- p
}


saveRDS(plot_smooths, "saved_files/plot_smooths.rds")

#manual axis specifications of some ill-behaved smooths to make plots easier to understand
plot_smooths$stream_distance <- plot_smooths$stream_distance + ylim(-0.03,0.1)
plot_smooths$lake_view <- plot_smooths$lake_view + xlim(0.0,10000) + ylim(0, 0.6)
plot_smooths$forest_size <- plot_smooths$forest_size + xlim(0.0,40) + ylim(-0.02, 0.1)

wrap_plots(plot_smooths, ncol = 2) + 
           plot_layout(guides = "collect") &
           theme(legend.position = "bottom",
                 legend.text = element_text(size = 12),
                 legend.title = element_text(size = 13, face = "bold"),
                 legend.box = "horizontal",
                 legend.justification = "center")
```

#### Factor effects
The code below extracts and plots the factor effects of trained GAM-models. A point is shown for the effect of each model but error bars are shown only for Model 1. 

```{r Factor effects, eval=TRUE, echo=TRUE, warning=FALSE, results = 'hide', fig.width=20, fig.height=20}
factor_variables <- c(
  house_type = "House type",
  municipality_type = "Municipality type",
  roof_type = "Roof type",
  wall_type = "Wall type",
  heating_type = "Heating type")

make_termwise_df <- function(model, model_data, variable, interaction_by = "municipality_type") {

  v_vals <- unique(model_data[[variable]])
  if (is.null(interaction_by)) {
    df <- data.frame(setNames(list(v_vals), variable), stringsAsFactors = FALSE)
  } else {
    m_vals <- unique(model_data[[interaction_by]])
    grid_list <- setNames(list(v_vals, m_vals), c(variable, interaction_by))
    df <- expand.grid(grid_list, stringsAsFactors = FALSE)
    df[[interaction_by]] <- factor(df[[interaction_by]], 
                                   levels = levels(model_data[[interaction_by]]))
  }
  for (var in setdiff(all.vars(formula(model)), names(df))) {
    x <- model_data[[var]]
    if (is.factor(x)) {
      df[[var]] <- factor(rep(levels(x)[1], nrow(df)), levels = levels(x))
    } else {
      df[[var]] <- 0
    }
  }
  return(df)
}

combine_all_term_effects <- function(df, term, interaction_by = "municipality_type") {
  pattern <- paste0("^", term, "$|^", term, ":" ,"|:", term, "$")
  cols_to_sum <- grep(pattern, names(df), value = TRUE)
  df[[paste0("effect_", term)]] <- rowSums(as.matrix(df[cols_to_sum]))
  return(df)
}

combine_all_term_se <- function(df, term, interaction_by = "municipality_type") {
  pattern <- paste0("^", term, "$|^", term, ":" ,"|:", term, "$")
  cols_to_sum <- grep(pattern, names(df), value = TRUE)
  df[[paste0("se_", term)]] <- sqrt(rowSums(as.matrix(df[cols_to_sum])^2))
  return(df)
}

plot_term_effect <- function(df, term, interaction_by = "municipality_type") {
  effect_col <- paste0("effect_", term)
  lower_col <- paste0("lower_", effect_col)
  upper_col <- paste0("upper_", effect_col)
  df_sub <- df %>% filter(term == !!term)
  plot_title <- paste("Partial effect of", gsub("_", " ", term))

  p <- ggplot(df_sub, aes(x = .data[[term]], 
                          y = .data[[effect_col]], 
                          shape = Model,
                          group = interaction(Model, if (identical(term, interaction_by) || 
                                                        !(interaction_by %in% names(df_sub))) NULL 
                                                     else .data[[interaction_by]]))) +
      geom_pointrange(data = subset(df_sub, Model == "Model 1"),
                      aes(ymin = .data[[lower_col]], ymax = .data[[upper_col]]),
                      position = position_dodge(width = 0.5),
                      size = 0.8) +
      geom_point(data = subset(df_sub, Model != "Model 1"),
                 position = position_dodge(width = 0.5),
                 size = 2.5) +
      coord_flip() +
      labs(x = NULL, y = "Partial effect", title = plot_title, shape = "Model") +
      theme(axis.text.y = element_text(angle = 45, hjust = 1))

  if (interaction_by %in% names(df_sub)) {
      color_var <- if (term == interaction_by) term else interaction_by
      p <- p + aes(color = .data[[color_var]]) +
      scale_color_brewer(palette = "Dark2", name = "Municipality type")
  }

  return(p)
}

factor_dfs <- list()

for (model_name in names(models)) {

  model <- models[[model_name]]
  model_data <- model.frame(model)
  relevant_factors <- factor_variables[names(factor_variables) %in% all.vars(formula(model))]

  for (var in names(relevant_factors)) {

    interaction <- if (var == "municipality_type") NULL else "municipality_type"
    df <- make_termwise_df(model, model_data, var, interaction_by = interaction)
    df$term <- var

    terms <- predict.gam(model, df, type = "terms", se.fit = TRUE)
    df_predictions <- as.data.frame(terms$fit)
    df_se <- as.data.frame(terms$se.fit)

    combined_predictions <- combine_all_term_effects(df_predictions, 
                                                     var, 
                                                     interaction_by = interaction) %>% 
                            select(starts_with("effect_"))
    
    combined_se <- combine_all_term_se(df_se, 
                                       var, 
                                       interaction_by = interaction) %>% 
                   select(starts_with("se_"))

    df_final <- df %>% select(all_of(c(var, interaction)), term) %>%
                       bind_cols(combined_predictions) %>%
                       bind_cols(combined_se) %>%
                       mutate(across(starts_with("effect_"),
                                     ~ . - 1.96 * get(str_replace(cur_column(), 
                                     "effect_", 
                                     "se_")),
                                     .names = "lower_{.col}")) %>%
                       mutate(across(starts_with("effect_"),
                                     ~ . + 1.96 * get(str_replace(cur_column(), "effect_", "se_")),
                                     .names = "upper_{.col}"))

    df_final$Model <- sub("^Model", "Model ", model_name)
    factor_dfs[[length(factor_dfs) + 1]] <- df_final
  }
}

df_all_factors <- bind_rows(factor_dfs)
plot_factors <- lapply(names(factor_variables), 
                       function(term) {plot_term_effect(df_all_factors, term)})
names(plot_factors) <- names(factor_variables)

saveRDS(plot_factors, "saved_files/plot_factors.rds")

wrap_plots(plot_factors, ncol = 2) +
           plot_layout(guides = "collect") &
           theme(legend.position = "bottom",
                 legend.text = element_text(size = 12),
                 legend.title = element_text(size = 13, face = "bold"),
                 legend.box = "horizontal",
                 legend.justification = "center")
```


## 3. Model performance
The code below extracts various residual types and calculates different performance measures for the training and validation set. Note that using N-fold cross-validation means that there are N-1 estimates for each training point, and the training error is simply taken as the average over these. 


```{r Model performance, eval=TRUE, echo=TRUE, warning=FALSE}

get_residuals <- function(models, dataset, fold_from_name = TRUE) {
  
  residual_list <- future_lapply(seq_along(models), function(i) {
    
    model_name <- names(models)[i]
    model <- models[[i]]
    
    if (fold_from_name) {
      fold <- as.integer(sub("^Model(\\d+).*$", "\\1", model_name))
    } else {
      fold <- i
    }
    
    predicted_sales_price <- predict(model, newdata = dataset, type = "response")
    
    tibble(sales_price = dataset$sales_price,
           sales_date_numeric = dataset$sales_date_numeric,
           predicted_sales_price = predicted_sales_price,
           residual_sales_price = predicted_sales_price - dataset$sales_price,
           residual_sales_price_percent = (predicted_sales_price - dataset$sales_price) / dataset$sales_price * 100,
           residual_sales_price_log = log(predicted_sales_price) - log(dataset$sales_price),
           ratio = predicted_sales_price / dataset$sales_price,
           validation = dataset$validation_set == fold,
           id = dataset$id,
           municipality_type = dataset$municipality_type,
           Model = model_name)
    
  })
  
  bind_rows(residual_list)
}

residual_df <- get_residuals(models, dataset_sales_cv)
saveRDS(residual_df, "saved_files/residual_df.rds")

performance_countrywide <- residual_df %>%
                           mutate(dataset_type = if_else(validation, "Validation", "Training")) %>%
                           group_by(dataset_type) %>%
                           summarise(municipality_type = "Countrywide",
                                     MALE = mean(abs(residual_sales_price_log)),
                                     MAPE = mean(abs(residual_sales_price_percent)),
                                     MAE = mean(abs(residual_sales_price)),
                                     PM20 = mean(abs(residual_sales_price_percent) <= 20) * 100,
                                     COD = mean(abs(ratio - median(ratio))) / median(ratio) * 100,
                                     PRD = mean(ratio) / (sum(predicted_sales_price) / sum(sales_price)),
                                     .groups = "drop")

performance_municipality_type <- residual_df %>%
                                 mutate(dataset_type = if_else(validation, "Validation", "Training")) %>%
                                 group_by(dataset_type, municipality_type) %>%
                                 summarise(MALE = mean(abs(residual_sales_price_log)),
                                           MAPE = mean(abs(residual_sales_price_percent)),
                                           MAE = mean(abs(residual_sales_price)),
                                           PM20 = mean(abs(residual_sales_price_percent) <= 20) * 100,
                                           COD = mean(abs(ratio - median(ratio))) / median(ratio) * 100,
                                           .groups = "drop")

performance <- bind_rows(performance_countrywide, performance_municipality_type)

municipality_levels <- c("Countrywide", sort(setdiff(unique(performance$municipality_type), "Countrywide")))

performance_sorted <- performance %>%
                      mutate(dataset_type = factor(dataset_type, levels = c("Training", "Validation")),
                             municipality_type = factor(municipality_type, levels = municipality_levels),
                             MAE = round(MAE,0),
                             PRD = ifelse(is.na(PRD), "", round(PRD, 2))) %>%
                      arrange(municipality_type, dataset_type)  %>% 
                      rename(Type = dataset_type,
                             `Municipality type` = municipality_type,
                             `MAPE [%]` = MAPE,
                             `MAE [DKK]` = MAE,
                             `PM20 [%]` = PM20,
                             `COD [%]` = COD)

#standard table
#knitr::kable(performance_sorted, digits = 2, caption = "Model Performance by Municipality Type")

#nicer table
kable(performance_sorted %>% 
select(-`Municipality type`), 
       digits = 2, 
       caption = "Model Performance by Municipality Type", 
       booktabs = TRUE) %>%
kable_styling(full_width = FALSE, position = "center") %>%
group_rows("Countrywide", 1, 2) %>%
group_rows("Capital", 3, 4) %>%
group_rows("Commuter", 5, 6) %>%
group_rows("Large city", 7, 8) %>%
group_rows("Provincial", 9, 10) %>%
group_rows("Rural", 11, 12)

tbl_latex <- kable(performance_sorted %>% select(-`Municipality type`),
                   format = "latex", booktabs = TRUE, digits = 2,
                   caption = "Model performance by municipality type.",
                   label = "perf_by_muni") %>%
             kable_styling(full_width = FALSE, 
                           position = "center", 
                           latex_options = c("hold_position")) %>%
             group_rows("Countrywide", 1, 2) %>%
             group_rows("Capital", 3, 4) %>%
             group_rows("Commuter", 5, 6) %>%
             group_rows("Large city", 7, 8) %>%
             group_rows("Provincial", 9, 10) %>%
             group_rows("Rural", 11, 12)

writeLines(tbl_latex, "saved_files/performance_by_muni.tex")

residual_df_validation <- residual_df %>% 
                          filter(validation) %>% 
                          mutate(pm20 = as.integer(abs(residual_sales_price_percent) <= 20))

### Various residual plots for test set
plot_validation_residuals_log <- ggplot(residual_df_validation,
                                        aes(x = sales_price / 1e6, 
                                            y = residual_sales_price_log)) +
                                 geom_point(alpha = 0.6, color = "#6BAED6", size = 0.25) +
                                 geom_smooth(aes(x = sales_price / 1e6, 
                                                 y = abs(residual_sales_price_log), 
                                                 linetype = "MALE"),
                                             se = FALSE, 
                                             color = "black", 
                                             size = 0.8, 
                                             inherit.aes = FALSE) +
                                scale_linetype_manual(name = NULL,
                                                      values = c(MALE = "solid"),
                                guide = guide_legend(override.aes = list(color = "black", size = 0.8),
                                                     nrow = 1, 
                                                     byrow = TRUE)) +
                                ylim(-1, 1) +
                                labs(x = "Sales price [million DKK]", 
                                     y = "Log residuals") +
                                theme(legend.position = c(0.5, 0.4))

saveRDS(plot_validation_residuals_log, "saved_files/plot_validation_residuals_log.rds")
print(plot_validation_residuals_log)



plot_validation_residuals_percent <- ggplot(residual_df_validation, 
                                            aes(x = sales_price / 1e6, 
                                                y = residual_sales_price_percent)) +
                                     geom_point(alpha = 0.6, color = "#6BAED6", size = 0.25) +
  
                                     # Absolute residuals line (MAPE) — inherit.aes = FALSE to decouple from color
                                     geom_smooth(aes(x = sales_price / 1e6, 
                                                     y = abs(residual_sales_price_percent), 
                                                     linetype = "MAPE"),
                                                 se = FALSE, 
                                                 color = "black", 
                                                 size = 0.8, 
                                                 inherit.aes = FALSE) +
  
                                     # pm20 probability line 
                                     geom_smooth(aes(x = sales_price / 1e6, 
                                                     y = pm20 * 100, 
                                                     linetype = "PM20"),
                                                 se = FALSE, 
                                                 color = "black", 
                                                 size = 0.8, 
                                                 inherit.aes = FALSE) +
                                     scale_linetype_manual(name = NULL,
                                                           values = c(MAPE = "solid", PM20 = "dashed"),
                                                           guide = guide_legend(override.aes = list(color = "black", 
                                                                                size = 0.8),
                                                           nrow = 1, byrow = TRUE)) +
                                     ylim(-100, 100) +
                                     labs(x = "Sales price [million DKK]", 
                                          y = "Percentage residuals [%]") +
                                     theme(legend.position = c(0.5, 0.4))


saveRDS(plot_validation_residuals_percent, "saved_files/plot_validation_residuals_percent.rds")
print(plot_validation_residuals_percent)

plot_validation_residuals <- ggplot(residual_df_validation, 
                                    aes(x = sales_price / 1e6, 
                                        y = residual_sales_price / 1e6)) +
                             geom_point(alpha = 0.6, color = "#6BAED6", size = 0.25) +
  
                             # MAE line, decoupled from color mapping
                             geom_smooth(aes(x = sales_price / 1e6, 
                                             y = abs(residual_sales_price) / 1e6, 
                                             linetype = "MAE"),
                                         se = FALSE, 
                                         color = "black", 
                                         size = 0.8, 
                                         inherit.aes = FALSE) +
                      
                             scale_linetype_manual(name = NULL,
                                                   values = c(MAE = "solid"),
                                                   guide = guide_legend(override.aes = list(color = "black", size = 0.8), 
                                                   nrow = 1, 
                                                   byrow = TRUE)) +
                             ylim(-3, 3) +
                             labs(x = "Sales price [million DKK]", 
                                  y = "Residuals [million DKK]") +
                             theme(legend.position = c(0.5, 0.4)) # inside top-right



saveRDS(plot_validation_residuals, "saved_files/plot_validation_residuals.rds")
print(plot_validation_residuals)

### Smoothed error plots for train, validation and test error

residual_df$type <- with(residual_df, ifelse(validation, "Validation error", "Training error"))


plot_log_errors <- ggplot(residual_df) + 
                   geom_smooth(aes(x = sales_date_numeric, 
                                   y = abs(residual_sales_price_log), 
                                   linetype = type),
                                   se = FALSE, 
                                   color = "black", 
                                   size = 0.8, 
                                   inherit.aes = FALSE) +
                   labs(x = "Sales date", 
                        y = "MALE") +
                   theme(legend.title = element_blank(),
                         legend.position = c(0.5, 0.4))


saveRDS(plot_log_errors, "saved_files/plot_log_errors.rds")
print(plot_log_errors)

plot_percentage_errors <- ggplot(residual_df) + 
                          geom_smooth(aes(x = sales_date_numeric, 
                                          y = 100*as.integer(abs(residual_sales_price_percent) <= 20), 
                                          linetype = type),
                                          se = FALSE, 
                                          color = "black", 
                                          size = 0.8, 
                                          inherit.aes = FALSE) +
                          labs(x = "Sales date", 
                               y = "PM20 [%]") +
                          theme(legend.title = element_blank(),
                                legend.position = c(0.5, 0.4))


saveRDS(plot_percentage_errors, "saved_files/plot_percentage_errors.rds")
print(plot_percentage_errors)

plot_errors <- ggplot(residual_df) + 
               geom_smooth(aes(x = sales_date_numeric, 
                               y = abs(residual_sales_price) / 1e6, 
                               linetype = type),
                               se = FALSE, 
                               color = "black", 
                               size = 0.8, 
                               inherit.aes = FALSE) +
               labs(x = "Sales date", 
                    y = "MAE [million DKK]") +
               theme(legend.title = element_blank(),
                     legend.position = c(0.5, 0.4))

saveRDS(plot_errors, "saved_files/plot_errors.rds")
print(plot_errors)

```


## 4. Model interpretation

```{r Interpretation, eval=TRUE, echo=TRUE, warning=FALSE}

make_interpretation_results <- function(model, model_name) {
  
  model_data <- model.frame(model)
  
  # generate fictitious properties
  grid <- data.frame(coordinate_east = c(697000, 462037, 469142, 725599),
                     coordinate_north = c(6150000, 6150225, 6322298, 6185920))
  
  point_constraints <- list()
  for (sm in model$smooth) {
    if (!is.null(sm$point.con)) {point_constraints <- c(point_constraints, sm$point.con)}
  }
  
  vars_in_model <- all.vars(formula(model))
  
  # fill missing variables in grid
  for (var in setdiff(vars_in_model, names(grid))) {
    
    if (is.factor(model_data[[var]])) {
      
      grid[[var]] <- factor(rep(levels(model_data[[var]])[1], nrow(grid)),
                            levels = levels(model_data[[var]]))
      
    } else {
      value <- if (var %in% names(point_constraints)) {
        point_constraints[[var]]
      } else {
        mean(model_data[[var]], na.rm = TRUE)
      }
      grid[[var]] <- rep(value, nrow(grid))
    }
  }
  
  # augment properties
  grid$coast_distance[2:4] <- 100
  grid$year_built[2:4] <- 2020
  grid$municipality_type[2:4] <- factor(c("Provincial", "Rural", "Capital"), levels = levels(model_data$municipality_type))
  grid$type <- c("Reference", "Esbjerg", "Nordjylland", "Strandvejen")
  
  terms <- as.data.frame(predict.gam(model, grid, type = "terms"))
  
  # helper for summing possibly missing columns
  sum_terms <- function(df, cols) {
    cols_present <- intersect(cols, names(df))
    if (length(cols_present) == 0) {return(rep(0, nrow(df)))}
    rowSums(df[, cols_present, drop = FALSE])
  }
  
  adjustments <- terms %>%
                 mutate(standard_price = exp(as.vector(coef(model))[1]),
      
                        adjustment_coordinates = exp(`s(coordinate_east,coordinate_north)`),
      
                        # Smooths with municipality_type interactions – 
                        #might as well sum over all is it only non-zero for the relevant smooth
                         
                        adjustment_sales_date = exp(`s(sales_date_numeric):municipality_typeCapital` +
                                                    `s(sales_date_numeric):municipality_typeRural` +
                                                    `s(sales_date_numeric):municipality_typeCommuter` +
                                                    `s(sales_date_numeric):municipality_typeProvincial` +
                                                    `s(sales_date_numeric):municipality_typeLarge city`),
      
                        adjustment_year_built = exp(`s(year_built):municipality_typeCapital` +
                                                    `s(year_built):municipality_typeRural` +
                                                    `s(year_built):municipality_typeCommuter` +
                                                    `s(year_built):municipality_typeProvincial` +
                                                    `s(year_built):municipality_typeLarge city`),
      
                        adjustment_year_rebuilt = exp(`s(year_rebuilt):municipality_typeCapital` +
                                                      `s(year_rebuilt):municipality_typeRural` +
                                                      `s(year_rebuilt):municipality_typeCommuter` +
                                                      `s(year_rebuilt):municipality_typeProvincial` +
                                                      `s(year_rebuilt):municipality_typeLarge city`),
      
                        adjustment_living_space = exp(`s(living_space):municipality_typeCapital` +
                                                      `s(living_space):municipality_typeRural` +
                                                      `s(living_space):municipality_typeCommuter` +
                                                      `s(living_space):municipality_typeProvincial` +
                                                      `s(living_space):municipality_typeLarge city`),
      
                        adjustment_lot_size = exp(`s(lot_size):municipality_typeCapital` +
                                                  `s(lot_size):municipality_typeRural` +
                                                  `s(lot_size):municipality_typeCommuter` +
                                                  `s(lot_size):municipality_typeProvincial` +
                                                  `s(lot_size):municipality_typeLarge city`),
                                              
                        adjustment_coast_distance = exp(`s(coast_distance)`),
          
                        adjustment_lake_distance = exp(`s(lake_distance)`),
          
                        adjustment_stream_distance = exp(`s(stream_distance)`),
          
                        adjustment_road_distance = exp(`s(road_distance)`),
          
                        adjustment_highway_distance = exp(`s(highway_distance)`),
          
                        adjustment_train_station_distance = exp(`s(train_station_distance)`),
          
                        adjustment_railway_distance = exp(`s(railway_distance)`),
          
                        adjustment_windturbine_distance = exp(`s(windturbine_distance)`),
          
                        adjustment_ocean_view = exp(`s(ocean_view)`),
          
                        adjustment_lake_view = exp(`s(lake_view)`),
          
                        adjustment_forest_size = exp(`s(forest_size)`),
          
                        adjustment_number_of_bathrooms = exp(`s(number_of_bathrooms):municipality_typeCapital` +
                                                             `s(number_of_bathrooms):municipality_typeRural` +
                                                             `s(number_of_bathrooms):municipality_typeCommuter` +
                                                             `s(number_of_bathrooms):municipality_typeProvincial` +
                                                             `s(number_of_bathrooms):municipality_typeLarge city`),
      
                        #Factor terms with interactions
                        adjustment_municipality_type = exp(municipality_type),
                        adjustment_house_type = exp(house_type + `house_type:municipality_type`),
                        adjustment_roof_type = exp(roof_type + `municipality_type:roof_type`),
                        adjustment_wall_type = exp(wall_type + `municipality_type:wall_type`),
                        adjustment_heating_type = exp(heating_type + `municipality_type:heating_type`),
 
                        adjustment_location = adjustment_coordinates *
                                              adjustment_sales_date *
                                              adjustment_lot_size *
                                              adjustment_coast_distance *
                                              adjustment_lake_distance *
                                              adjustment_stream_distance *
                                              adjustment_road_distance *
                                              adjustment_highway_distance *
                                              adjustment_train_station_distance *
                                              adjustment_railway_distance *
                                              adjustment_windturbine_distance *
                                              adjustment_ocean_view *
                                              adjustment_lake_view *
                                              adjustment_municipality_type,
      
                        adjustment_property = adjustment_year_built *
                                              adjustment_year_rebuilt *
                                              adjustment_living_space *
                                              adjustment_roof_type *
                                              adjustment_wall_type *
                                              adjustment_heating_type,
      
                        adjustment_total = adjustment_location * adjustment_property,
                        predicted_price = standard_price * adjustment_total) %>%
                        select(standard_price,
                               adjustment_municipality_type,
                               adjustment_coordinates,
                               adjustment_sales_date,
                               adjustment_year_built,
                               adjustment_year_rebuilt,
                               adjustment_living_space,
                               adjustment_lot_size,
                               adjustment_coast_distance,
                               adjustment_lake_distance,
                               adjustment_stream_distance,
                               adjustment_road_distance,
                               adjustment_highway_distance,
                               adjustment_train_station_distance,
                               adjustment_railway_distance,
                               adjustment_windturbine_distance,
                               adjustment_ocean_view,
                               adjustment_lake_view,
                               adjustment_forest_size,
                               adjustment_number_of_bathrooms,
                               adjustment_house_type,
                               adjustment_roof_type,
                               adjustment_wall_type,
                               adjustment_heating_type,
                               adjustment_location,
                               adjustment_property,
                               adjustment_total,
                               predicted_price)
  
  grid_out <- grid %>% select(coast_distance, year_built, municipality_type, type)
  
  bind_cols(grid_out, adjustments) %>% mutate(Model = model_name, .before = 1)
}

results_all_models <- bind_rows(
  lapply(names(models), function(model_name) {
    make_interpretation_results(models[[model_name]], model_name)
  })
)

saveRDS(results_all_models, "saved_files/results_interpretation_all_models.rds")

```



## 4. Uncertainty detection

The code below flags outliers based on both the GAM-based uncertainty estimates as well as the ensemble-based robustness measure. 

```{r Uncertainty detection, eval=TRUE, echo=TRUE, warning=FALSE, results = 'hide'}

predictions <- imap(models, function(model, name) {
  preds <- predict.gam(model, dataset_properties, type = "link", se.fit = TRUE)
  tibble(!!paste0("pred_", name) := as.vector(preds$fit),
         !!paste0("sigma_", name) := as.vector(preds$se.fit))
})

pred_matrix <- bind_cols(predictions)
property_predictions <- bind_cols(dataset_properties, pred_matrix)
saveRDS(property_predictions, "saved_files/property_predictions.rds")

pred_cols <- grep("^pred_", names(property_predictions), value = TRUE)
sigma_limit <- 0.2

property_predictions <- property_predictions %>%
                        rowwise() %>%
                        mutate(sigma_Ensemble = sd(c_across(all_of(pred_cols))), # ensemble spread
                               sigma_combined = pmax(sigma_Model1, sigma_Ensemble),   # max-rule
                               above_limit    = sigma_combined > sigma_limit) %>%
                        ungroup()

n_total <- nrow(property_predictions)
n_flagged <- sum(property_predictions$above_limit)
pct_flagged <- 100 * n_flagged / n_total


label_expr <- sprintf(
  "atop('Flagged: %s / %s (%.2f%%)', 'Rule: max(' * sigma[\"Model 1\"] * ', ' * sigma[Ensemble] * ') > %.2f')",
  format(n_flagged, big.mark = ","),
  format(n_total,   big.mark = ","),
  pct_flagged,
  sigma_limit
)

plot_uncertainties <- ggplot(property_predictions, aes(x = sigma_Model1, y = sigma_Ensemble)) +
                      geom_point(alpha = 0.1, size = 0.8, color = "#08306B") +
                      geom_point(data = subset(property_predictions, above_limit),
                                 shape = 21, fill = NA, color = "#08306B",
                                 size = 1.5, stroke = 0.7, alpha = 1.0) +
                      geom_density_2d(color = "grey80", linewidth = 0.6, alpha = 0.8) +
                      geom_vline(xintercept = sigma_limit, linetype = "dotted", linewidth = 0.7) +
                      geom_hline(yintercept = sigma_limit, linetype = "dotted", linewidth = 0.7) +
                      annotate("label",
                               x = 0.075, y = 0.24,
                               label = label_expr,
                               parse = TRUE,
                               hjust = 0, vjust = 1,
                               size = 3.5, label.size = 0.2) +
                      coord_equal(xlim = c(0, 0.5), ylim = c(0, 0.5), expand = FALSE) +
                      labs(x = expression(sigma["Model 1"]),
                           y = expression(sigma["Ensemble"]))

saveRDS(plot_uncertainties, "saved_files/plot_uncertainties.rds")
print(plot_uncertainties)

#Plot outliers geographically
outliers <- property_predictions %>% filter(above_limit)

plot_outliers_geographically <- ggplot() +
  
                                # Add land/ocean contour
                                geom_contour(data = coordinate_grid,
                                             aes(x = coordinate_east, 
                                                 y = coordinate_north, 
                                                 z = as.numeric(!ocean)),
                                             color = "black", size = 0.4) +
                                # Outliers
                                geom_point(data = outliers,
                                           aes(x = coordinate_east,
                                               y = coordinate_north,
                                               color = sigma_combined),
                                           size = 1.0, alpha = 0.7) +
                                scale_color_viridis_c(option = "plasma", name = expression(sigma["Combined"])) + 
                                coord_fixed() +
                                theme_void(base_size = 10) +
                                theme(plot.background = element_rect(fill = "white", color = NA),
                                      panel.background = element_rect(fill = "white", color = NA))

saveRDS(plot_outliers_geographically, "saved_files/plot_outliers_geographically.rds")
print(plot_outliers_geographically)


```




## 5. Model checking

### Standard Model Checks
The code below plots various standard diagnostics for Model 1 using the appraise() function from the gratia library as described in https://gavinsimpson.github.io/gratia/reference/appraise.html. 

```{r Standard model checks, eval=TRUE, echo=TRUE, warning=FALSE}
k.check(models$Model1)
plot_appraise <- appraise(models$Model1) 
saveRDS(plot_appraise, "saved_files/plot_appraise.rds")

```


### The Distributional Assumption
The code simulates a series of posterior distributions of the data based on Model 1 and compares it to the actual distribution of sales prices (see https://gavinsimpson.github.io/gratia/articles/posterior-simulation.html for details on posterior simulation with gratia).

```{r Posterior distribution, eval=TRUE, echo=TRUE, warning=FALSE, results = 'hide'}

n_draws <- 10
samples <-  posterior_samples(models$Model1, n = n_draws) %>%
            select(.draw, .response) %>%
            mutate(draw = paste("Draw", .draw),
                   type = "Posterior draws",
                   sales_price = .response) %>%
            select(draw, type, sales_price)

samples <- samples %>%
           group_by(draw) %>%
           ungroup()

data <- as.data.frame(list(type = "Empirical distribution",
                           draw = "Empirical distribution",
                           sales_price = models$Model1$y))


plot_distribution <- ggplot(data = data,
                            aes(sales_price / 1e6, group = draw, color = type)) +
                     geom_density(linewidth = 1.0, key_glyph = "path") +
                     geom_density(data = samples,
                                  aes(sales_price / 1e6, color = type),
                                      linewidth = 0.25,
                                      key_glyph = "path") +
                     xlim(0, 15) +
                     xlab('Sales price [million DKK]') +
                     ylab('Density') +
                     theme(legend.position = c(0.8, 0.9),
                           legend.text = element_text(size = 12)) +
                     scale_color_manual(name = NULL,
                                        values = c("Empirical distribution" = "black",
                                                   "Posterior draws" = "#F8766D"))

saveRDS(plot_distribution, "saved_files/plot_distribution.rds")
print(plot_distribution)


```




### Concurvity Analysis
The code investigates concurvity between the various smooths for Model 1 based on https://stat.ethz.ch/R-manual/R-devel/library/mgcv/html/concurvity.html. 

```{r Concurvity analysis, eval=TRUE, echo=TRUE, warning=FALSE, results = 'hide', fig.width=20, fig.height=30}

cc <- concurvity(models$Model1, full = FALSE)

collapse_concurvity <- function(conc_mat, pretty_names) {
  
  # remove self-concurvity first
  diag(conc_mat) <- NA
  
  # base smooth names (strip :municipality_typeXXXX)
  base_names <- sub(":municipality_type.*$", "", colnames(conc_mat))
  uniq_terms <- unique(base_names)
  
  # prepare new collapsed matrix
  conc_grouped <- matrix(
    NA_real_,
    nrow = length(uniq_terms),
    ncol = length(uniq_terms),
    dimnames = list(uniq_terms, uniq_terms)
  )
  
  # fill blocks
  for (i in seq_along(uniq_terms)) {
    for (j in seq_along(uniq_terms)) {
      rows <- base_names == uniq_terms[i]
      cols <- base_names == uniq_terms[j]
      block <- conc_mat[rows, cols, drop = FALSE]
      conc_grouped[i, j] <- max(block)
    }
  }
  
  # drop parametric part
  conc_grouped <- conc_grouped[
    rownames(conc_grouped) != "para",
    colnames(conc_grouped) != "para",
    drop = FALSE
  ]
  
  # apply nice labels
  rownames(conc_grouped) <- pretty_names[rownames(conc_grouped)]
  colnames(conc_grouped) <- pretty_names[colnames(conc_grouped)]
  
  as.data.frame(conc_grouped)
}

# 0) Your pretty names
pretty_names <- c("s(coordinate_east,coordinate_north)" = "Spatial",
                  "s(sales_date_numeric)" = "Sales date",
                  "s(living_space)" = "Living space",
                  "s(lot_size)" = "Lot size",
                  "s(number_of_bathrooms)" = "Number of bathrooms",
                  "s(year_built)" = "Year built",
                  "s(year_rebuilt)" = "Year rebuilt",
                  "s(coast_distance)" = "Coast distance",
                  "s(lake_distance)" = "Lake distance",
                  "s(stream_distance)" = "Stream distance",
                  "s(train_station_distance)" = "Train distance",
                  "s(railway_distance)" = "Railway distance",
                  "s(road_distance)" = "Road distance",
                  "s(highway_distance)" = "Highway distance",
                  "s(windturbine_distance)" = "Wind turbine distance",
                  "s(forest_size)" = "Forest size",
                  "s(ocean_view)" = "Ocean view",
                  "s(lake_view)" = "Lake view")

# 1) collapse concurvity 
conc_grouped_worst <- collapse_concurvity(cc$worst, pretty_names)
conc_grouped_observed <- collapse_concurvity(cc$observed, pretty_names)
conc_grouped_estimate <- collapse_concurvity(cc$estimate, pretty_names)

# 2) pheatmaps WITHOUT legends
concurvity_heatmap_worst <- pheatmap(conc_grouped_worst,
                                     cluster_rows = FALSE,
                                     cluster_cols = FALSE,
                                     na_col       = "black",
                                     fontsize_row = 12,
                                     fontsize_col = 12,
                                     labels_col   = rep("", ncol(conc_grouped_worst)),
                                     color        = viridis(100),
                                     legend       = FALSE,                          
                                     breaks       = seq(0, 1, length.out = 100),
                                     border_color = NA,
                                     silent       = TRUE)

concurvity_heatmap_observed <- pheatmap(conc_grouped_observed,
                                        cluster_rows = FALSE,
                                        cluster_cols = FALSE,
                                        na_col       = "black",
                                        fontsize_row = 12,
                                        fontsize_col = 12,
                                        labels_col   = rep("", ncol(conc_grouped_observed)),
                                        color        = viridis(100),
                                        legend       = FALSE,                         
                                        breaks       = seq(0, 1, length.out = 100),
                                        border_color = NA,
                                        silent       = TRUE)

concurvity_heatmap_estimate <- pheatmap(conc_grouped_estimate,
                                        cluster_rows = FALSE,
                                        cluster_cols = FALSE,
                                        na_col       = "black",
                                        fontsize_row = 12,
                                        fontsize_col = 12,
                                        color        = viridis(100),
                                        legend       = FALSE,                           
                                        breaks       = seq(0, 1, length.out = 100),
                                        border_color = NA,
                                        silent       = TRUE)

#one with legend true to be saved separately
concurvity_heatmap_est <- pheatmap(conc_grouped_estimate,
                                   cluster_rows = FALSE,
                                   cluster_cols = FALSE,
                                   na_col       = "black",
                                   fontsize_row = 12,
                                   fontsize_col = 12,
                                   color        = viridis(100),
                                   legend       = FALSE,                           
                                   breaks       = seq(0, 1, length.out = 100),
                                   border_color = NA,
                                   silent       = TRUE)

# 3) stack the three heatmaps

concurvity_heatmaps <- plot_grid(concurvity_heatmap_worst$gtable,
                                 concurvity_heatmap_observed$gtable,
                                 concurvity_heatmap_estimate$gtable,
                                 ncol = 1,
                                 align = "v",
                                 rel_heights = c(1, 1, 1.2))


legend_df <- data.frame(x = 1, y = 1, z = 0.5)  

legend_plot <- ggplot(legend_df, aes(x = x, y = y, fill = z)) +
               geom_tile() +
               scale_fill_viridis_c(limits = c(0, 1),
                                    breaks = c(0, 0.5, 1),
                                    labels = c("0", "0.5", "1"),
                                    name   = "Concurvity") +
               theme_void() +
               theme(legend.position   = "bottom",
                     legend.title      = element_text(size = 11),
                     legend.text       = element_text(size = 10),
                     legend.key.width  = unit(4.0, "cm"),
                     legend.key.height = unit(0.6, "cm"),
                     legend.margin = margin(t = -20, b = -20))

g <- ggplotGrob(legend_plot)
guide_ix  <- which(sapply(g$grobs, function(x) x$name) == "guide-box")
legend_grob <- g$grobs[[guide_ix]]

concurvity_heatmaps <- plot_grid(concurvity_heatmaps,
                                 legend_grob,
                                 ncol = 1,
                                 rel_heights = c(1, 0.04))

saveRDS(concurvity_heatmap_est, "saved_files/concurvity_heatmap_est.rds")
saveRDS(concurvity_heatmaps, "saved_files/concurvity_heatmaps.rds")
print(concurvity_heatmaps)


```


### Convergence of the Spatial Smooth
The code below fits the model ensemble for different basis set sizes of the spatial smooth to examine the convergence of the test error.

```{r Convergence basis set, eval=TRUE, echo=TRUE, warning=FALSE, results='hide'}
k_values <- c(100, 200, 500, 1000, 2000, 3000, 4000, 5000, 6000, 7000, 8000, 9000, 10000)
k_values <- k_values[k_values <= k_spatial]
  
model_grid_basis <- expand.grid(fold = 1:N, k_spatial = k_values)


if (load) {
  
  models_basis <- readRDS("saved_files/models_basis.rds")
  
} else {
  
  models_basis <- future_lapply(seq_len(nrow(model_grid_basis)), function(idx) {
    row <- model_grid_basis[idx, ]
    fit_fold(
      i = row$fold,
      data = dataset_sales_cv,
      k_spatial = row$k_spatial,
      k_other = k_other,
      fraction = 1.0,
      checkpoint_prefix = paste0("basis_k", row$k_spatial)
    )
  })
  
  names(models_basis) <- paste0("Model", model_grid_basis$fold, "_", model_grid_basis$k_spatial)
  saveRDS(models_basis, "saved_files/models_basis.rds")
}

get_residuals_grid <- function(models, dataset, grid_df, grid_name) {
  
  residual_list <- future_lapply(seq_along(models), function(i) {
    model_name <- names(models)[i]
    model <- models[[i]]
    row <- grid_df[i, ]
    fold <- row$fold
    
    predicted_sales_price <- predict(model, newdata = dataset, type = "response")
    
    tibble(sales_price = dataset$sales_price,
           predicted_sales_price = predicted_sales_price,
           residual_sales_price = predicted_sales_price - dataset$sales_price,
           residual_sales_price_percent = (predicted_sales_price - dataset$sales_price) / dataset$sales_price * 100,
           residual_sales_price_log = log(predicted_sales_price) - log(dataset$sales_price),
           validation = dataset$validation_set == fold,
           municipality_type = dataset$municipality_type,
           Model = model_name,
           value = row[[grid_name]])
  })
  
  bind_rows(residual_list)
}

residual_df_basis <- get_residuals_grid(models = models_basis,
                                        dataset = dataset_sales_cv,
                                        grid_df = model_grid_basis,
                                        grid_name = "k_spatial")

saveRDS(residual_df_basis, "saved_files/residual_df_basis.rds")

performance_basis <- residual_df_basis %>%
                     mutate(type = if_else(validation, "Validation error", "Training error")) %>%
                     group_by(type, municipality_type, value) %>%
                     summarise(MALE = mean(abs(residual_sales_price_log)),
                               .groups = "drop") %>%
                     rename(k = value)

plot_basis <- ggplot(performance_basis,
                     aes(x = k, y = MALE, color = municipality_type, lty = type)) +
              geom_line() +
              geom_point() +
              labs(x = "Number of spatial basis functions (k)",
                   y = "MALE",
                   color = "Municipality type")

saveRDS(plot_basis, "saved_files/plot_basis.rds")
print(plot_basis)

```



## Learning curves
The code below fits a list of models to increasing fractions of the training set and assesses their validation MALE. Learning curves for each municipality type are then plotted.

```{r Learning curves, eval=TRUE, echo=TRUE, warning=FALSE, results = 'hide'}
fractions <- c(0.1, 0.2, 0.5, 1.0)
folds <- 1:N

model_grid_fraction <- expand.grid(fold = folds, fraction = fractions)

if (load) {
  models_fraction <- readRDS("saved_files/models_fraction.rds")
} else {
  
  models_fraction <- future_lapply(seq_len(nrow(model_grid_fraction)), function(idx) {
    row <- model_grid_fraction[idx, ]
    
    fit_fold(
      i = row$fold,
      data = dataset_sales_cv,
      k_spatial = k_spatial,
      k_other = k_other,
      fraction = row$fraction,
      checkpoint_prefix = paste0("fraction_", row$fraction)
    )
  })
  
  names(models_fraction) <- paste0(
    "Model", model_grid_fraction$fold,
    "_", model_grid_fraction$fraction
  )
  
  saveRDS(models_fraction, "saved_files/models_fraction.rds")
}


get_residuals_fraction <- function(models, dataset, grid_df) {
  
  residual_list <- future_lapply(seq_along(models), function(i) {
    
    model_name <- names(models)[i]
    model <- models[[i]]
    row <- grid_df[i, ]
    fold <- row$fold
    frac <- row$fraction
    
    predicted_sales_price <- predict(model, newdata = dataset, type = "response")
    
    tibble(
      sales_price = dataset$sales_price,
      predicted_sales_price = predicted_sales_price,
      residual_sales_price = predicted_sales_price - dataset$sales_price,
      residual_sales_price_percent = (predicted_sales_price - dataset$sales_price) / dataset$sales_price * 100,
      residual_sales_price_log = log(predicted_sales_price) - log(dataset$sales_price),
      validation = dataset$validation_set == fold,
      municipality_type = dataset$municipality_type,
      Model = paste0("Model", fold),
      fraction = frac
    )
  })
  
  bind_rows(residual_list)
}

residual_df_fraction <- get_residuals_fraction(
  models = models_fraction,
  dataset = dataset_sales_cv,
  grid_df = model_grid_fraction
)

saveRDS(residual_df_fraction, "saved_files/residual_df_fraction.rds")

performance_fraction <- residual_df_fraction %>%
                        filter(validation) %>%
                        group_by(fraction, municipality_type) %>%
                        summarise(MALE = mean(abs(residual_sales_price_log)),
                                  .groups = "drop")

plot_learning_curves <- ggplot(performance_fraction,
                               aes(x = fraction, 
                                   y = MALE, 
                                   color = municipality_type)) +
                        geom_line(linewidth = 1) +
                        geom_point() +
                        scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
                        labs(x = "Percentage of training set",
                             y = "Validation MALE",
                             color = "Municipality type") +
                        theme_minimal()

saveRDS(plot_learning_curves, "saved_files/plot_learning_curves.rds")
print(plot_learning_curves)

```


## Market noise and other models
The code below estimates the market noise based on properties sold twice and compares this to the error obtained using our GAM model as well as an XGboost model. 

```{r Market noise, eval=TRUE, echo=TRUE, warning=FALSE, results = 'hide'}


### Identify double sales and discard those with changes to interior between sales
dataset_sales <- dataset_sales %>%
                 group_by(id) %>%
                 mutate(n = n()) %>%
                 ungroup() %>%
                 arrange(id)

other_sales <- dataset_sales %>% filter(n != 2)
double_sales <- dataset_sales %>% filter(n == 2)

double_sales <- double_sales %>%
                group_by(id) %>%
                filter(year_rebuilt < 2014,
                       n_distinct(living_space) == 1,
                       n_distinct(lot_size) == 1,
                       n_distinct(year_built) == 1,
                       n_distinct(house_type) == 1,
                       n_distinct(number_of_floors) == 1,
                       n_distinct(number_of_rooms) == 1,
                       n_distinct(number_of_bathrooms) == 1,
                       n_distinct(roof_type) == 1,
                       n_distinct(heating_type) == 1,
                       n_distinct(wall_type) == 1) %>%
                ungroup()


### Fit models to other sales and evaluate on double sales

# 1. Indexation model

if (load){

  time_model <- readRDS("saved_files/time_model.rds")

} else{

  time_model <- gam(sales_price ~ 
                    municipality_type + 
                    s(sales_date_numeric, k = k_other, by = municipality_type, pc = 2024),
                    data = dataset_sales,
                    family = tw(link=log),
                    method = "REML")

  saveRDS(time_model, "saved_files/time_model.rds")

}


time_model_terms <- as.data.frame(predict.gam(time_model, double_sales, type = "terms"))
time_adjustment <- as.vector(rowSums(time_model_terms[, grepl("sales_date_numeric", 
                                     colnames(time_model_terms)), 
                                     drop = FALSE]))

double_sales <- double_sales %>% mutate(log_sales_price = log(sales_price),
                                        log_sales_price_indexed = log_sales_price + time_adjustment,
                                        sales_price_indexed = exp(log_sales_price_indexed))

other_sales <- other_sales %>% mutate(log_sales_price = log(sales_price),
                                      log_sales_price_indexed = log_sales_price + time_adjustment,
                                      sales_price_indexed = exp(log_sales_price_indexed))

double_sales_lagged <- double_sales %>%
                       arrange(id, sales_date) %>%
                       group_by(id) %>%
                       mutate(sales_price_diff_indexed_log = log_sales_price_indexed - lag(log_sales_price_indexed),
                              avg_price = (sales_price_indexed + lag(sales_price_indexed)) / 2,
                              sales_date_diff = sales_date_numeric - lag(sales_date_numeric)) %>%
                       ungroup()


# 2. GAM model

if (load){

  gam_model <- readRDS("saved_files/gam_model.rds")

} else{

  gam_model <- bam(sales_price_indexed ~ 
                   s(coordinate_east, coordinate_north, k = k_spatial, pc = c(697000, 6150000)) +
                   s(year_built, k = k_other, by = municipality_type, pc = 1970) +
                   s(year_rebuilt, k = k_other, by = municipality_type, pc = 1970) +
                   s(living_space, k = k_other, by = municipality_type, pc = 140) +
                   s(lot_size, k = k_other, by = municipality_type, pc = 800) +
                   s(coast_distance, k = k_other, pc = 1500) +
                   s(lake_distance, k = k_other, pc = 500) +
                   s(stream_distance, k = k_other, pc = 1000) +
                   s(road_distance, k = k_other, pc = 2000) +
                   s(highway_distance, k = k_other, pc = 2000) +
                   s(train_station_distance, k = k_other, pc = 2000) +
                   s(railway_distance, k = k_other, pc = 1000) +
                   s(windturbine_distance, k = k_other, pc = 3000) +
                   s(ocean_view, k = k_other, pc = 0) +
                   s(lake_view, k = k_other, pc = 0) +
                   s(forest_size, k = k_other, pc = 0) +
                   s(number_of_bathrooms, k = 2, by = municipality_type, pc = 1) +
                   house_type*municipality_type +
                   roof_type*municipality_type +
                   wall_type*municipality_type +
                   heating_type*municipality_type,
                   data = other_sales,
                   family = tw(link=log),
                   method = "fREML")


  saveRDS(gam_model, "saved_files/gam_model.rds")
}

double_sales <- double_sales %>% 
                mutate(residuals_gam = 
                       log_sales_price_indexed - predict(gam_model, double_sales, type = "link"))

other_sales <- other_sales %>% 
               mutate(residuals_gam = 
                      log_sales_price_indexed - predict(gam_model, other_sales, type = "link"))


# 3. XGboost Models (fitted to log sales prices and GAM residuals)

dummy_model <- dummyVars(~ roof_type +
                           wall_type +
                           heating_type +
                           house_type +
                           municipality_type + 
                           municipality +
                           province +
                           region,
                         data = other_sales)

categorical_encoded <- predict(dummy_model, newdata = other_sales)

X <- other_sales %>%
     select(coordinate_east,
            coordinate_north,
            number_of_bathrooms,
            number_of_rooms,
            number_of_floors,
            year_built,
            year_rebuilt,
            living_space,
            lot_size,
            coast_distance,
            lake_distance,
            stream_distance,
            road_distance,
            highway_distance,
            train_station_distance,
            railway_distance,
            windturbine_distance,
            forest_size,
            ocean_view,
            lake_view) %>%
      as.matrix()

X_full <- cbind(X, categorical_encoded)

if (load){

  xgb_model <- readRDS("saved_files/xgb_model.rds")
  xgb_model_residuals <- readRDS("saved_files/xgb_model_residuals.rds")

} else{

  #on log sales price
  dtrain <- xgb.DMatrix(data = X_full, label = other_sales$log_sales_price_indexed)

  xgb_model <- xgboost(
    data = dtrain,
    objective = "reg:squarederror",
    nrounds = 1000,
    max_depth = 12,
    eta = 0.1,
    subsample = 0.8,
    colsample_bytree = 0.8,
    verbose = 1
  )

  saveRDS(xgb_model, "saved_files/xgb_model.rds")
  
  #on log residuals
  dtrain <- xgb.DMatrix(data = X_full, label = other_sales$residuals_gam)

  xgb_model_residuals <- xgboost(
    data = dtrain,
    objective = "reg:squarederror",
    nrounds = 1000,
    max_depth = 12,
    eta = 0.1,
    subsample = 0.8,
    colsample_bytree = 0.8,
    verbose = 1
  )

  saveRDS(xgb_model_residuals, "saved_files/xgb_model_residuals.rds")

}

categorical_encoded <- predict(dummy_model, newdata = double_sales)

X <- double_sales %>%
     select(coordinate_east,
            coordinate_north,
            number_of_bathrooms,
            number_of_rooms,
            number_of_floors,
            year_built,
            year_rebuilt,
            living_space,
            lot_size,
            coast_distance,
            lake_distance,
            stream_distance,
            road_distance,
            highway_distance,
            train_station_distance,
            railway_distance,
            windturbine_distance,
            forest_size,
            ocean_view,
            lake_view) %>%
      as.matrix()

X_full <- cbind(X, categorical_encoded)

double_sales <- double_sales %>% 
                mutate(residuals_xgb = log_sales_price_indexed - predict(xgb_model, X_full),
                       residuals_hybrid = residuals_gam - predict(xgb_model_residuals, X_full))


# 4. Noise model

if (load){

  noise_model <- readRDS("saved_files/noise_model.rds")

} else{

  noise_model <- gam(sales_price_diff_indexed_log ~ s(avg_price),
                     data = double_sales_lagged,
                     family = gaussian())


  saveRDS(noise_model, "saved_files/noise_model.rds")

}


### Make scatter plot with errors
double_sales_lagged$residuals_noise <- NA
double_sales_lagged$residuals_noise[which(!is.na(double_sales_lagged$sales_date_diff))] <- residuals(noise_model)/sqrt(2)

smooth_data <- bind_rows(double_sales_lagged %>% 
                         transmute(x = avg_price / 1e6, y = abs(residuals_noise), Model = "Market noise"),
                         double_sales %>%
                         transmute(x = sales_price / 1e6, y = abs(residuals_gam), Model = "GAM"),
                         double_sales %>%
                         transmute(x = sales_price / 1e6, y = abs(residuals_xgb), Model = "XGBoost"),
                         double_sales %>%
                         transmute(x = sales_price / 1e6, y = abs(residuals_hybrid), Model = "Hybrid model"))

plot_market_noise <- ggplot(data = double_sales_lagged,
                     aes(x = avg_price / 1e6, y = residuals_noise)) +
                     geom_point(size = 0.3) +
                     geom_smooth(data = smooth_data,
                                 aes(x = x, y = y, color = Model),
                                 method = "gam", formula = y ~ s(x),
                                 se = TRUE, 
                                 linewidth = 1) +
                     labs(x = "Average price [million DKK]",
                          y = "Log residuals",
                          color = "Model",
                          linetype = NULL) +
                     ylim(-1, 1) +
                     xlim(0, 10) +
                     theme_minimal() +
                     theme(text = element_text(size = 12),
                           legend.position = c(0.95, 0.65),
                           legend.title = element_text(size = 12),  
                           legend.text = element_text(size = 10))

saveRDS(plot_market_noise, "saved_files/plot_market_noise.rds")
print(plot_market_noise)



```


### Number of folds for model ensemble
The code below fits the model ensemble for numbers of fold for the model ensembles and examines the ensemble spread af a function of this.

```{r Number of folds, eval=TRUE, echo=TRUE, warning=FALSE, results='hide'}
if (load) {
  models_folds <- readRDS("saved_files/models_folds.rds")
  data_folds <- readRDS("saved_files/data_folds.rds")
} else {

  N_values <- c(2,3,4,6,8,10,20,50)

  data_folds <- lapply(N_values, function(N_current) {assign_cv_folds(dataset_sales, N = N_current, seed = 123)})
  
  names(data_folds) <- paste0("N", N_values)

  models_folds <- lapply(seq_along(N_values), function(j) {
    
    N_current <- N_values[j]
    data_current <- data_folds[[j]]

    models_current <- future_lapply(1:N_current, function(i) {
      
      fit_fold(i = i,
               data = data_current,
               k_spatial = k_spatial,
               k_other = k_other,
               fraction = 1.0,
               checkpoint_prefix = paste0("N", N_current))
      })

    names(models_current) <- paste0("Model", 1:N_current)
    models_current
    
    })
  
  
  names(models_folds) <- paste0("N", N_values)

  saveRDS(models_folds, "saved_files/models_folds.rds")
  saveRDS(data_folds, "saved_files/data_folds.rds")
}

get_ensemble_spread <- function(model_list, dataset, N_value) {

  predictions <- imap(model_list, function(model, name) {
    preds <- predict.gam(model, dataset, type = "link", se.fit = FALSE)
    tibble(!!paste0("pred_", name) := as.vector(preds))
  })

  pred_matrix <- bind_cols(predictions)
  pred_cols <- grep("^pred_", names(pred_matrix), value = TRUE)

  spread_df <- bind_cols(dataset, pred_matrix) %>%
               rowwise() %>%
               mutate(sigma_Ensemble = sd(c_across(all_of(pred_cols)))) %>%
               ungroup() %>%
               mutate(N = N_value)

  spread_df
}

property_spread_folds <- bind_rows(
  lapply(names(models_folds), function(name) {
    N_current <- as.numeric(sub("^N", "", name))
    get_ensemble_spread(
      model_list = models_folds[[name]],
      dataset = dataset_properties,
      N_value = N_current
    )
  })
)

saveRDS(property_spread_folds, "saved_files/property_spread_folds.rds")

spread_summary <- property_spread_folds %>%
                  group_by(N) %>%
                  summarise(q05 = quantile(sigma_Ensemble, 0.05, na.rm = TRUE),
                            median = median(sigma_Ensemble, na.rm = TRUE),
                            q95 = quantile(sigma_Ensemble, 0.95, na.rm = TRUE),
                            mean = mean(sigma_Ensemble, na.rm = TRUE),
                            .groups = "drop")

saveRDS(spread_summary, "saved_files/spread_summary_folds.rds")

plot_folds_spread <- ggplot(spread_summary, aes(x = N)) +
                     geom_ribbon(aes(ymin = q05, ymax = q95), alpha = 0.2, fill = "grey70") +
                     geom_line(aes(y = median), linewidth = 1) +
                     geom_point(aes(y = median), size = 2) +
                     labs(x = "Number of folds",
                          y = expression(sigma[Ensemble]),
                          title = "Ensemble spread across valuation properties by number of folds")

saveRDS(plot_folds_spread, "saved_files/plot_folds_spread.rds")
print(plot_folds_spread)

```


## Interplay between market noise and data scarcity
The code below examines the interplay between and noise and data scarcity by examining the validation MALE in the 2D space of KNN-averaged sales densities and sales prices. 

```{r Interplay, eval=FALSE, echo=TRUE, warning=FALSE, results = 'hide'}

# Parameters
K <- 100 #number of neighbors to average over
N <- 3 #number of clusters in each direction
log_scale <- TRUE #perform partitioning on log scale

# Split data
test_set <- dataset %>% filter(test_set)
training_set <- dataset %>% filter(validation_set != 1, !test_set)

# Predict log price
test_set$sales_price_predicted_log <- as.vector(
  predict.gam(models$Model1, test_set, type = "link")
)

# Find spatial neighbors
neighbors <- knn(
  data = training_set[, c("coordinate_east", "coordinate_north")],
  query = test_set[, c("coordinate_east", "coordinate_north")],
  k = K,
  radius = 0
)

distance_matrix <- neighbors$nn.dists
id_matrix <- neighbors$nn.idx

# Compute local sales densities
test_set$sales_density <- apply(distance_matrix, 1, function(dists) {
  K / (pi * (0.001 * max(dists))^2)
})

# Compute average sales price of neighbors
test_set$avg_sales_price <- vapply(
  1:nrow(test_set),
  function(j) mean(training_set$sales_price[id_matrix[j, ]]),
  numeric(1)
)

# Compute residuals in log space
test_set$resid_sales_price_log <- test_set$sales_price_predicted_log - log(test_set$sales_price)

#Find clusters
price_vec <- if (log_scale) log1p(test_set$avg_sales_price) else test_set$avg_sales_price
density_vec <- if (log_scale) log1p(test_set$sales_density) else test_set$sales_density
price_clusters <- kmeans(price_vec, centers = N)$cluster
density_clusters <- kmeans(density_vec, centers = N)$cluster
price_order <- rank(tapply(test_set$avg_sales_price, price_clusters, mean, na.rm = TRUE))
density_order <- rank(tapply(test_set$sales_density, density_clusters, mean, na.rm = TRUE))

test_set <- test_set %>%
            mutate(price_cluster = price_clusters,
                   density_cluster = density_clusters,
                   price_group = factor(price_order[price_cluster],
                   labels = c("Low price", "Medium price", "High price")),
                   density_group = factor(density_order[density_cluster],
                   labels = c("Low density", "Medium density", "High density")),
                   price_density_label = paste(price_group, density_group, sep = ", "))

plot_clusters <- ggplot(test_set, aes(x = sales_density, 
                                      y = avg_sales_price, 
                                      color = price_density_label, 
                                      shape = municipality_type)) +
                 geom_point(alpha = 0.6, size = 1.5) +
                 scale_color_brewer(palette = "Set1") +
                 scale_x_continuous(trans = "log1p") +
                 scale_y_continuous(trans = "log1p") +
                 scale_shape_discrete(name = "Municipality type") +  
                 labs(x = expression("Sales density [sales per km"^2*"]"),
                      y = "Average sales price [DKK]",
                      color = "Cluster") + 
                 theme_minimal()

saveRDS(plot_clusters, "saved_files/plot_clusters.rds")
print(plot_clusters)


#Plot clusters geographically
plot_clusters_geographically <- ggplot() +
                                # Add land/ocean contour
                                geom_contour(data = coordinate_grid,
                                             aes(x = coordinate_east, 
                                                 y = coordinate_north, 
                                                 z = as.numeric(!ocean)),
                                             color = "black", size = 0.4) +
                                # Add price-density group points
                                geom_point(data = test_set,
                                           aes(x = coordinate_east, 
                                               y = coordinate_north, 
                                               color = price_density_label),
                                           size = 0.5, alpha = 0.7) +
                                scale_color_brewer(palette = "Set1") +
                                coord_fixed() +
                                theme_void(base_size = 10) +
                                theme(legend.position = "none",
                                      plot.background = element_rect(fill = "white", color = NA),
                                      panel.background = element_rect(fill = "white", color = NA))

saveRDS(plot_clusters_geographically, "saved_files/plot_clusters_geographically.rds")
print(plot_clusters_geographically)



#Examine interplay between model performance and clusters
binned_data <- test_set %>%
               filter(!is.na(price_group), !is.na(density_group)) %>%
               group_by(price_group, density_group) %>%
               summarise(count = n(),
                         mean_abs_resid = mean(abs(resid_sales_price_log), na.rm = TRUE),
                         .groups = "drop")

plot_interplay <- ggplot(binned_data, aes(x = density_group, 
                                          y = price_group, 
                                          fill = mean_abs_resid)) +
                  geom_tile(color = "white") +
                  geom_text(aes(label = count), color = "black", size = 3) +
                  scale_fill_viridis_c(name = "MALE") +
                  labs(x = NULL,
                       y = NULL) + 
                  theme_minimal()

saveRDS(plot_interplay, "saved_files/plot_interplay.rds")
print(plot_interplay)

```

